### 2017-09-12 paper/main.tex
Add paper

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
\PassOptionsToPackage{numbers, compress}{natbib}
% before loading nips_2017
%
% to avoid loading the natbib package, add option nonatbib:
% \usepackage[nonatbib]{nips_2017}

%\usepackage{nips_2017}
\usepackage[final]{nips_2017}
\bibliographystyle{abbrvnat}

% to compile a camera-ready version, add the [final] option, e.g.:
%\usepackage[final]{nips_2017}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{graphicx}

\providecommand{\scan}{\text{SCAN}}
\providecommand{\reduce}{\text{REDUCE}}
% declaration of the new block
\algblock{ParFor}{EndParFor}
% customising the new block
\algnewcommand\algorithmicparfor{\textbf{parfor}}
\algnewcommand\algorithmicpardo{\textbf{do}}
\algnewcommand\algorithmicendparfor{\textbf{end\ parfor}}
\algrenewtext{ParFor}[1]{\algorithmicparfor\ #1\ \algorithmicpardo}
\algrenewtext{EndParFor}{\algorithmicendparfor}

\title{Parallelizing Linear Recurrent Neural Nets Over Sequence Length}

% The \author macro works with any number of authors. There are two
% commands used to separate the names and addresses of multiple
% authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to
% break the lines. Using \AND forces a line break at that point. So,
% if LaTeX puts 3 of 4 authors names on the first line, and the last
% on the second line, try using \AND instead of \And before the third
% author name.

\author{
  Eric Martin \\%\thanks{Use footnote for providing further
    %information about author (webpage, alternative
    %address)---\emph{not} for acknowledging funding agencies.} \\
  %Some Affiliation\\
  \texttt{eric@ericmart.in}
  %% examples of more authors
  \And
  Chris Cundy \\
  \texttt{c.cundy@berkeley.edu}
  %%\texttt{some@email.com}
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \AND
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}

\begin{document}
% \nipsfinalcopy is no longer used

\maketitle

\begin{abstract}
Recurrent neural networks (RNNs) are widely used to model sequential data 
but their non-linear dependencies between sequence elements prevent parallelizing
training over sequence length. We show the training of RNNs with only 
linear sequential dependencies can be parallelized over the
sequence length using the parallel scan algorithm, leading to rapid training
on long sequences with small minibatch size. We abstract prior linear sequence models
into a new framework of linear surrogate RNNs and develop
a linear surrogate long short-term memory (LS-LSTM) powered by a parallel linear recurrence
CUDA kernel we implemented.
We evaluate the LS-LSTM on a long sequence noisy autoregressive task and find
the LS-LSTM achieves slightly superior train and test performance to a similar 
sized LSTM in 4x less training time. We analyze latency and throughput of
the LS-LSTM and find the LS-LSTM reaches up to 
175x the throughput of the LSTM in the small minibatch long sequence regime.
\end{abstract}

\section{Introduction}

Recurrent neural networks (RNNs) are widely used for sequence modelling tasks in domains such as 
natural language processing \cite{sutskever2014sequence}, 
speech recognition \cite{amodei2015deep},
and reinforcement learning \cite{hausknecht2015deep}. Most RNNs, including popular variants such as long short-term memories (LSTMs) \cite{hochreiter1997long} and gated recurrent units (GRUs) \cite{cho2014learning}, contain a non-linear dependency
between sequential inputs. These non-linear dependencies create a very flexible class of
models but limit the feasibility of training RNNs on long sequences as each sequence element
must be processed sequentially.
Modelling sequences of thousands to millions of elements is important to domains
such as robotics, remote sensing, control systems, speech recognition, medicine, and finance.

The RNN serial evaluation inefficiency problem is usually mitigated by parallelizing the forward
and backward pass over a minibatch of inputs. Without minibatches, RNN evaluation is a sequence of matrix-vector multiplications. Minibatches transform RNN computation into a sequence of more efficient matrix-matrix multiplications, but minibatches within RNNs present many issues.
RNN model size is often limited by GPU memory size, and running a forward 
and backward pass on a minibatch requires memory linear in
the minibatch size.
Grouping data into minibatches increases the latency of each pass and reduces the rate of optimization steps. Finally, training with larger minibatches damages generalization ability 
\cite{keskar2017large}. Persistent RNNs \cite{diamos2016persistent} use a novel implementation that can achieve high GPU utilization with very small 
minibatch sizes when the recurrent state is larger than 500 elements, but even
persistent RNNs become limited by the serial evaluation
inefficiency at smaller hidden sizes.

Numerous prior works have shown strong performance from neural sequential models with only linear dependence on earlier sequence elements. \citet{balduzzi2016strongly} investigated RNNs with only elementwise linear recurrence relations 
$h_t = \alpha_t \odot h_{t-1} + (1-\alpha_t) \odot x_t$ and developed
linear variants of LSTM and GRU that perform similarly to
standard non-linear RNNs on text generation tasks. \citet{bradbury2017quasi},  \citet{kalchbrenner2016neural}, \citet{gehring2017convolutional}, and \citet{van2016wavenet} have successfully applied networks of convolutions over sequences for tasks
such as machine translation, language modelling, and audio generation. 
These works have observed up to an order of magnitude 
increase in training throughput compared to RNN alternatives. Convolutional sequence models typically rely on either an attention mechanism or a (possibly linear) recurrent layer
to integrate information at scales larger than the filter width. Introduction of a recurrent
layer prevents full parallelization over the sequence length while attention mechanisms are
expensive to apply on long sequences in online inference use cases.
One dimensional convolution can be viewed as a learnable linear
finite impulse response (FIR) filter with a parallel evaluation algorithm, while
linear recurrence is a learnable linear infinite impulse response (IIR). This work
parallelizes evaluation of linear recurrences through application of the parallel
scan algorithm.

Scans and reductions are computations involving repeated application of a binary
operator $\oplus$ over an array of data. Computing the sum or maximum
of an array is an example of a reduction, while a cumulative sum is a common
example of a scan operation. Throughout this work, the scan of $\oplus$ with
initial value $b$ is defined as
\begin{align*}
\scan(\oplus, [a_1, a_2, ..., a_n], b) = [(a_1 \oplus b), (a_2 \oplus a_1 \oplus b), ..., (a_n \oplus a_{n-1} ... \oplus a_1 \oplus b)]
\end{align*}
The reduction of $\oplus$ over array $A$ and initial value $b$ is denoted 
$\reduce(\oplus, A, b)$ and is the final element of $\scan(\oplus, A, b)$.
Despite their dependent computation graph, algorithms exist to parallelize scans 
and reductions when $\oplus$ is associative \cite{ladner1980parallel}.

\citet{blelloch1990prefix} shows that first order recurrences of the form
$h_t = (\Lambda_t \otimes h_{t-1}) \oplus x_t$ can be parallelized with
the parallel scan algorithm if three conditions are met:

\begin{enumerate}
\item $\oplus$ is associative: $(a \oplus b) \oplus c = a \oplus (b \oplus c)$
\item $\otimes$ is semiassociative: there exists a binary associative operator
$\odot$ such that $a \otimes (b \otimes c) = (a \odot b) \otimes c$
\item $\otimes$ distributes over $\oplus$: $a\otimes(b\oplus c) = (a\otimes b) \oplus (a \otimes c)$ 
\end{enumerate}

Our primary contribution is the application of the parallel linear recurrence algorithm to RNNs.

\section{Parallel linear recurrence}

Using elementwise vector addition as $x \oplus y= x+y$, matrix-vector multiplication as $A \otimes x=Ax$,
and matrix-matrix multiplication as $A \odot B=AB$ satisfies Blelloch's three conditions, thus allowing linear
recurrence $h_t = \Lambda_t h_{t-1} + x_t$ to be evaluated in parallel over $t$ for
vectors $x_t$ and square matrices $\Lambda_t$. As the method is
essential to this work, Algorithm 1 presents the parallel linear recurrence algorithm
for the interested reader.
\begin{algorithm}
\caption{Parallel linear recurrence on $p$ processors}
\begin{algorithmic}[1]
  \State Let $y = [(\Lambda_1, x_1), (\Lambda_2, x_2), ..., (\Lambda_T, x_T)]$
  \State Let binary operator $\bullet$ act as $(\Lambda, x) \bullet h = \Lambda h + x$
  \State Let $S_0=1, S_i < E_i, E_i + 1 = S_{i+1}, E_{p-1}=T$ for $i$ in $0,p-1$

  \\
  \ParFor{$i \gets 0,p-1$}
    \State $P_i = \reduce(\odot, \Lambda_{S_i:E_i}, I)$
    \State $R_i = \reduce(\bullet, y_{S_i:E_i}, 0)$
  \EndParFor

  \\
  \State Let $z = [(P_0, R_0), (P_1, R_1), ..., (P_p, R_p)]$.
  \State $C = \scan(\bullet, z, h_0)$   \Comment{compute $C_i = P_i C_{i-1} + R_i$ with $C_{-1}=h_0$}

  \\
  \ParFor{$i \gets 0,p-1$}
    \State $h_{S_i:E_i} = \scan(\bullet, y_{S_i:E_i}, C_{i-1})$
  \EndParFor

  \State \Return $h$
\end{algorithmic}
\end{algorithm}

\subsection{Theoretical performance}
The cost of a serial scan over a sequence of length $T$ is 
$C_\text{sscan} = (C_\otimes + C_\oplus)T$, compared to the parallel scan cost
$C_\text{pscan} = 2(C_\odot + C_\otimes + C_\oplus)(T/p + \lg p)$ \cite{blelloch1990prefix}.
If $h_t$ is a vector of dimension $n$ then 
$C_\odot=n^3, C_\otimes=n^2, C_\oplus=n$ giving
$C_\text{pscan} = 2(n^3 + n^2 + n)(T/p + \lg p)$ and 
$C_\text{sscan} = (n^2 + n)T$. The $n^3$ cost of the matrix
multiplication in the parallel algorithm can destroy any parallel speedups for
sufficiently large hidden states and lead to a slower algorithm overall.

To avoid this problem, we will only consider diagonal matrices $\Lambda_t$, in which
case matrix-matrix and matrix-vector have cost $n$ and 
$C_\text{pscan}=6n(T/p + \lg p)$ and $C_\text{sscan}=2nT$. Assuming $p \ll T$, then 
$C_\text{pscan} \le C_\text{sscan}$ when $p \ge 3$.
As we are only considering diagonal matrices, the
linear recurrence will be written $h_t = \lambda_t \odot h_{t-1} + x_t$ where
$\odot$ indicates elementwise multiplication.

Limiting $\Lambda_t$ to be diagonal may seem like a severe constraint but there are
several reasons to do so beyond the unfavorable parallelization performance. Relatively few neural
network models use separate recurrent matrices for each sequence element and using these
separate matrices would require potentially prohibitive $n^2T$ memory. Applying
the same matrix $\Lambda$ to each sequence element is also unappealing considering that a matrix
multiplication can be thought of as a rotation and a scaling. The same rotation at every
element seems unlikely to be useful, and the scaling is exactly what's captured in diagonal
vectors $\lambda_t$. Recurrent coefficient vectors $\lambda_t$ provide enough flexibility
to implement schemes such as exponential moving averages or a gating mechanism.

\subsection{Backpropagation}
\begin{align*}
\nabla_{h_T}L &= \frac{\partial L}{\partial h_T} \\
\nabla_{h_t}L &= \frac{\partial h_{t+1}}{\partial h_t} \odot \nabla_{h_{t+1}} L + \frac{\partial L}{\partial h_t} \\ 
&= \lambda_{t+1} \odot \nabla_{h_{t+1}} L + \frac{\partial L}{\partial h_t} \\
\nabla_{\lambda_t}L &= \frac{\partial h_t}{\partial\lambda_t} \odot \nabla_{h_t}L = h_{t-1} \odot \nabla_{h_t}L \\
\nabla_{x_t}L &= \nabla_{h_t} L \\
\nabla_{h_0}L &=  \frac{\partial h_1}{\partial h_0} \odot \nabla_{h_1} L = \lambda_1 \odot \nabla_{h_1} L
\end{align*}

The backpropagation equations center around a linear recurrence over $\frac{\partial L}{\partial h_t}$ in the reverse order of the original sequence. This allows for parallelizing both the forwards and backwards pass of a linear RNN over the sequence length.

\subsection{Implementation}
A modern high-end NVIDIA GPU consists of between 640 and 3200 concurrently
executing warps. Each warp operates on 32 single precision floating point numbers
in parallel.

This work implemented parallel linear recurrence as a CUDA kernel with
bindings into the TensorFlow \cite{abadi2016tensorflow} framework. Each warp acts as a processor, which means algorithmic $p$ is up to 3200 and the theoretical parallelization speedup factor is up to several hundred.
The 32 lanes of each warp work on different
elements of the recurrence vector in parallel. These implementation details mean that
peak performance is only obtained on sequences of at least several thousand
steps on at least a 32 element vector.

\section{Models}
Parallel linear recurrence can be used to construct a wide variety of differentiable modules that can be evaluated in parallel. Common applications of linear recurrence include gating schemes and exponential moving averages. Although linear recurrence values can depend only linearly on previous elements, the stacking of linear recurrent layers separated by non-linearities allows for a non-linear dependence on the past. In this sense the non-linear depth of a linear recurrent network is the number of layers and not the sequence length.

\subsection{Gated impulse linear recurrent layer}
A gated impulse linear recurrent (GILR) layer transforms its $m$ dimensional inputs $x_t$ into a sequence of $n$ dimensional hidden states $h_t$:
\begin{align*}
g_t &= \sigma(Ux_t + b_g) \\
i_t &= \tau(Vx_t + b_z) \\
h_t &= g_t \odot h_{t-1} + (1-g_t)\odot i_t
\end{align*}
A GILR layer applies the same non-linear transform to each sequence element and then accumulates the sequence elements with a non-linear gating mechanism. Gate $g_t$ uses the sigmoid activation function to have values in [0,1] for reasonable gating semantics, while impulse $i_t$ can use any activation function $\tau$. Stacking GILR layers allows for rich non-linear dependence on previous events while still taking advantage of fast parallel sequence evaluation.

\subsubsection{Impact on effective "batch size"}
Consider evaluating a vanilla RNN $h_t = \sigma(Uh_{t-1} + Vx_t + b)$ from $m$ inputs to $n$ hidden units on a sequence of length $T$ with minibatch size $b$ using a  serial evaluation strategy. At each of $T$ iterations, the naive approach performs two $(b, m) \mathbf{x} (m, n)$ matrix multiplications. Larger matrix multiplications achieve higher throughput due to less IO overhead, so the better approach computes $Vx_t$ for all $t$ ahead of time in a single $(bT, m) \mathbf{x} (m, n)$ matrix multiply. The non-linear recurrence forces even the better approach to perform $T$ potentially small $(b, m) \mathbf{x} (m, n)$ matrix multiplications in serial which makes performance heavily dependent on minibatch size.

Now consider the GILR, noting that it has the same two matrix-vector multiplications per iteration as the vanilla RNN. $g$ and $i$ can each be evaluated for all $t$ with a single $(bT, m) \mathbf{x} (m, n)$ matrix multiplication each. Given $g$ and $i$, $h$ can be computed using a parallel linear recurrence over $T$ vectors each of $bn$ elements. Rather than $T$ small operations, the GILR can be evaluated over all sequence elements with two large matrix multiplies and a parallel linear recurrence. GILR performance is much less dependent on batch size as the matrix multiplies see an "effective batch size" of $bT$ and $T$ is typically large.

\subsection{Linear surrogate RNNs}
RNNs learn a transition function $s_t = f(s_{t-1}, x_t)$ which combines previous state $s_{t-1}$ with input $x_t$ to compute current state $s_t$. Non-linear $f$ prevents application of the parallel linear recurrence algorithm and forces slow serial evaluation. To work around this inefficiency, note that $s_t$ serves a dual purpose. In $s_t = f(s_{t-1}, x_t)$, $s_{t-1}$ serves as an input to $f$ summarizing the previous inputs while $s_t$ serves as the output of $f$ to be passed to other layers of the network. If we decouple these uses, we can instead compute $s_t = f(\tilde{s}_{t-1}, x_t)$ with $\tilde{s}_t$ as a linearly computable surrogate for $s_t$. With this linear surrogate, non-linear $f$ can still be evaluated. We refer to this class of model as a linear surrogate RNN (LS-RNN). Quasi-RNNs \cite{bradbury2017quasi} are LS-RNNs using $\tilde{h}_{t-1} = W_k x_{t-k} + ... W_1 x_{t-1}$
and strongly typed RNNs\cite{balduzzi2016strongly} are LS-RNNs with $\tilde{h}_t=x_{t-1}$. Although not a rule, LS-RNNs can often be parallelized over sequence length with either
convolution or linear recurrence.

As an example LS-RNN, consider an LSTM:
\begin{align*}
f_t, i_t, o_t &= \sigma(U_{f,i,o} h_{t-1} + V_{f,i,o} x_t + b_{f,i,o}) \\
z_t &= \tau(U_z h_{t-1} + V_z x_t + b_z) \\
c_t &= f_t \odot c_{t-1} + i_t \odot z_t \\
h_t &= o_t \odot c_t
\end{align*}
An LSTM has state $s_t = (h_t, c_t)$. $c_t$ depends only linearly on $c_{t-1}$, so no surrogate is needed for $c_t$. $h_t$ has a non-linear dependence on $h_{t-1}$, so $h_t$ needs a linear surrogate. With a GILR layer as surrogate, the linear surrogate LSTM (LS-LSTM) is
\begin{align*}
f_t, i_t, o_t &= \sigma(U_{f,i,o} \tilde{h}_{t-1} + V_{f,i,o} x_t + b_{f,i,o}) \\
z_t &= \tau(U_z \tilde{h}_{t-1} + V_z x_t + b_z) \\
c_t &= f_t \odot c_{t-1} + i_t \odot z_t \\
h_t &= o_t \odot c_t \\
g_t &= \sigma(V_g x_t + b_g) \\
\tilde{h}_t &= g_t \odot \tilde{h}_{t-1} + (1-g_t)\odot \tau(Wx_t + c)
\end{align*}

For $m$ inputs and hidden size $n$, the LS-LSTM contains $2n(n+m)$ more parameters than the equivalently sized LSTM to handle the mapping from $x$ to $\tilde{h}$. More generally, a LS-RNN contains all of the same parameters as the underlying RNN as well as some additional parameters to compute the linear surrogate.

\section{Experiments}
Experiments were performed on the N-MNIST \cite{orchard2015converting} dataset. N-MNIST captures the MNIST digit dataset \cite{lecun1998mnist} by panning
an event driven camera over each digit. Each example in N-MNIST is a sequence of single pixel events (x, y, polarity, timestamp) where
x and y each indicate a pixel position in [0, 33], polarity indicates whether the pixel was switching on or off, and timestamp is the time of the event in microseconds. Videos produced from the event data show that positive polarity events are often located on the leading edge of the digit motion and negative polarity events on the trailing edge. Between the 60,000 digits in the training set, N-MNIST contains approximately 250 million pixel events with sequence lengths ranging from 500 to 8000 and averaging 4000 events.

We attempt to forecast 50 events ahead in N-MNIST using a two layer LSTM with 256 units per layer and a two layer LS-LSTM with 234 units per layer. The LS-LSTM layer size was selected so that it had slightly fewer parameters than the LSTM. Each model transformed the incoming event position into a 40-dimensional embedding vector which was then combined with the polarity to produce a 41 dimensional input. Both models output a $2 \mathbf{x} 34^2$ matrix containing two future event location probability distributions conditioned on the polarity of the future event. The training algorithm only considers the probability distribution of the true future polarity and uses the cross entropy loss function. The Adam \cite{kingma2014adam} optimization algorithm and Glorot \cite{glorot2010understanding} initialization scheme were used. Training on a minibatch of size $b$ with sequence length $T$ consisted of uniformly sampling $b$ N-MNIST sequences and then extracting a single random $T$ element subsequence (and its 50 element ahead forecast) from the sequence. Sequences less than $T$ elements were padded out to length $T$, and there was a "burn-in" of 30 events at the start of each sequence where no predictions were made. An epoch was defined as a pass over as many pixel events as there are in the full training set.

All experiments were performed using TensorFlow 1.0 \cite{abadi2016tensorflow} on a single Nvidia K80 GPU running for up to 18 hours. The LSTM model was computed using TensorFlow's dynamic\_rnn and BasicLSTMCell routines which are slower than but algorithmically similar to the cuDNN LSTM implementation.

\subsection{Computational performance}
The computational performance of the LSTM and LS-LSTM models were compared across a wide range of minibatch sizes and sequence lengths. Although the 234 unit LS-LSTM has a similar number of parameters to the 256 unit LSTM, we measured the computational performance of a 256 unit LS-LSTM to observe any impact of the linear surrogate calculation. We define throughput as events/s and a minibatch of size $b$ and length $T$ to be $bT$ events. The serial evaluation of the LSTM causes its throughput to be independent of the sequence length as a doubling of sequence length causes a doubling of runtime.

Figure \ref{fig:tp_perf} shows the LS-LSTM model achieves a throughput between 1.17x and 175x that of the LSTM, with the greatest relative advantages occurring at small minibatch sizes and long sequences. Notably, the LS-LSTM can achieve a higher throughput by running on one 8192 event sequence at a time than an LSTM running with 256 sequence minibatches. Similar speedups were found for inference. These speedups indicate the use of parallel linear recurrence through a linear surrogate can massively accelerate RNN training.

\begin{figure}[t]
\centering
\includegraphics[width=12cm]{heatmap.png}
\caption{Throughput comparison between LSTM-256-256 with LS-LSTM-256-256. The LSTM only has a single row of data because its throughput is independent of sequence length. Entries are missing from the LS-LSTM table because there was not enough memory on the GPU to handle such large batch sizes and sequences.}
\label{fig:tp_perf}
\end{figure}

\subsection{Training performance}
The speed of neural net training is not solely determined by the training throughput but also by the frequency of optimization steps. As an example, a batch method may process inputs at the same rate as a minibatch method, but the minibatch method will generally converge much faster on large datasets. On an infinite dataset, the batch method never takes a single optimization step regardless of its throughput. With this example in mind, it is clear that achieving fast training is a balancing act between training data throughput and optimization step latency and that training performance should be evaluated with training curves and not just throughput numbers such as time per epoch.

The LSTM and LS-LSTM offer differ latency and throughput tradeoffs. LSTM throughput depends only on minibatch size but LSTM latency depends on both minibatch size and sequence length. Experiments were conducted training the LSTM models with minibatch size 256 and sequence lengths 128, 256, 512, and 1024. Figure \ref{fig:learning_curves} shows the smaller sequence lengths led to faster initial learning but inferior final performance.

The throughput and latency of the LS-LSTM are both influenced by batch size and sequence length. Several LS-LSTMs were trained with batch size 16 and sequence length 1024. This combination was chosen because of the nearly maximum throughput, the low latency, and the ease of building a minibatch given the distribution of sequence lengths in the data set. The 234 unit LS-LSTM reaches a better training loss in roughly 4.5 hours than any of the LSTMs could reach in 18 hours. This experimental evidence indicates the LS-LSTM is as powerful of a model as the LSTM and can be trained in a fraction of the time.

\begin{figure}[t]
\centering
\includegraphics[width=12cm]{lc2.png}
\caption{The LSTM models had a direct relationship between sequence length and training curve: the shorter the sequence length, the faster the initial learning and
the higher the final loss. The faster initial learning is explained by the decreased
latency of optimization steps and the higher final loss is explained by the
inability to learn long dependencies. The LS-LSTM has 234 units per hidden
layer to have the same number of parameters as the 256 unit LSTMs.}
\label{fig:learning_curves}
\end{figure}

\subsection{Test performance}
\begin{table}[h]
  \caption{N-MNIST test results}
  \label{test-results}
  \centering
  \begin{tabular}{lllll}
    \toprule
    & Train Cross Entropy     & Test Cross Entropy     & Top-5 Accuracy & Top-20 Accuracy \\
    \midrule
    LS-LSTM 			&  \textbf{4.567} & 4.856           & 10.51\% & 36.70\%   \\
    LSTM, seqlen=128    & 4.664 & \textbf{4.846}  & \textbf{10.62\%} & \textbf{37.08\%}   \\
    LSTM, seqlen=256    & 4.620                & 4.906           & 10.11\% & 35.40\%   \\
    LSTM, seqlen=512    & 4.602                & 4.883           & 10.45\% & 36.32\%   \\
    LSTM, seqlen=1024   & 4.611                & 4.880           & 10.32\% & 35.92\%   \\
    \bottomrule
  \end{tabular}
\end{table}
Beyond cross entropy on train and test, we also evaluated top-$k$ accuracy for $k=5, 20$. A probability distribution is top-$k$ correct if it assigns the realized location one of the
$k$ largest probabilities. N-MNIST contains 1156 pixel positions, so top-5 and top-20 accuracy are equivalent to localizing the future to 0.43\% and 1.7\% of pixels.

No regularization was attempted and all of the models overfit, as indicated by the best test performance from the model with the worst train performance. Table \ref{test-results} contains the full test results. Our focus was the fast training of powerful models, and we leave regularizing parallel linear recurrences and LS-RNNs to future work. Although not tested, it is possible that the LS-LSTM generalized better than the LSTMs trained on long sequences due to the much smaller minibatch size of LS-LSTM leading to a wider minima.

\section{Conclusion}
Parallel linear recurrence is an extremely powerful algorithm and the LS-LSTM is just one
of many possible models that can be built with it. Future applications of parallel linear recurrence could include sequences orders of magnitude longer than N-MNIST, the development of parallel computable differentiable memory modules, and the combination of linear recurrence with convolutional sequence models. Besides future research, existing models such as Quasi-RNNs and strongly typed RNNs that already contain linear recurrences can immediately benefit from parallel linear recurrence. This work demonstrates the LS-LSTM significantly accelerates the training of small to medium sized LSTMs. Although similar techniques have been used before, the now explicit concept of linear surrogacy provides a framework for future development and analysis of fast sequence models.

We intend to expand upon the experiments section and open-source the parallel linear recurrence kernel in the near future.

\subsubsection*{Acknowledgments}
We would like to acknowledge Kevin Bowers, Alex Meiburg, JD Co-Reyes, Carson McNeil, and several others for fruitful conversations and guidance.

\bibliography{references}

\end{document}

### 2017-10-24 perf_plot.py
Add plotting code

@@ -1,0 +1,45 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import pandas as pd
+import seaborn as sns
+
+def to_df(d):
+    return pd.DataFrame([[k[0], k[1], v] for k, v in d.items()],
+                        columns=['bs', 'seqlen', 'val'])
+
+def plot(fast_tp, slow_tp):
+    # fast_tp and slow_tp are dicts from {(bs, seqlen): throughput} with throughput in steps/s
+    # it's ok if values are missing in the dicts. See Seaborn heatmap docs.
+    # for plot in pre-print, slow_tp is LSTM throughput and fast is LS-LSTM
+
+    # plot throughput values in tp_scale steps/s
+    tp_scale = 1000.
+
+    max_tp = max(max(fast_tp.values()), max(slow_tp.values())) / tp_scale
+
+    props = {'height_ratios': (.1, .9)}
+    f, ax = plt.subplots(2, gridspec_kw=props)
+    plt.subplots_adjust(left=.2, right=.85)
+
+    df = to_df(slow_tp)
+    df = df[df.seqlen == 512]    # only plot 1 row of slow throughput (because LSTM throughput doesn't depend on seqlen)
+    df = df.pivot('seqlen', 'bs', 'val') / tp_scale
+    sns.heatmap(df, ax=ax[0], annot=True, fmt='.3g', cbar=False, yticklabels=[''],
+                vmin=0, vmax=max_tp)
+    ax[0].yaxis.label.set_visible(False)
+    ax[0].set_xlabel('batch size')
+    ax[0].xaxis.set_label_coords(-.18, -.25)
+    ax[0].annotate('LSTM', xy=(1, 0), xytext=(1.03, .4), textcoords='axes fraction',
+                   fontsize=14)
+
+    df = to_df(fast_tp).pivot('seqlen', 'bs', 'seqlen') / tp_scale
+    sns.heatmap(df, ax=ax[1], annot=True, fmt='.3g', cbar=False, xticklabels=[''],
+                vmin=0, vmax=max_tp)
+    ax[1].set_ylabel('seq length', rotation=0)
+    ax[1].yaxis.set_label_coords(-.18, .485)
+    plt.setp(ax[1].yaxis.get_majorticklabels(), rotation=0)
+    ax[1].set_xlabel('')
+    ax[1].annotate('LS-LSTM', xy=(1, 0), xytext=(1.03, .485), textcoords='axes fraction',
+                   fontsize=14)
+
+    plt.show()

### 2017-10-26 paper/main.tex
Rewrote some of paper, added in QRNN implementation,

@@
 \begin{abstract}
-Recurrent neural networks (RNNs) are widely used to model sequential data 
-but their non-linear dependencies between sequence elements prevent parallelizing
-training over sequence length. We show the training of RNNs with only 
-linear sequential dependencies can be parallelized over the
-sequence length using the parallel scan algorithm, leading to rapid training
-on long sequences with small minibatch size. We abstract prior linear sequence models
-into a new framework of linear surrogate RNNs and develop
-a linear surrogate long short-term memory (LS-LSTM) powered by a parallel linear recurrence
-CUDA kernel we implemented.
-We evaluate the LS-LSTM on a long sequence noisy autoregressive task and find
-the LS-LSTM achieves slightly superior train and test performance to a similar 
-sized LSTM in 4x less training time. We analyze latency and throughput of
-the LS-LSTM and find the LS-LSTM reaches up to 
-175x the throughput of the LSTM in the small minibatch long sequence regime.
+Recurrent neural networks (RNNs) are widely used to model sequential data but
+their non-linear dependencies between sequence elements prevent parallelizing
+training over sequence length. We show the training of RNNs with only linear
+sequential dependencies can be parallelized over the sequence length using the
+parallel scan algorithm, leading to rapid training on long sequences with small
+minibatch size. We develop a parallel linear recurrence CUDA kernel and show
+that it can be applied to immediately speed up several RNN architectures by up
+to 9x. We abstract prior linear sequence models into a new framework of linear
+surrogate RNNs and develop a linear surrogate for the long short-term memory 
+unit, the LS-LSTM, which is able to achieve throughput 30x greater than the
+currently fastest-available LSTM implementation. We show that the LS-LSTM is
+able to perform well in the 
 \end{abstract}
@@
 
-Our primary contribution is the application of the parallel linear recurrence algorithm to RNNs.
+Our contributions are as follows:
+\begin{itemize}
+\item{We describe the class of RNNs which satisfy the conditions above, and
+    show that many of the RNNs used in practice such as the Quasi-RNN are
+    contained in this class}
+\item{We provide an implementation of the parallel linear recurrence
+    algorithm as a CUDA kernel, and show that it speeds up training of
+    the SRU and QRNN architectures by factors of up to 9x}
+\item{We describe how several recent linear RNNs can be described as
+    linear surrogates for nonlinear architectures. We introduce a linear
+    surrogate for the LSTM and show that we are able to train it
+    with a speedup of 5-10x compared to the CuDNN-LSTM when we use the plr
+    algorithm.}
+\end{itemize}
 
 \section{Parallel linear recurrence}
-
-Using elementwise vector addition as $x \oplus y= x+y$, matrix-vector multiplication as $A \otimes x=Ax$,
-and matrix-matrix multiplication as $A \odot B=AB$ satisfies Blelloch's three conditions, thus allowing linear
-recurrence $h_t = \Lambda_t h_{t-1} + x_t$ to be evaluated in parallel over $t$ for
-vectors $x_t$ and square matrices $\Lambda_t$. As the method is
-essential to this work, Algorithm 1 presents the parallel linear recurrence algorithm
-for the interested reader.
+Considering the familiar operations in linear algebra, we see that the
+associative operation of vector addition ($\boldsymbol{x} \oplus \boldsymbol{y}
+= \boldsymbol{x} + \boldsymbol{y}$), the semiassociative operation of
+matrix-vector multiplication ($A \otimes \boldsymbol{x} = A\boldsymbol{x}$) and
+the associative operation of matrix-matrix multiplication ($A \odot B=AB$) form
+recurrences which satisfy Blelloch's three conditions, thus allowing the linear
+recurrence $h_t = \Lambda_t h_{t-1} + x_t$ to be evaluated in parallel over $t$
+for vectors $x_t$ and square matrices $\Lambda_t$. As the method is essential to
+this work, Algorithm 1 presents the parallel linear recurrence algorithm for the
+interested reader.
 \begin{algorithm}
+  \label{alg:plr}
 \caption{Parallel linear recurrence on $p$ processors}
@@
 $C_\text{sscan} = (C_\otimes + C_\oplus)T$, compared to the parallel scan cost
-$C_\text{pscan} = 2(C_\odot + C_\otimes + C_\oplus)(T/p + \lg p)$ \cite{blelloch1990prefix}.
+$C_\text{pscan} = 2(C_\odot + C_\otimes + C_\oplus)(T/p + \lg p)$ \cite{blelloch1990prefix},
+where \(p\) is the number of available parallel cores. 
 If $h_t$ is a vector of dimension $n$ then 
-$C_\odot=n^3, C_\otimes=n^2, C_\oplus=n$ giving
-$C_\text{pscan} = 2(n^3 + n^2 + n)(T/p + \lg p)$ and 
-$C_\text{sscan} = (n^2 + n)T$. The $n^3$ cost of the matrix
-multiplication in the parallel algorithm can destroy any parallel speedups for
+$C_\odot\propto n^3, C_\otimes\propto n^2, C_\oplus\propto n$ giving
+$C_\text{pscan} \propto 2(n^3 + n^2 + n)(T/p + \lg p)$ and 
+$C_\text{sscan} \propto (n^2 + n)T$. The $n^3$ cost of the matrix
+multiplication in the parallel algorithm can counter-act any parallel speedups for
 sufficiently large hidden states and lead to a slower algorithm overall.
 
-To avoid this problem, we will only consider diagonal matrices $\Lambda_t$, in which
-case matrix-matrix and matrix-vector have cost $n$ and 
-$C_\text{pscan}=6n(T/p + \lg p)$ and $C_\text{sscan}=2nT$. Assuming $p \ll T$, then 
-$C_\text{pscan} \le C_\text{sscan}$ when $p \ge 3$.
-As we are only considering diagonal matrices, the
-linear recurrence will be written $h_t = \lambda_t \odot h_{t-1} + x_t$ where
-$\odot$ indicates elementwise multiplication.
+To avoid this problem, we will only consider diagonal matrices $\Lambda_t$, in
+which case both matrix-matrix and matrix-vector multiplication have cost
+proportional to $n$ and $C_\text{pscan}\propto 6n(T/p + \lg p)$ and
+$C_\text{sscan}=2nT$. Assuming $p \ll T$, then $C_\text{pscan} \le
+C_\text{sscan}$ when $p \ge 3$.\footnote{Can include the actual speedup
+  factor, \(\alpha = \frac{PT}{3(T + \lg(p))}\).}
+As we are only considering diagonal matrices,
+we write the linear recurrence as $h_t = \lambda_t \odot h_{t-1} + x_t$
+where $\odot$ indicates elementwise multiplication.
 
@@
 network models use separate recurrent matrices for each sequence element and using these
-separate matrices would require potentially prohibitive $n^2T$ memory. Applying
+separate matrices would require potentially prohibitive $n^2T$ memory.\footnote{Ideally we can
+  find a citation for someone thinking these points before}
+Applying
 the same matrix $\Lambda$ to each sequence element is also unappealing considering that a matrix
@@
 \subsection{Implementation}
-A modern high-end NVIDIA GPU consists of between 640 and 3200 concurrently
-executing warps. Each warp operates on 32 single precision floating point numbers
-in parallel.
+In 2017, GPUs commonly used in deep learning consist of between 640 and 3200
+concurrently executing warps. Each warp operates on 32 single precision floating
+point numbers in parallel.
 
-This work implemented parallel linear recurrence as a CUDA kernel with
-bindings into the TensorFlow \cite{abadi2016tensorflow} framework.
-Each warp acts as a processor, which means algorithmic $p$ is up to 3200 and the theoretical parallelization speedup factor is up to several hundred.
-The 32 lanes of each warp work on different
-elements of the recurrence vector in parallel. These implementation details mean that
-peak performance is only obtained on sequences of at least several thousand
-steps on at least a 32 element vector.
+This work implemented parallel linear recurrence as a CUDA kernel with bindings
+into the TensorFlow \cite{abadi2016tensorflow} framework. Each warp acts as a
+processor, which means the algorithmic parallelizability $p$ is up to 3200 and the theoretical
+parallelization speedup factor is up to several hundred.  The 32 lanes of each
+warp work on different elements of the recurrence vector in parallel. These
+implementation details mean that peak performance is only obtained on sequences
+of at least several thousand steps on at least a 32 element vector.
 
@@
 \subsubsection{Impact on effective "batch size"}
-Consider evaluating a vanilla RNN $h_t = \sigma(Uh_{t-1} + Vx_t + b)$ from $m$ inputs to $n$ hidden units on a sequence of length $T$ with minibatch size $b$ using a  serial evaluation strategy.
-At each of $T$ iterations, the naive approach performs two $(b, m) \mathbf{x} (m, n)$ matrix multiplications.
-Larger matrix multiplications achieve higher throughput due to less IO overhead, so the better approach computes $Vx_t$ for all $t$ ahead of time in a single $(bT, m) \mathbf{x} (m, n)$ matrix multiply.
-The non-linear recurrence forces even the better approach to perform $T$ potentially small $(b, m) \mathbf{x} (m, n)$ matrix multiplications in serial which makes performance heavily dependent on minibatch size.
+Consider evaluating an RNN with recurrence $h_t = \sigma(Uh_{t-1} + Vx_t + b)$,
+from $m$ inputs to $n$ hidden units on a sequence of length $T$ with minibatch
+size $b$ using a serial evaluation strategy. At each of $T$ iterations, the
+naive approach performs two $(b, m) \times (m, n)$ matrix
+multiplications. Larger matrix multiplications achieve higher throughput due to
+less IO overhead, so the better approach computes $Vx_t$ for all $t$ ahead of
+time in a single $(bT, m) \times (m, n)$ matrix multiply. The non-linear
+recurrence forces even the better approach to perform $T$ potentially small $(b,
+m) \times (m, n)$ matrix multiplications in serial which makes performance
+heavily dependent on minibatch size.
 
-Now consider the GILR, noting that it has the same two matrix-vector multiplications per iteration as the vanilla RNN.
-$g$ and $i$ can each be evaluated for all $t$ with a single $(bT, m) \mathbf{x} (m, n)$ matrix multiplication each.
-Given $g$ and $i$, $h$ can be computed using a parallel linear recurrence over $T$ vectors each of $bn$ elements.
-Rather than $T$ small operations, the GILR can be evaluated over all sequence elements with two large matrix multiplies and a parallel linear recurrence.
-GILR performance is much less dependent on batch size as the matrix multiplies see an "effective batch size" of $bT$ and $T$ is typically large.
+Now consider the GILR, noting that it has the same two matrix-vector
+multiplications per iteration as the above RNN. $g$ and $i$ can each be
+evaluated for all $t$ with a single $(bT, m) \times (m, n)$ matrix
+multiplication each. Given $g$ and $i$, $h$ can be computed using a parallel
+linear recurrence over $T$ vectors each of $bn$ elements. Rather than $T$ small
+operations, the GILR can be evaluated over all sequence elements with two large
+matrix multiplies and a parallel linear recurrence. GILR performance is much
+less dependent on batch size as the matrix multiplies see an "effective batch
+size" of $bT$ and $T$ is typically large.
 
 \subsection{Linear surrogate RNNs}
-RNNs learn a transition function $s_t = f(s_{t-1}, x_t)$ which combines previous state $s_{t-1}$ with input $x_t$ to compute current state $s_t$.
-Non-linear $f$ prevents application of the parallel linear recurrence algorithm and forces slow serial evaluation.
-To work around this inefficiency, note that $s_t$ serves a dual purpose.
-In $s_t = f(s_{t-1}, x_t)$, $s_{t-1}$ serves as an input to $f$ summarizing the previous inputs while $s_t$ serves as the output of $f$ to be passed to other layers of the network.
-If we decouple these uses, we can instead compute $s_t = f(\tilde{s}_{t-1}, x_t)$ with $\tilde{s}_t$ as a linearly computable surrogate for $s_t$.
-With this linear surrogate, non-linear $f$ can still be evaluated.
-We refer to this class of model as a linear surrogate RNN (LS-RNN).
-Quasi-RNNs \cite{bradbury2017quasi} are LS-RNNs using $\tilde{h}_{t-1} = W_k x_{t-k} + ...
-W_1 x_{t-1}$
-and strongly typed RNNs\cite{balduzzi2016strongly} are LS-RNNs with $\tilde{h}_t=x_{t-1}$. Although not a rule, LS-RNNs can often be parallelized over sequence length with either
+\label{sec:ls-rnns}
+RNNs learn a transition function $s_t = f(s_{t-1}, x_t)$ which combines previous
+state $s_{t-1}$ with input $x_t$ to compute current state $s_t$. Non-linear $f$
+prevents application of the parallel linear recurrence algorithm and forces slow
+serial evaluation. To work around this inefficiency, note that $s_t$ serves a
+dual purpose. In $s_t = f(s_{t-1}, x_t)$, $s_{t-1}$ serves as an input to $f$
+summarizing the previous inputs while $s_t$ serves as the output of $f$ to be
+passed to other layers of the network. If we decouple these uses, we can instead
+compute $s_t = f(\tilde{s}_{t-1}, x_t)$ with $\tilde{s}_t$ as a linearly
+computable surrogate for $s_t$. With this linear surrogate, non-linear $f$ can
+still be evaluated. We refer to this class of model as a linear surrogate RNN
+(LS-RNN). Quasi-RNNs \cite{bradbury2017quasi} are LS-RNNs using $\tilde{h}_{t-1}
+= W_k x_{t-k} + ... W_1 x_{t-1}$ and strongly typed
+RNNs\cite{balduzzi2016strongly} are LS-RNNs with $\tilde{h}_t=x_{t-1}$. Although
+not a rule, LS-RNNs can often be parallelized over sequence length with either
 convolution or linear recurrence.
 
-As an example LS-RNN, consider an LSTM:
+As an example LS-RNN, consider an LSTM\footnote{Want a citation here I guess}:
 \begin{align*}
@@
 An LSTM has state $s_t = (h_t, c_t)$.
-$c_t$ depends only linearly on $c_{t-1}$, so no surrogate is needed for $c_t$.
+Since $c_t$ depends only linearly on $c_{t-1}$, no surrogate is needed for $c_t$.
 $h_t$ has a non-linear dependence on $h_{t-1}$, so $h_t$ needs a linear surrogate.
@@
 g_t &= \sigma(V_g x_t + b_g) \\
-\tilde{h}_t &= g_t \odot \tilde{h}_{t-1} + (1-g_t)\odot \tau(Wx_t + c)
+\tilde{h}_t &= g_t \odot \tilde{h}_{t-1} + (1-g_t)\odot \tau(Wx_t + b_h)
 \end{align*}
@@
 \section{Experiments}
-Experiments were performed on the N-MNIST \cite{orchard2015converting} dataset. N-MNIST captures the MNIST digit dataset \cite{lecun1998mnist} by panning
-an event driven camera over each digit. Each example in N-MNIST is a sequence of single pixel events (x, y, polarity, timestamp) where
-x and y each indicate a pixel position in [0, 33], polarity indicates whether the pixel was switching on or off, and timestamp is the time of the event in microseconds.
+We perform several experiments, first confirming that our implementation of the
+parallel linear recurrence algorithm is able to achieve higher throughput than a
+similarly implemented serial version of the algorithm. We then show that the
+LS-LSTM architecture, trained with the PLR algorithm, is able to outperform the
+optimized CuDNN LSTM implementation on a pathological example from the original
+LSTM paper. Finally we show that our LS-LSTM is able to quickly achieve state of
+the art results on an existing medical data set.
+
+\subsection{Throughput Benchmarks}
+**** Throughput of pure PLR kernel vs SLR kernel
+**** Show how it speeds up the SRU, QRNN, and LS-LSTM
+**** Show how the LS-LSTM has much better throughput than the CudnnLSTM
+
+We first illustrate the throughput advantage of the parallel scan algorithm for
+evaluating the linear recurrence. We define throughput as events/s and a
+minibatch of size $b$ and length $T$ to be $bT$ events. We implement two CUDA
+kernels, one which evaluates the parallel linear recurrence described in
+algorithm \ref{alg:plr}, and one which evaluates the linear recurrence using the
+straightforward serial approach. This comparison is performed directly at the
+kernel level, avoiding any overhead due to calling Tensorflow. We find that at
+long sequence lengths, the PLR kernel has a distinct
+advantage, as shown in table \ref{table:kernel-throughput}.
+
+\begin{figure}
+\begin{center}
+  \begin{tabular}{lrrr}
+    \label{table:rnn-throughput}
+Sequence Length\footnote{What's the batch size etc here??} & 4 Dimensions & 32 Dimensions & 128 Dimensions \\
+\hline
+16 & 0.06 & 0.06 & 0.05\\ 
+256 & 0.22 & 0.22 & 0.86\\ 
+4,096 & 1.02 & 2.94 & 3.36\\ 
+65,536 & 38.5 & 41.8 & 17.5\\
+  \end{tabular}  
+    \end{center}
+    \caption{Speedup of the PLR kernel compared to the SLR kernel, without any
+    overhead from Tensorflow.}    
+  \end{figure}
+
+Many RNN architectures that have been recently introduced are amenable to
+implementing through parallel linear recurrence. Table
+\ref{table:rnn-throughput} shows the throughput advantage by using a parallel
+linear recurrence when compared to a serial linear recurrence. Since PLR and SLR
+are simply different methods for computing the same output, there is no penalty
+incurred for this speedup: it is a pure speed gain. Typically experiments will aim
+to maximize the memory utilisation of the GPU, so we simulate this by choosing the
+product of the batch size and the sequence length to be constant. We see that the
+architectures which have fewer terms (for which the linear recurrence is a higher
+proportion of the computational burden) are more effected by the switch to PLR.
+This is particularly clear in the case of the QRNN, where including a wider
+convolutional window results in less time spent on the recurrence and less to gain
+from the parallelization. 
+\begin{figure}
+\begin{center}
+  \begin{tabular}{lrrrr}
+    \label{table:rnn-throughput}
+Sequence Length & LS-LSTM & SRU & QRNN(2) & QRNN(10)\\
+\hline
+16 & 0.61 & 0.28 & 0.38 & 0.78\\
+256 & 0.91 & 0.84 & 0.86 & 0.99\\
+4,096 & 0.98 & 1.38 & 1.18 & 1.05\\
+65,536 & 1.41 & 9.21 & 6.68 & 2.05\\
+  \end{tabular}  
+\end{center}
+\caption{Speedup obtained when using parallel linear recurrence compared
+    to serial linear recurrence. We analyse this for several recent architectures:
+    the Simple Recurrent Unit from [cite], the Quasi-RNN from [cite], with a
+    convolutional filter of size 2 and 10, and the Linear Surrogate LSTM that
+    we introduce in section \ref{sec:ls-rnns}. To give a fair model of a
+    typical use-case, we set batch size \(b\) such that \(Tb = 65536\)
+    for a sequence length \(T\).}
+\end{figure}
+
+Finally, we compare the throughput of the linear surrogate LSTM to the current
+fastest available implementation of the LSTM, the CuDNN-LSTM. We observe that our
+LS-LSTM has higher throughput for most of the low-batch-size regime, with a 30x
+speedup at a sequence length of 65,536. Recent work (cite https://arxiv.org/pdf/1609.04836.pdf)
+has suggested that small mini-batches may aid generalization by encouraging
+the discovery of shallow minima. The LS-LSTM allows us to achieve high throughput
+for all choices of minibatch size and sequence length.
+
+\ref{fig:tp_perf}.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% The computational performance of the LSTM and LS-LSTM models were compared       %
+% across a wide range of minibatch sizes and sequence lengths. Although the 234    %
+% unit LS-LSTM has a similar number of parameters to the 256 unit LSTM, we         %
+% measured the computational performance of a 256 unit LS-LSTM to observe any      %
+% impact of the linear surrogate calculation. The serial evaluation of the LSTM    %
+% causes its throughput to be independent of the sequence length as a doubling of  %
+% sequence length causes a doubling of runtime. Notably, the LS-LSTM can achieve a %
+% higher throughput by running on one 8192 event sequence at a time than an LSTM   %
+% running with 256 sequence minibatches. Similar speedups were found for           %
+% inference.                                                                       %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Unlike the previous case, these speedups do come at a cost. The LS-LSTM
+architecture is different to the canonical LSTM, and it's possible that the loss
+of model capacity outweighs the speedup in training. The following experiment
+shows that the LS-LSTM is able to outperform the LSTM on an example originally
+designed to illustrate the advantages of the LSTM.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=12cm]{cudnn_heatmap.png}
+\caption{Throughput comparison between LSTM-256-256 with LS-LSTM-256-256, with a
+  32-dimensional input and a 2-dimensional output, for various batch sizes and
+  sequence lengths.The LSTM only has a single row of data because its throughput
+  is independent of sequence length. Entries are missing from the LS-LSTM table
+  because there was not enough memory on the GPU to handle such large batch
+  sizes and sequences.}
+\label{fig:tp_perf}
+\end{figure}
+
+% Experiments were performed on the N-MNIST \cite{orchard2015converting} dataset. N-MNIST captures the MNIST digit dataset \cite{lecun1998mnist} by panning
+% an event driven camera over each digit. Each example in N-MNIST is a sequence of single pixel events (x, y, polarity, timestamp) where
+% x and y each indicate a pixel position in [0, 33], polarity indicates whether the pixel was switching on or off, and timestamp is the time of the event in microseconds.
 Videos produced from the event data show that positive polarity events are often located on the leading edge of the digit motion and negative polarity events on the trailing edge.
@@
 
-We attempt to forecast 50 events ahead in N-MNIST using a two layer LSTM with 256 units per layer and a two layer LS-LSTM with 234 units per layer.
+% We attempt to forecast 50 events ahead in N-MNIST using a two layer LSTM with 256 units per layer and a two layer LS-LSTM with 234 units per layer.
 The LS-LSTM layer size was selected so that it had slightly fewer parameters than the LSTM.
@@
 
-All experiments were performed using TensorFlow 1.0 \cite{abadi2016tensorflow} on a single Nvidia K80 GPU running for up to 18 hours.
+% All experiments were performed using TensorFlow 1.0 \cite{abadi2016tensorflow} on a single Nvidia K80 GPU running for up to 18 hours.
 The LSTM model was computed using TensorFlow's dynamic\_rnn and BasicLSTMCell routines which are slower than but algorithmically similar to the cuDNN LSTM implementation.
 
-\subsection{Computational performance}
-The computational performance of the LSTM and LS-LSTM models were compared across a wide range of minibatch sizes and sequence lengths.
-Although the 234 unit LS-LSTM has a similar number of parameters to the 256 unit LSTM, we measured the computational performance of a 256 unit LS-LSTM to observe any impact of the linear surrogate calculation.
-We define throughput as events/s and a minibatch of size $b$ and length $T$ to be $bT$ events.
-The serial evaluation of the LSTM causes its throughput to be independent of the sequence length as a doubling of sequence length causes a doubling of runtime.
-
-Figure \ref{fig:tp_perf} shows the LS-LSTM model achieves a throughput between 1.17x and 175x that of the LSTM, with the greatest relative advantages occurring at small minibatch sizes and long sequences.
-Notably, the LS-LSTM can achieve a higher throughput by running on one 8192 event sequence at a time than an LSTM running with 256 sequence minibatches.
-Similar speedups were found for inference.
-These speedups indicate the use of parallel linear recurrence through a linear surrogate can massively accelerate RNN training.
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=12cm]{heatmap.png}
-\caption{Throughput comparison between LSTM-256-256 with LS-LSTM-256-256.
-The LSTM only has a single row of data because its throughput is independent of sequence length.
-Entries are missing from the LS-LSTM table because there was not enough memory on the GPU to handle such large batch sizes and sequences.}
-\label{fig:tp_perf}
-\end{figure}
-
-\subsection{Training performance}
-The speed of neural net training is not solely determined by the training throughput but also by the frequency of optimization steps.
+% \subsection{Training performance}
+% The speed of neural net training is not solely determined by the training throughput but also by the frequency of optimization steps.
 As an example, a batch method may process inputs at the same rate as a minibatch method, but the minibatch method will generally converge much faster on large datasets.
@@
 
-The LSTM and LS-LSTM offer differ latency and throughput tradeoffs.
+% The LSTM and LS-LSTM offer differ latency and throughput tradeoffs.
 LSTM throughput depends only on minibatch size but LSTM latency depends on both minibatch size and sequence length.
@@
 
-The throughput and latency of the LS-LSTM are both influenced by batch size and sequence length.
+% The throughput and latency of the LS-LSTM are both influenced by batch size and sequence length.
 Several LS-LSTMs were trained with batch size 16 and sequence length 1024.
@@
 
-\begin{figure}[t]
-\centering
-\includegraphics[width=12cm]{lc2.png}
-\caption{The LSTM models had a direct relationship between sequence length and training curve: the shorter the sequence length, the faster the initial learning and
-the higher the final loss. The faster initial learning is explained by the decreased
-latency of optimization steps and the higher final loss is explained by the
-inability to learn long dependencies. The LS-LSTM has 234 units per hidden
-layer to have the same number of parameters as the 256 unit LSTMs.}
-\label{fig:learning_curves}
-\end{figure}
+% \begin{figure}[t]
+% \centering
+% \includegraphics[width=12cm]{lc2.png}
+% \caption{The LSTM models had a direct relationship between sequence length and training curve: the shorter the sequence length, the faster the initial learning and
+% the higher the final loss. The faster initial learning is explained by the decreased
+% latency of optimization steps and the higher final loss is explained by the
+% inability to learn long dependencies. The LS-LSTM has 234 units per hidden
+% layer to have the same number of parameters as the 256 unit LSTMs.}
+% \label{fig:learning_curves}
+% \end{figure}
 
-\subsection{Test performance}
-\begin{table}[h]
-  \caption{N-MNIST test results}
-  \label{test-results}
-  \centering
-  \begin{tabular}{lllll}
-    \toprule
-    & Train Cross Entropy     & Test Cross Entropy     & Top-5 Accuracy & Top-20 Accuracy \\
-    \midrule
-    LS-LSTM 			&  \textbf{4.567} & 4.856           & 10.51\% & 36.70\%   \\
-    LSTM, seqlen=128    & 4.664 & \textbf{4.846}  & \textbf{10.62\%} & \textbf{37.08\%}   \\
-    LSTM, seqlen=256    & 4.620                & 4.906           & 10.11\% & 35.40\%   \\
-    LSTM, seqlen=512    & 4.602                & 4.883           & 10.45\% & 36.32\%   \\
-    LSTM, seqlen=1024   & 4.611                & 4.880           & 10.32\% & 35.92\%   \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-Beyond cross entropy on train and test, we also evaluated top-$k$ accuracy for $k=5, 20$. A probability distribution is top-$k$ correct if it assigns the realized location one of the
-$k$ largest probabilities. N-MNIST contains 1156 pixel positions, so top-5 and top-20 accuracy are equivalent to localizing the future to 0.43\% and 1.7\% of pixels.
+% \subsection{Test performance}
+% \begin{table}[h]
+%   \caption{N-MNIST test results}
+%   \label{test-results}
+%   \centering
+%   \begin{tabular}{lllll}
+%     \toprule
+%     & Train Cross Entropy     & Test Cross Entropy     & Top-5 Accuracy & Top-20 Accuracy \\
+%     \midrule
+%     LS-LSTM 			&  \textbf{4.567} & 4.856           & 10.51\% & 36.70\%   \\
+%     LSTM, seqlen=128    & 4.664 & \textbf{4.846}  & \textbf{10.62\%} & \textbf{37.08\%}   \\
+%     LSTM, seqlen=256    & 4.620                & 4.906           & 10.11\% & 35.40\%   \\
+%     LSTM, seqlen=512    & 4.602                & 4.883           & 10.45\% & 36.32\%   \\
+%     LSTM, seqlen=1024   & 4.611                & 4.880           & 10.32\% & 35.92\%   \\
+%     \bottomrule
+%   \end{tabular}
+% \end{table}
+% Beyond cross entropy on train and test, we also evaluated top-$k$ accuracy for $k=5, 20$. A probability distribution is top-$k$ correct if it assigns the realized location one of the
+% $k$ largest probabilities. N-MNIST contains 1156 pixel positions, so top-5 and top-20 accuracy are equivalent to localizing the future to 0.43\% and 1.7\% of pixels.
 
-No regularization was attempted and all of the models overfit, as indicated by the best test performance from the model with the worst train performance.
+% No regularization was attempted and all of the models overfit, as indicated by the best test performance from the model with the worst train performance.
 Table \ref{test-results} contains the full test results.
@@
 Although not tested, it is possible that the LS-LSTM generalized better than the LSTMs trained on long sequences due to the much smaller minibatch size of LS-LSTM leading to a wider minima.
+
+\subsection{Synthetic Example}
+In order to demonstrate how PLR may be used to speed up tasks that the LSTM is
+well-suited for, we tackle the pathological example 2 from cite Hochreiter and
+Schmidhtbuer which involves storing a number for a very long sequence of time
+steps. We have an alphabet of size \(p\) with each character represented as a
+particular one-hot vector in this space. We input a sequence of \(n\) of these
+vectors, chosen at random (with replacement). The first vector in the sequence
+is always the same, up to the sign of the component (i.e. it is \(\pm p_0
+\).
+
+The two signs on the first component separate the sequences into two
+sets. The whole sequence is fed into the LSTM and we aim to learn to classify
+the sequences. This requires remembering the first element over the length of
+the sequence. Due to the large throughput of the LS-LSTM, we would expect that
+it would do better when the sequence length is large. In the original
+formulation of the problem, \(p\) is set equal to \(n\). Since this would make
+the size of the input data grow impracticaly large as \(\mathcal{O}(n^2)\) for
+long sequences, we fix \(p = 128\) and vary \(n\).
+
+We generated three sets of data: for \(n\) equal to 1,024, 8,192, and
+1,048,576. For each of these we compared a two-layer LS-LSTM with 512 hidden
+units to a canonical LSTM network with roughly the same number of parameters. We
+could imagine using a two-layer LSTM with 512 hidden units in each layer, or a
+one-layer LSTM with 1024 hidden units.
+
+Initial experiments showed that the two-layer LSTM was much quicker to converge
+than the one-layer LSTM, so here we compare the LS-LSTM to the two-layer
+CudnnLSTM.  We ran all experiments on a NVIDIA K80 GPU, choosing the largest
+minibatch size which fit into the GPU memory, with five runs per configuration
+allowing us to find the average and standard deviation of the time and number of
+iterations to convergence. For all CUDA runs, a brief search over learning rate
+and batch size was carried out to find the parameters which allow the network to
+converge most rapidly. The criterion for convergence was five consecutive
+minibatches giving 100\% accuracy. As can be seen from the learning curves
+in figure \ref{fig:synthetic_training}, this was a reasonable criterion.
+\begin{figure}
+  \centering
+  \begin{minipage}{0.5\textwidth}
+    \includegraphics[width=1.0\textwidth]{./1k_synthetic.png}
+  \end{minipage}%
+    \begin{minipage}{0.5\textwidth}
+    \includegraphics[width=1.0\textwidth]{./8k_synthetic.png}      
+  \end{minipage}
+  \begin{minipage}{0.6\textwidth}
+    \includegraphics[width=1.0\textwidth]{./1M_synthetic.png}      
+  \end{minipage}
+  \caption{Training curves for LS-LSTM and CUDNNLSTM architectures for various
+  sequence lengths.}
+    \label{fig:synthetic_training}
+  \end{figure}
+
+\begin{table}[]
+\centering
+\caption{Performance of the LS-LSTM compared to CUDA-optimised CudnnLSTM
+  implementation on problem 2b from (cite '97). We see that the LS-LSTM has a clear
+  advantage in training speed as the sequence length increases.
+  {\scriptsize *For the longest sequence length, the number of hidden units was decreased to
+    64 for both architectures so that the net could fit in memory. }    }
+\label{my-label} 
+\begin{tabular}{@{}lllllll@{}} \toprule Sequence Length &
+\multicolumn{2}{c}{\textbf{1,024}} & \multicolumn{2}{c}{\textbf{8,192}} &
+\multicolumn{2}{c}{\textbf{1,048,576\(^*\)}} \\ \midrule & CudnnLSTM & LS-LSTM &
+CudnnLSTM & LS-LSTM & CudnnLSTM & LS-LSTM \\ \cmidrule(l){2-7} Iterations (1000s) &
+                     1.0 %bs=8
+                     \(\pm\) 0.4 & 0.55 %bs=8
+                                    \(\pm\) 0.04 & 0.44 %bs=8
+                                                   \(\pm\) 0.05 & 0.56 %bs=8
+                                                                 \(\pm\) 0.16
+                                                               & & 14 \(\pm\) 3 \\
+\begin{tabular}[c]{@{}l@{}}Wall Clock time (hours)\end{tabular} &
+0.278 \(\pm\) 0.083 & 0.0314 \(\pm\) 0.0019 & 0.58 \(\pm\) 0.06 & 0.10 \(\pm\) 0.03
+& & 9.7 \(\pm\) 1.7 \\ \bottomrule
+\end{tabular}
+\end{table}
+
+\subsection{Medical Example}
+A common application of RNNs is in medicine. In order to illustrate the increased
+performance that the plr algorithm allows, we implement an LS-LSTM on the problem
+presented in \texttt{CITE}, the 2016 PhysioNet/Computing in Cardiology Challenge 2016.
+The aim of the Classification of Normal/Abnormal Heart Sound Recordings challenge
+is to determine which of a set of EEGs are produced by people with abnormal hearts.
+
+The training set of data is unbalanced, with 75\% of the data corresponding to abnormal
+hearts.
+
+Key points are:
+\begin{itemize}
+\item{With the fast training of the LS-LSTM / PLR, we can train quickly on very
+    long sequence-lengths}
+\item{Allows us to get state-of-the-art results, even when performing very quickly
+    and with little hyperparameter tuning}
+\end{itemize}
+
+NEED TO GET THOSE RESULTS!
 
@@
 \subsubsection*{Acknowledgments}
-We would like to acknowledge Kevin Bowers, Alex Meiburg, JD Co-Reyes, Carson McNeil, and several others for fruitful conversations and guidance.
+We would like to acknowledge Kevin Bowers, Alex Meiburg, JD Co-Reyes, Carson
+McNeil, Andy Palan, and several others for fruitful conversations and guidance.
 
-\bibliography{references}
+\begin{thebibliography}{19}
+\providecommand{\natexlab}[1]{#1}
+\providecommand{\url}[1]{\texttt{#1}}
+\expandafter\ifx\csname urlstyle\endcsname\relax
+  \providecommand{\doi}[1]{doi: #1}\else
+  \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi
+
+\bibitem[Abadi et~al.(2016)Abadi, Agarwal, Barham, Brevdo, Chen, Citro,
+  Corrado, Davis, Dean, Devin, et~al.]{abadi2016tensorflow}
+M.~Abadi, A.~Agarwal, P.~Barham, E.~Brevdo, Z.~Chen, C.~Citro, G.~S. Corrado,
+  A.~Davis, J.~Dean, M.~Devin, et~al.
+\newblock Tensorflow: Large-scale machine learning on heterogeneous distributed
+  systems.
+\newblock \emph{arXiv preprint arXiv:1603.04467}, 2016.
+
+\bibitem[Amodei et~al.(2015)Amodei, Anubhai, Battenberg, Case, Casper,
+  Catanzaro, Chen, Chrzanowski, Coates, Diamos, et~al.]{amodei2015deep}
+D.~Amodei, R.~Anubhai, E.~Battenberg, C.~Case, J.~Casper, B.~Catanzaro,
+  J.~Chen, M.~Chrzanowski, A.~Coates, G.~Diamos, et~al.
+\newblock Deep speech 2: End-to-end speech recognition in english and mandarin.
+\newblock \emph{arXiv preprint arXiv:1512.02595}, 2015.
+
+\bibitem[Balduzzi and Ghifary(2016)]{balduzzi2016strongly}
+D.~Balduzzi and M.~Ghifary.
+\newblock Strongly-typed recurrent neural networks.
+\newblock In \emph{Proceedings of The 33rd International Conference on Machine
+  Learning}, pages 1292--1300, 2016.
+
+\bibitem[Blelloch(1990)]{blelloch1990prefix}
+G.~E. Blelloch.
+\newblock Prefix sums and their applications.
+\newblock 1990.
+
+\bibitem[Bradbury et~al.(2017)Bradbury, Merity, Xiong, and
+  Socher]{bradbury2017quasi}
+J.~Bradbury, S.~Merity, C.~Xiong, and R.~Socher.
+\newblock Quasi-recurrent neural networks.
+\newblock In \emph{International Conference on Learning Representations
+  (ICLR)}, 2017.
+
+\bibitem[Cho et~al.(2014)Cho, Van~Merri{\"e}nboer, Gulcehre, Bahdanau,
+  Bougares, Schwenk, and Bengio]{cho2014learning}
+K.~Cho, B.~Van~Merri{\"e}nboer, C.~Gulcehre, D.~Bahdanau, F.~Bougares,
+  H.~Schwenk, and Y.~Bengio.
+\newblock Learning phrase representations using rnn encoder-decoder for
+  statistical machine translation.
+\newblock \emph{arXiv preprint arXiv:1406.1078}, 2014.
+
+\bibitem[Diamos et~al.(2016)Diamos, Sengupta, Catanzaro, Chrzanowski, Coates,
+  Elsen, Engel, Hannun, and Satheesh]{diamos2016persistent}
+G.~Diamos, S.~Sengupta, B.~Catanzaro, M.~Chrzanowski, A.~Coates, E.~Elsen,
+  J.~Engel, A.~Hannun, and S.~Satheesh.
+\newblock Persistent rnns: Stashing recurrent weights on-chip.
+\newblock In \emph{International Conference on Machine Learning}, pages
+  2024--2033, 2016.
+
+\bibitem[Gehring et~al.(2017)Gehring, Auli, Grangier, Yarats, and
+  Dauphin]{gehring2017convolutional}
+J.~Gehring, M.~Auli, D.~Grangier, D.~Yarats, and Y.~N. Dauphin.
+\newblock Convolutional sequence to sequence learning.
+\newblock \emph{arXiv preprint arXiv:1705.03122}, 2017.
+
+\bibitem[Glorot and Bengio(2010)]{glorot2010understanding}
+X.~Glorot and Y.~Bengio.
+\newblock Understanding the difficulty of training deep feedforward neural
+  networks.
+\newblock In \emph{Aistats}, volume~9, pages 249--256, 2010.
+
+\bibitem[Hausknecht and Stone(2015)]{hausknecht2015deep}
+M.~Hausknecht and P.~Stone.
+\newblock Deep recurrent q-learning for partially observable mdps.
+\newblock In \emph{2015 AAAI Fall Symposium Series}, 2015.
+
+\bibitem[Hochreiter and Schmidhuber(1997)]{hochreiter1997long}
+S.~Hochreiter and J.~Schmidhuber.
+\newblock Long short-term memory.
+\newblock \emph{Neural computation}, 9\penalty0 (8):\penalty0 1735--1780, 1997.
+
+\bibitem[Kalchbrenner et~al.(2016)Kalchbrenner, Espeholt, Simonyan, Oord,
+  Graves, and Kavukcuoglu]{kalchbrenner2016neural}
+N.~Kalchbrenner, L.~Espeholt, K.~Simonyan, A.~v.~d. Oord, A.~Graves, and
+  K.~Kavukcuoglu.
+\newblock Neural machine translation in linear time.
+\newblock \emph{arXiv preprint arXiv:1610.10099}, 2016.
+
+\bibitem[Keskar et~al.(2017)Keskar, Mudigere, Nocedal, Smelyanskiy, and
+  Tang]{keskar2017large}
+N.~S. Keskar, D.~Mudigere, J.~Nocedal, M.~Smelyanskiy, and P.~T.~P. Tang.
+\newblock On large-batch training for deep learning: Generalization gap and
+  sharp minima.
+\newblock 2017.
+
+\bibitem[Kingma and Ba(2014)]{kingma2014adam}
+D.~Kingma and J.~Ba.
+\newblock Adam: A method for stochastic optimization.
+\newblock \emph{arXiv preprint arXiv:1412.6980}, 2014.
+
+\bibitem[Ladner and Fischer(1980)]{ladner1980parallel}
+R.~E. Ladner and M.~J. Fischer.
+\newblock Parallel prefix computation.
+\newblock \emph{Journal of the ACM (JACM)}, 27\penalty0 (4):\penalty0 831--838,
+  1980.
+
+\bibitem[LeCun et~al.(1998)LeCun, Cortes, and Burges]{lecun1998mnist}
+Y.~LeCun, C.~Cortes, and C.~J. Burges.
+\newblock The mnist database of handwritten digits, 1998.
+
+\bibitem[Orchard et~al.(2015)Orchard, Jayawant, Cohen, and
+  Thakor]{orchard2015converting}
+G.~Orchard, A.~Jayawant, G.~Cohen, and N.~Thakor.
+\newblock Converting static image datasets to spiking neuromorphic datasets
+  using saccades.
+\newblock \emph{arXiv preprint arXiv:1507.07629}, 2015.
+
+\bibitem[Sutskever et~al.(2014)Sutskever, Vinyals, and
+  Le]{sutskever2014sequence}
+I.~Sutskever, O.~Vinyals, and Q.~V. Le.
+\newblock Sequence to sequence learning with neural networks.
+\newblock In \emph{Advances in neural information processing systems}, pages
+  3104--3112, 2014.
+
+\bibitem[van~den Oord et~al.(2016)van~den Oord, Dieleman, Zen, Simonyan,
+  Vinyals, Graves, Kalchbrenner, Senior, and Kavukcuoglu]{van2016wavenet}
+A.~van~den Oord, S.~Dieleman, H.~Zen, K.~Simonyan, O.~Vinyals, A.~Graves,
+  N.~Kalchbrenner, A.~Senior, and K.~Kavukcuoglu.
+\newblock Wavenet: A generative model for raw audio.
+\newblock \emph{CoRR abs/1609.03499}, 2016.
+
+\end{thebibliography}
 

### 2017-10-26 paper/to_do.tex
Rewrote some of paper, added in QRNN implementation,

% Created 2017-10-23 Mon 00:20
\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{marvosym}
\usepackage{wasysym}
\usepackage{amssymb}
\usepackage{hyperref}
\tolerance=1000
\author{Christopher Cundy}
\date{\today}
\title{to\_do}
\hypersetup{
  pdfkeywords={},
  pdfsubject={},
  pdfcreator={Emacs 25.1.1 (Org mode 8.2.10)}}
\begin{document}

\maketitle
\tableofcontents

\section{To do for paper}
\label{sec-1}
\subsection{Edit all}
\label{sec-1-1}
\subsection{Look up templates etc, double-check there's not one out there}
\label{sec-1-2}
\subsection{Do and write up the experiments}
\label{sec-1-3}
\subsubsection{Synthetic Task}
\label{sec-1-3-1}
\begin{enumerate}
\item Already done, written up. Need to review
\label{sec-1-3-1-1}
\end{enumerate}
\subsubsection{Medical Task}
\label{sec-1-3-2}
\begin{enumerate}
\item Being done, need to check on testing set
\label{sec-1-3-2-1}
\item Need to write up
\label{sec-1-3-2-2}
\end{enumerate}
\subsubsection{Throughput Task}
\label{sec-1-3-3}
\begin{enumerate}
\item Chat to Eric about this
\label{sec-1-3-3-1}
\item Just generate random data
\label{sec-1-3-3-2}
\end{enumerate}
\subsubsection{PLR/SLR comparison}
\label{sec-1-3-4}
\begin{enumerate}
\item Have the pure kernel data
\label{sec-1-3-4-1}

\item Do that for random data
\label{sec-1-3-4-2}
\item Do a 2-layer network, with varying seq$_{\text{len}}$
\label{sec-1-3-4-3}
\item Report the throughput
\label{sec-1-3-4-4}

\begin{center}
\begin{tabular}{lrrrr}
Sequence Length & LS-LSTM & SRU & QRNN(2) & QRNN(10\\
\hline
16 & 0.61 & 0.28 & 0.38 & 0.78\\
256 & 0.91 & 0.84 & 0.86 & 0.99\\
4,096 & 0.98 & 1.38 & 1.18 & 1.05\\
65,536 & 1.41 & 9.21 & 6.68 & 2.05\\
\end{tabular}
\end{center}
For an input$_{\text{size}}$ of 24, output 2, hidden$_{\text{size}}$ 256, 
batch$_{\text{size}}$ 65536 / seq$_{\text{len}}$
\end{enumerate}

\subsection{Collate the experiments and smooth}
\label{sec-1-4}
\subsection{Think about the overall message of the piece and rewrite as necessary}
\label{sec-1-5}
\subsection{Send to Adam, Andy to check}
\label{sec-1-6}
\subsection{Submit}
\label{sec-1-7}
\subsubsection{Then go to party and relax the fuck out}
\label{sec-1-7-1}

\section{Timeline}
\label{sec-2}
\subsection{Saturday evening:}
\label{sec-2-1}
\subsubsection{Done interesting medical results, tested on testing set}
\label{sec-2-1-1}
\subsubsection{Looked at throughput task}
\label{sec-2-1-2}
\subsubsection{Done PLR / SLR throughput task}
\label{sec-2-1-3}
\subsection{Sunday evening:}
\label{sec-2-2}
\subsubsection{Fully written up medical task}
\label{sec-2-2-1}
\subsubsection{Written up throuput and PLR/SLR task}
\label{sec-2-2-2}
\subsubsection{Start editing the paper, smoothing into a cohesive whole}
\label{sec-2-2-3}
\subsection{Monday evening:}
\label{sec-2-3}
\subsubsection{Finish editing, send off to everyone else}
\label{sec-2-3-1}

\section{Paper structure}
\label{sec-3}
\subsection{Introduction, abstract, background}
\label{sec-3-1}
\subsubsection{Describe}
\label{sec-3-1-1}
We can apply the PLR method to any architecture that satisfies the 
constraints. E.g. SRU, QRNN, etc. Introduce LS-LSTM as good substitute.
Describe how previous papers have shown that linear LSTMs work well, this
approach allows us to speed it up.
\subsection{Experiments}
\label{sec-3-2}
\subsubsection{Benchmarks}
\label{sec-3-2-1}
\begin{enumerate}
\item Throughput of pure PLR kernel vs SLR kernel
\label{sec-3-2-1-1}
\item Show how it speeds up the SRU, QRNN, and LS-LSTM
\label{sec-3-2-1-2}
\item Show how the LS-LSTM has much better throughput than the CudnnLSTM
\label{sec-3-2-1-3}
\end{enumerate}
\subsubsection{Synthetic task to show that linear LSTMs can still work well}
\label{sec-3-2-2}
\subsubsection{Medical task}
\label{sec-3-2-3}

31.4 * (4.3 * 6) = 
\textasciitilde{}50 * \textasciitilde{}74
135 * 190
% Emacs 25.1.1 (Org mode 8.2.10)
\end{document}

### 2017-10-26 layers_new.py

@@ -1,0 +1,238 @@
+import tensorflow as tf
+from linear_recurrent_net.tensorflow_binding import linear_recurrence
+
+def vscope(name):
+    return tf.variable_scope(None, default_name=name)
+
+# contracts on the innermost dimension
+def matmul(X, W):
+    res = tf.tensordot(X, W, [[-1], [0]])
+    res.set_shape(X.get_shape().as_list()[:-1] +
+                  W.get_shape().as_list()[1:])
+    return res
+
+def embedding_layer(X, size, dims, name='embedding'):
+    with vscope(name):
+        W = tf.get_variable('W', [dims, size])
+        return tf.nn.embedding_lookup(W, X)
+
+def fc_layer(X, hidden_size, nonlin=tf.nn.elu,
+             use_bias=True, use_layer_norm=False, ln_eps=1e-3,
+             name='fc', sn=0.05, forget_bias=5.0):
+    n_dims = X.get_shape()[-1].value
+    with vscope(name):
+        W = tf.get_variable('W', initializer=tf.random_uniform([n_dims, hidden_size], maxval=sn, minval=-sn))
+
+        if use_bias and name == 'pre_fc':
+            b = tf.get_variable('b', initializer=tf.concat([tf.constant(forget_bias, shape=[hidden_size/4]),
+                                                            tf.zeros([3*(hidden_size/4)])],axis=0))
+        elif use_bias and name == 'sru_pre':
+            b = tf.get_variable('b', initializer=tf.concat([tf.zeros([(hidden_size/3)]), 
+                                                             tf.constant(forget_bias, shape=[hidden_size/3]),
+                                                             tf.zeros([(hidden_size/3)])],axis=0))            
+        elif use_bias:
+            b = tf.get_variable('b', initializer=tf.zeros([hidden_size]))
+        else:
+            b = 0
+
+        prod = matmul(X, W)
+        if use_layer_norm:
+            idx = ([None] * (len(X.shape) - 1)) + [slice(None)]
+            g = tf.get_variable('g', [hidden_size])[idx]
+
+            mu, sigma = tf.nn.moments(prod, [-1], keep_dims=True)
[…]

### 2017-10-26 plr_slr.py

@@ -1,0 +1,447 @@
+def plr_slr(bs_seq_len_list):
+    """Given a list of pairs (batch size, seq_len), 
+    calculate the throughput of an LS-LSTM, an SRU, a QRNN(2),
+    and QRNN(10) using the parallel kernel as opposed to the serial
+    one"""
+    import tensorflow as tf
+    import numpy as np
+    import scipy.io.wavfile
+    from tensorflow.contrib import rnn
+    import math
+    from layers_new import linear_surrogate_lstm
+    from layers_new import s_linear_surrogate_lstm
+    from layers_new import SRU
+    from layers_new import s_SRU
+    from layers_new import QRNN
+    from layers_new import s_QRNN        
+    import time
+    import os
+    import random
+
+    throughput_list = []
+
+    #TODO:
+    #Make LS_LSTM with PLR
+    #Make SRU with PLR
+    #Make QRNN with PLR
+    #Make LS_LSTM with SLR
+    #Make SRU with SLR
+    #Make QRNN with SLR
+    
+
+    for seq_len in seq_len_list:
+        #First generate the LS-LSTM and work out the throughput
+        tf.reset_default_graph()        
+        n_hidden = 256
+        n_classes = 2
+        n_steps = seq_len
+        batch_size = 65536 / seq_len
+        bs = batch_size
+        print "Batch size is {} and sequence length is {}".format(bs, seq_len)
+        n_input = 24
+        n_layers = 2
+        forget_gate_init = 1.0                          # = 1/(n_in). We use uniform p(x)
+        #Training Parameters
+        sn = 1.0 / math.sqrt(n_hidden)
+        learning_rate = 0.001
+        training_iters = 5000000
+
+        x = tf.placeholder("float", [n_steps, batch_size, n_input])
+        y = tf.placeholder("float", [batch_size, n_classes])
+        tf.get_variable_scope().reuse == True
+        W1 = tf.get_variable('W1', initializer=
+                             tf.random_normal([n_hidden, n_classes]), dtype='float')
+        b1 = tf.get_variable('b1', initializer=tf.zeros([n_classes]), dtype='float')
+
+        layer1 = linear_surrogate_lstm(x, n_hidden, name='ls-lstm')
+        outputs = linear_surrogate_lstm(layer1, n_hidden, name='ls-lstm2')    
+        pred = tf.matmul(outputs[-1], W1) + b1
+        #Evaluate network, run adam and clip gradients
[…]

### 2017-10-27 paper/main.tex
Paper edits etc

@@
 \usepackage{graphicx}
+\usepackage{todonotes}
+\usepackage{xargs}
+\usepackage{booktabs}
 
@@
 \algrenewtext{EndParFor}{\algorithmicendparfor}
+
+\newcommandx{\Chris}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
+\newcommandx{\Eric}[2][1=]{\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue,#1]{#2}}
+\newcommandx{\thiswillnotshow}[2][1=]{\todo[disable,#1]{#2}}
 
@@
 parallel scan algorithm, leading to rapid training on long sequences with small
-minibatch size. We develop a parallel linear recurrence CUDA kernel and show
-that it can be applied to immediately speed up several RNN architectures by up
-to 9x. We abstract prior linear sequence models into a new framework of linear
-surrogate RNNs and develop a linear surrogate for the long short-term memory 
-unit, the LS-LSTM, which is able to achieve throughput 30x greater than the
-currently fastest-available LSTM implementation. We show that the LS-LSTM is
-able to perform well in the 
+minibatch size.\Chris{Thought: we could rephrase this as allowing us to break
+the necessity of using large batches to train fast} We develop a
+parallel linear recurrence CUDA kernel and show that it can be applied to
+immediately speed up several RNN architectures by up to a factor of 9. We abstract
+recent work on linear RNNs into a new framework of linear surrogate RNNs and develop
+a linear surrogate for the long short-term memory unit, the GILR-LSTM, which is
+able to achieve throughput 30 times greater than the currently fastest-available LSTM
+implementation. We show that the GILR-LSTM is able to perform well on real-world
+problems, 
 \end{abstract}
@@
 \section{Introduction}
+Recurrent neural networks (RNNs) are widely used for sequence modelling tasks in
+domains such as natural language processing \cite{sutskever2014sequence}, speech
+recognition \cite{amodei2015deep}, and reinforcement learning
+\cite{hausknecht2015deep}. Most RNNs, including popular variants such as long
+short-term memories (LSTMs) \cite{hochreiter1997long} and gated recurrent units
+(GRUs) \cite{cho2014learning}, contain a non-linear dependency between
+sequential inputs. These non-linear dependencies create a very flexible class of
+models but limit the feasibility of training RNNs on long sequences as each
+sequence element must be processed sequentially.  Modelling sequences of
+thousands to millions of elements is important to domains such as robotics,
+remote sensing, control systems, speech recognition, medicine, and finance.
 
-Recurrent neural networks (RNNs) are widely used for sequence modelling tasks in domains such as 
-natural language processing \cite{sutskever2014sequence}, 
-speech recognition \cite{amodei2015deep},
-and reinforcement learning \cite{hausknecht2015deep}.
-Most RNNs, including popular variants such as long short-term memories (LSTMs) \cite{hochreiter1997long} and gated recurrent units (GRUs) \cite{cho2014learning}, contain a non-linear dependency
-between sequential inputs. These non-linear dependencies create a very flexible class of
-models but limit the feasibility of training RNNs on long sequences as each sequence element
-must be processed sequentially.
-Modelling sequences of thousands to millions of elements is important to domains
-such as robotics, remote sensing, control systems, speech recognition, medicine, and finance.
+The RNN serial evaluation inefficiency problem is usually mitigated by
+parallelizing the forward and backward pass over a minibatch of inputs. Without
+minibatches, RNN evaluation is a sequence of matrix-vector
+multiplications. Minibatches transform RNN computation into a sequence of more
+efficient matrix-matrix multiplications, but this speed-up brings several
+disadvantages.  RNN model size is often limited by GPU memory size, and running
+a forward and backward pass on a minibatch requires memory linear in the
+minibatch size.  Grouping data into minibatches increases the latency of each
+pass and reduces the rate of optimization steps. Finally, training with larger
+minibatches damages generalization ability \cite{keskar2017large}. It is clearly
+important to investigate if it is possible to parallelise RNN training without
+introducing minibatches. Persistent RNNs \cite{diamos2016persistent} use a novel
+implementation that can achieve high GPU utilization with very small minibatch
+sizes when the recurrent state is larger than 500 elements, but even persistent
+RNNs become limited by the serial evaluation inefficiency at smaller hidden
+sizes.
 
-The RNN serial evaluation inefficiency problem is usually mitigated by parallelizing the forward
-and backward pass over a minibatch of inputs.
-Without minibatches, RNN evaluation is a sequence of matrix-vector multiplications.
-Minibatches transform RNN computation into a sequence of more efficient matrix-matrix multiplications, but minibatches within RNNs present many issues.
-RNN model size is often limited by GPU memory size, and running a forward 
-and backward pass on a minibatch requires memory linear in
-the minibatch size.
-Grouping data into minibatches increases the latency of each pass and reduces the rate of optimization steps. Finally, training with larger minibatches damages generalization ability 
-\cite{keskar2017large}. Persistent RNNs \cite{diamos2016persistent} use a novel implementation that can achieve high GPU utilization with very small 
-minibatch sizes when the recurrent state is larger than 500 elements, but even
-persistent RNNs become limited by the serial evaluation
-inefficiency at smaller hidden sizes.
+Numerous prior works have shown strong performance from neural sequential models
+with only linear dependence on earlier sequence
+elements. \citet{balduzzi2016strongly} investigated RNNs with only elementwise
+linear recurrence relations $h_t = \alpha_t \odot h_{t-1} + (1-\alpha_t) \odot
+x_t$ and developed linear variants of LSTM and GRU that perform similarly to
+standard non-linear RNNs on text generation tasks. \citet{bradbury2017quasi},
+\citet{kalchbrenner2016neural}, \citet{gehring2017convolutional}, and
+\citet{van2016wavenet} have successfully applied networks of convolutions over
+sequences for tasks such as machine translation, language modelling, and audio
+generation.  These works have observed up to an order of magnitude increase in
+training throughput compared to RNN alternatives. Convolutional sequence models
+typically rely on either an attention mechanism or a (possibly linear) recurrent
+layer to integrate information at scales larger than the filter
+width. Introduction of a recurrent layer prevents full parallelization over the
+sequence length while attention mechanisms are expensive to apply on long
+sequences in online inference use cases.  One dimensional convolution can be
+viewed as a learnable linear finite impulse response (FIR) filter with a
+parallel evaluation algorithm, while linear recurrence is a learnable linear
+infinite impulse response (IIR).\Chris{Seems like an aside - take out?}
 
-Numerous prior works have shown strong performance from neural sequential models with only linear dependence on earlier sequence elements.
-\citet{balduzzi2016strongly} investigated RNNs with only elementwise linear recurrence relations
-$h_t = \alpha_t \odot h_{t-1} + (1-\alpha_t) \odot x_t$ and developed
-linear variants of LSTM and GRU that perform similarly to
-standard non-linear RNNs on text generation tasks.
-\citet{bradbury2017quasi},  \citet{kalchbrenner2016neural}, \citet{gehring2017convolutional}, and \citet{van2016wavenet} have successfully applied networks of convolutions over sequences for tasks
-such as machine translation, language modelling, and audio generation. 
-These works have observed up to an order of magnitude 
-increase in training throughput compared to RNN alternatives. Convolutional sequence models typically rely on either an attention mechanism or a (possibly linear) recurrent layer
-to integrate information at scales larger than the filter width. Introduction of a recurrent
-layer prevents full parallelization over the sequence length while attention mechanisms are
-expensive to apply on long sequences in online inference use cases.
-One dimensional convolution can be viewed as a learnable linear
-finite impulse response (FIR) filter with a parallel evaluation algorithm, while
-linear recurrence is a learnable linear infinite impulse response (IIR). This work
-parallelizes evaluation of linear recurrences through application of the parallel
-scan algorithm.
+  This work parallelizes evaluation of linear recurrences through application of
+the parallel scan algorithm.
 
@@
 
-Our contributions are as follows:
+Considering the familiar operations in linear algebra, we see that the
+associative operation of vector addition
+($\boldsymbol{x} \oplus \boldsymbol{y} = \boldsymbol{x} + \boldsymbol{y}$), the
+semiassociative operation of matrix-vector multiplication
+($A \otimes \boldsymbol{x} = A\boldsymbol{x}$) and the associative operation of
+matrix-matrix multiplication ($A \odot B=AB$) satisfy Blelloch's three
+conditions, allowing $h_t = \Lambda_t h_{t-1} + x_t$ to be evaluated in parallel
+over time steps \(t\) for vectors $x_t$ and square matrices $\Lambda_t$.
+
+We investigate this idea further and deliver the following contributions:
 \begin{itemize}
-\item{We describe the class of RNNs which satisfy the conditions above, and
+\item{We classify the RNNs which satisfy the conditions above, and
     show that many of the RNNs used in practice such as the Quasi-RNN are
@@
     surrogate for the LSTM and show that we are able to train it
-    with a speedup of 5-10x compared to the CuDNN-LSTM when we use the plr
-    algorithm.}
+    with a speedup of 5-10x compared to the CuDNN-LSTM when we use the
+    parallel linear recurrence algorithm.}
 \end{itemize}
@@
 \section{Parallel linear recurrence}
-Considering the familiar operations in linear algebra, we see that the
-associative operation of vector addition ($\boldsymbol{x} \oplus \boldsymbol{y}
-= \boldsymbol{x} + \boldsymbol{y}$), the semiassociative operation of
-matrix-vector multiplication ($A \otimes \boldsymbol{x} = A\boldsymbol{x}$) and
-the associative operation of matrix-matrix multiplication ($A \odot B=AB$) form
-recurrences which satisfy Blelloch's three conditions, thus allowing the linear
-recurrence $h_t = \Lambda_t h_{t-1} + x_t$ to be evaluated in parallel over $t$
-for vectors $x_t$ and square matrices $\Lambda_t$. As the method is essential to
-this work, Algorithm 1 presents the parallel linear recurrence algorithm for the
-interested reader.
+As the method is
+essential to this work, Algorithm 1 presents the parallel linear recurrence
+algorithm for the interested reader.
 \begin{algorithm}
@@
 The cost of a serial scan over a sequence of length $T$ is 
-$C_\text{sscan} = (C_\otimes + C_\oplus)T$, compared to the parallel scan cost
-$C_\text{pscan} = 2(C_\odot + C_\otimes + C_\oplus)(T/p + \lg p)$ \cite{blelloch1990prefix},
+$C_\text{sscan} \in \mathcal{O}((C_\otimes + C_\oplus)T)$, compared to the parallel scan cost
+$C_\text{pscan} \in \mathcal{O}(2(C_\odot + C_\otimes + C_\oplus)(T/p + \lg p))$ \cite{blelloch1990prefix},
 where \(p\) is the number of available parallel cores. 
 If $h_t$ is a vector of dimension $n$ then 
-$C_\odot\propto n^3, C_\otimes\propto n^2, C_\oplus\propto n$ giving
-$C_\text{pscan} \propto 2(n^3 + n^2 + n)(T/p + \lg p)$ and 
-$C_\text{sscan} \propto (n^2 + n)T$. The $n^3$ cost of the matrix
+$C_\odot \in  \mathcal{O}(n^3), C_\otimes \in \mathcal{O}(n^2), C_\oplus \in \mathcal{O}(n)$ giving
+$C_\text{pscan} \in \mathcal{O}(2(n^3 + n^2 + n)(T/p + \lg p))$ and 
+$C_\text{sscan} \in \mathcal{O}((n^2 + n)T)$. The $\mathcal{O}(n^3)$ cost of the matrix
 multiplication in the parallel algorithm can counter-act any parallel speedups for
@@
 which case both matrix-matrix and matrix-vector multiplication have cost
-proportional to $n$ and $C_\text{pscan}\propto 6n(T/p + \lg p)$ and
-$C_\text{sscan}=2nT$. Assuming $p \ll T$, then $C_\text{pscan} \le
-C_\text{sscan}$ when $p \ge 3$.\footnote{Can include the actual speedup
+proportional to $n$ and $C_\text{pscan}\in \mathcal{O}(6n(T/p + \lg p))$ and
+$C_\text{sscan} \in \mathcal{O}(2nT)$. Assuming $p \ll T$, then $C_\text{pscan} \le
+C_\text{sscan}$ when $p \ge 3$.\Chris{Can include the actual speedup
   factor, \(\alpha = \frac{PT}{3(T + \lg(p))}\).}
@@
 network models use separate recurrent matrices for each sequence element and using these
-separate matrices would require potentially prohibitive $n^2T$ memory.\footnote{Ideally we can
-  find a citation for someone thinking these points before}
+separate matrices would require potentially prohibitive $n^2T$ memory.\Chris{I can't
+seem to find anyone explicitly saying this before from a quick google.}
 Applying
@@
 \end{align*}
-A GILR layer applies the same non-linear transform to each sequence element and then accumulates the sequence elements with a non-linear gating mechanism.
-Gate $g_t$ uses the sigmoid activation function to have values in [0,1] for reasonable gating semantics, while impulse $i_t$ can use any activation function $\tau$.
-Stacking GILR layers allows for rich non-linear dependence on previous events while still taking advantage of fast parallel sequence evaluation.
+A GILR layer applies the same non-linear transform to each sequence element and
+then accumulates the sequence elements with a non-linear gating mechanism. Gate
+$g_t$ uses the sigmoid activation function to give values in [0,1] for
+reasonable gating semantics, while impulse $i_t$ can use any activation function
+$\tau$. Stacking GILR layers allows for rich non-linear dependence on previous
+events while still taking advantage of fast parallel sequence evaluation.
 
@@
 Now consider the GILR, noting that it has the same two matrix-vector
-multiplications per iteration as the above RNN. $g$ and $i$ can each be
-evaluated for all $t$ with a single $(bT, m) \times (m, n)$ matrix
-multiplication each. Given $g$ and $i$, $h$ can be computed using a parallel
-linear recurrence over $T$ vectors each of $bn$ elements. Rather than $T$ small
-operations, the GILR can be evaluated over all sequence elements with two large
-matrix multiplies and a parallel linear recurrence. GILR performance is much
-less dependent on batch size as the matrix multiplies see an "effective batch
-size" of $bT$ and $T$ is typically large.
+multiplications per iteration as the above RNN. The intermediate variables $g$
+and $i$ can be evaluated for all $t$ with a single $(bT, m) \times (m, n)$
+matrix multiplication each. Given $g$ and $i$, $h$ can be computed using a
+parallel linear recurrence over $T$ vectors each of $bn$ elements. Rather than
+$T$ small operations, the GILR can be evaluated over all sequence elements with
+two large matrix multiplications and a parallel linear recurrence. GILR performance
+is much less dependent on batch size as the matrix multiplies see an "effective
+batch size" of $bT$ and $T$ is typically large.
 
@@
 summarizing the previous inputs while $s_t$ serves as the output of $f$ to be
-passed to other layers of the network. If we decouple these uses, we can instead
-compute $s_t = f(\tilde{s}_{t-1}, x_t)$ with $\tilde{s}_t$ as a linearly
-computable surrogate for $s_t$. With this linear surrogate, non-linear $f$ can
-still be evaluated. We refer to this class of model as a linear surrogate RNN
+passed to other layers of the network. We can decouple these uses and introduce
+independent variables for each purpose: \(s_t\) is passed onto other layers of the network
+and we introduce the linear surrogate \(\tilde{s}_t\) which is passed onto the
+next state, with \(s_t = f(\tilde{s}_{t-1}, x_t)\). We are still able to choose a 
+nonlinear \(f\), our only limitation being that \(\tilde{s}_t\) must be linearly
+computable.  We refer to this class of model as a linear surrogate RNN
 (LS-RNN). Quasi-RNNs \cite{bradbury2017quasi} are LS-RNNs using $\tilde{h}_{t-1}
@@
 
-As an example LS-RNN, consider an LSTM\footnote{Want a citation here I guess}:
-\begin{align*}
-f_t, i_t, o_t &= \sigma(U_{f,i,o} h_{t-1} + V_{f,i,o} x_t + b_{f,i,o}) \\
-z_t &= \tau(U_z h_{t-1} + V_z x_t + b_z) \\
-c_t &= f_t \odot c_{t-1} + i_t \odot z_t \\
-h_t &= o_t \odot c_t
-\end{align*}
-An LSTM has state $s_t = (h_t, c_t)$.
-Since $c_t$ depends only linearly on $c_{t-1}$, no surrogate is needed for $c_t$.
-$h_t$ has a non-linear dependence on $h_{t-1}$, so $h_t$ needs a linear surrogate.
-With a GILR layer as surrogate, the linear surrogate LSTM (LS-LSTM) is
-\begin{align*}
-f_t, i_t, o_t &= \sigma(U_{f,i,o} \tilde{h}_{t-1} + V_{f,i,o} x_t + b_{f,i,o}) \\
-z_t &= \tau(U_z \tilde{h}_{t-1} + V_z x_t + b_z) \\
-c_t &= f_t \odot c_{t-1} + i_t \odot z_t \\
-h_t &= o_t \odot c_t \\
-g_t &= \sigma(V_g x_t + b_g) \\
-\tilde{h}_t &= g_t \odot \tilde{h}_{t-1} + (1-g_t)\odot \tau(Wx_t + b_h)
+As an example LS-RNN, consider an LSTM:
+\begin{align*} f_t, i_t, o_t &= \sigma(U_{f,i,o} h_{t-1} + V_{f,i,o} x_t +
+b_{f,i,o}) \\ z_t &= \tau(U_z h_{t-1} + V_z x_t + b_z) \\ c_t &= f_t \odot
+c_{t-1} + i_t \odot z_t \\ h_t &= o_t \odot c_t
+\end{align*} An LSTM has state $s_t = (h_t, c_t)$. Since $c_t$ depends only
+linearly on $c_{t-1}$, no surrogate is needed for $c_t$. $h_t$ has a non-linear
+dependence on $h_{t-1}$, so $h_t$ needs a linear surrogate. Introducing a GILR layer as
+the surrogate, we obtain the GILR-LSTM:
+\begin{align*} f_t, i_t, o_t &= \sigma(U_{f,i,o} \tilde{h}_{t-1} + V_{f,i,o} x_t
++ b_{f,i,o}) \\ z_t &= \tau(U_z \tilde{h}_{t-1} + V_z x_t + b_z) \\ c_t &= f_t
+\odot c_{t-1} + i_t \odot z_t \\ h_t &= o_t \odot c_t \\ g_t &= \sigma(V_g x_t +
+b_g) \\ \tilde{h}_t &= g_t \odot \tilde{h}_{t-1} + (1-g_t)\odot \tau(Wx_t + b_h)
 \end{align*}
 
-For $m$ inputs and hidden size $n$, the LS-LSTM contains $2n(n+m)$ more parameters than the equivalently sized LSTM to handle the mapping from $x$ to $\tilde{h}$.
-More generally, a LS-RNN contains all of the same parameters as the underlying RNN as well as some additional parameters to compute the linear surrogate.
+For $m$ inputs and hidden size $n$, the GILR-LSTM contains $2n(n+m)$ more
+parameters than the equivalently sized LSTM to handle the mapping from $x$ to
+$\tilde{h}$. More generally, a LS-RNN contains all of the same parameters as the
+underlying RNN as well as some additional parameters to compute the linear
+surrogate.
 
@@
 parallel linear recurrence algorithm is able to achieve higher throughput than a
-similarly implemented serial version of the algorithm. We then show that the
-LS-LSTM architecture, trained with the PLR algorithm, is able to outperform the
-optimized CuDNN LSTM implementation on a pathological example from the original
-LSTM paper. Finally we show that our LS-LSTM is able to quickly achieve state of
-the art results on an existing medical data set.
+similarly implemented serial version of the algorithm. We show that the plr kernel
+is up to 40 times faster than the serial equivalent for long sequence lengths, and
+that this speedup translates to implementations of LS-RNNs such as the QRNN (where
+we can get a speedup of up to 9 times).
+
+In order to illustrate that the linearisation does not necessarily come at the
+cost of expressibility, we show that the GILR-LSTM architecture, trained with
+the PLR algorithm, is able to outperform the optimized CuDNN LSTM implementation
+on a pathological example from the original LSTM paper
+\cite{hochreiter1997long}. Finally we show that our GILR-LSTM is able to
+quickly achieve state of the art results on an existing medical data
+set. \Chris{I hope so\ldots}
 
 \subsection{Throughput Benchmarks}
-**** Throughput of pure PLR kernel vs SLR kernel
-**** Show how it speeds up the SRU, QRNN, and LS-LSTM
-**** Show how the LS-LSTM has much better throughput than the CudnnLSTM
+We first illustrate the throughput advantage of the parallel scan algorithm for
+evaluating the linear recurrence. For a minibatch comprised of \(b\) sequences
+of length \(T\), we define the number of events as \(bT\) and the throughput as
+the number of events per second. We implement two CUDA kernels, one which
+evaluates the parallel linear recurrence described in algorithm \ref{alg:plr},
+and one which evaluates the linear recurrence using the straightforward serial
+approach. This comparison is performed directly at the kernel level, avoiding
+any overhead due to calling Tensorflow. We find that at long sequence lengths,
+the PLR kernel has a distinct advantage, as shown in table
+\ref{table:kernel-throughput}, with a speedup factor of 40. At short sequence
+lengths, the parallel kernel does not perform well due to the extra overhead
+involved in marshalling the parallel algorithm. \Chris{Important:
+  What's the batch size here? Is it just one sequence? Need to say that}
+\begin{table}[]
+  \centering
+\caption{Speedup of the PLR kernel compared to the SLR kernel, without any
+  overhead from Tensorflow, for various sequence lengths and input size \(m\).}      
+\label{table:kernel-throughput}
+\begin{tabular}{@{}llll@{}}
+Sequence Length & \(m=4\)  & \(m=32\) & \(m=128\) \\ \midrule
+16              & 0.06 & 0.06 & 0.05  \\
+256             & 0.22 & 0.22 & 0.86  \\
+4,096           & 1.02 & 2.94 & 3.36  \\
+65,536          & 38.5 & 41.8 & 17.5  \\ \bottomrule
+\end{tabular}
+\end{table}
 
-We first illustrate the throughput advantage of the parallel scan algorithm for
-evaluating the linear recurrence. We define throughput as events/s and a
-minibatch of size $b$ and length $T$ to be $bT$ events. We implement two CUDA
-kernels, one which evaluates the parallel linear recurrence described in
-algorithm \ref{alg:plr}, and one which evaluates the linear recurrence using the
-straightforward serial approach. This comparison is performed directly at the
-kernel level, avoiding any overhead due to calling Tensorflow. We find that at
-long sequence lengths, the PLR kernel has a distinct
-advantage, as shown in table \ref{table:kernel-throughput}.
+Many of the RNNs that have been recently introduced can be implemented with a
+parallel linear recurrence. Table \ref{table:rnn-throughput} shows the
+throughput advantage by using a parallel linear recurrence when compared to a
+serial linear recurrence. Since PLR and SLR are simply different methods for
+computing the recurrence, there is no penalty incurred for this speedup - we
+obtain exactly the same output.
 
+Typically experiments will aim to maximize the memory
+utilisation of the GPU, so we simulate this by choosing the product of the batch
+size and the sequence length to be constant. We see that the architectures which
+have fewer terms (for which the linear recurrence is a higher proportion of the
+computational burden) are more effected by the switch to PLR.  This is
+particularly clear in the case of the QRNN, where including a wider
+convolutional window results in less time spent on the recurrence and less to
+gain from the parallelization.
 \begin{figure}
 \begin{center}
-  \begin{tabular}{lrrr}
+  \begin{tabular}{@{}lrrrr@{}}
     \label{table:rnn-throughput}
-Sequence Length\footnote{What's the batch size etc here??} & 4 Dimensions & 32 Dimensions & 128 Dimensions \\
-\hline
-16 & 0.06 & 0.06 & 0.05\\ 
-256 & 0.22 & 0.22 & 0.86\\ 
-4,096 & 1.02 & 2.94 & 3.36\\ 
-65,536 & 38.5 & 41.8 & 17.5\\
-  \end{tabular}  
-    \end{center}
-    \caption{Speedup of the PLR kernel compared to the SLR kernel, without any
-    overhead from Tensorflow.}    
-  \end{figure}
-
-Many RNN architectures that have been recently introduced are amenable to
-implementing through parallel linear recurrence. Table
-\ref{table:rnn-throughput} shows the throughput advantage by using a parallel
-linear recurrence when compared to a serial linear recurrence. Since PLR and SLR
-are simply different methods for computing the same output, there is no penalty
-incurred for this speedup: it is a pure speed gain. Typically experiments will aim
-to maximize the memory utilisation of the GPU, so we simulate this by choosing the
-product of the batch size and the sequence length to be constant. We see that the
-architectures which have fewer terms (for which the linear recurrence is a higher
-proportion of the computational burden) are more effected by the switch to PLR.
-This is particularly clear in the case of the QRNN, where including a wider
-convolutional window results in less time spent on the recurrence and less to gain
-from the parallelization. 
-\begin{figure}
-\begin{center}
-  \begin{tabular}{lrrrr}
-    \label{table:rnn-throughput}
-Sequence Length & LS-LSTM & SRU & QRNN(2) & QRNN(10)\\
-\hline
+Sequence Length & GILR-LSTM & SRU & QRNN(2) & QRNN(10)\\ \midrule
 16 & 0.61 & 0.28 & 0.38 & 0.78\\
@@
 4,096 & 0.98 & 1.38 & 1.18 & 1.05\\
-65,536 & 1.41 & 9.21 & 6.68 & 2.05\\
+65,536 & 1.41 & 9.21 & 6.68 & 2.05\\ \bottomrule
   \end{tabular}  
 \end{center}
-\caption{Speedup obtained when using parallel linear recurrence compared
-    to serial linear recurrence. We analyse this for several recent architectures:
-    the Simple Recurrent Unit from [cite], the Quasi-RNN from [cite], with a
-    convolutional filter of size 2 and 10, and the Linear Surrogate LSTM that
-    we introduce in section \ref{sec:ls-rnns}. To give a fair model of a
-    typical use-case, we set batch size \(b\) such that \(Tb = 65536\)
-    for a sequence length \(T\).}
+\caption{Speedup obtained when using parallel linear recurrence compared to
+  serial linear recurrence. We analyse this for several recent architectures:
+  the Simple Recurrent Unit from \cite{lei2017}, the Quasi-RNN from
+  \cite{bradbury2017quasi}, with a convolutional filter of size 2 and 10, and
+  the GILR-LSTM which we introduce in section \ref{sec:ls-rnns}. To give a fair
+  model of a typical use-case maximising the memory usage of the GPU, we set
+  batch size \(b\) such that \(Tb = 65536\) for a sequence length \(T\).}
 \end{figure}
@@
 Finally, we compare the throughput of the linear surrogate LSTM to the current
-fastest available implementation of the LSTM, the CuDNN-LSTM. We observe that our
-LS-LSTM has higher throughput for most of the low-batch-size regime, with a 30x
-speedup at a sequence length of 65,536. Recent work (cite https://arxiv.org/pdf/1609.04836.pdf)
-has suggested that small mini-batches may aid generalization by encouraging
-the discovery of shallow minima. The LS-LSTM allows us to achieve high throughput
-for all choices of minibatch size and sequence length.
+fastest available implementation of the LSTM, the CuDNN-LSTM. Figure
+\ref{fig:tp_perf} shows that our GILR-LSTM has higher throughput for most of the
+low-batch-size regime, with a 30x speedup at a sequence length of 65,536. Recent
+work \cite{keskar2017large} has suggested that small mini-batches may aid
+generalization by encouraging the discovery of shallow minima. The GILR-LSTM
+allows us to achieve high throughput for all choices of minibatch size and
+sequence length.
 
-\ref{fig:tp_perf}.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% The computational performance of the LSTM and LS-LSTM models were compared       %
+% The computational performance of the LSTM and GILR-LSTM models were compared       %
 % across a wide range of minibatch sizes and sequence lengths. Although the 234    %
-% unit LS-LSTM has a similar number of parameters to the 256 unit LSTM, we         %
-% measured the computational performance of a 256 unit LS-LSTM to observe any      %
+% unit GILR-LSTM has a similar number of parameters to the 256 unit LSTM, we         %
+% measured the computational performance of a 256 unit GILR-LSTM to observe any      %
 % impact of the linear surrogate calculation. The serial evaluation of the LSTM    %
 % causes its throughput to be independent of the sequence length as a doubling of  %
-% sequence length causes a doubling of runtime. Notably, the LS-LSTM can achieve a %
+% sequence length causes a doubling of runtime. Notably, the GILR-LSTM can achieve a %
 % higher throughput by running on one 8192 event sequence at a time than an LSTM   %
@@
 
-Unlike the previous case, these speedups do come at a cost. The LS-LSTM
+Unlike the previous case, these speedups do come at a cost. The GILR-LSTM
 architecture is different to the canonical LSTM, and it's possible that the loss
 of model capacity outweighs the speedup in training. The following experiment
-shows that the LS-LSTM is able to outperform the LSTM on an example originally
+shows that the GILR-LSTM is able to outperform the LSTM on an example originally
 designed to illustrate the advantages of the LSTM.
@@
 \includegraphics[width=12cm]{cudnn_heatmap.png}
-\caption{Throughput comparison between LSTM-256-256 with LS-LSTM-256-256, with a
+\caption{Throughput comparison between LSTM-256-256 with GILR-LSTM-256-256, with a
   32-dimensional input and a 2-dimensional output, for various batch sizes and
   sequence lengths.The LSTM only has a single row of data because its throughput
-  is independent of sequence length. Entries are missing from the LS-LSTM table
+  is independent of sequence length. Entries are missing from the GILR-LSTM table
   because there was not enough memory on the GPU to handle such large batch
@@
 
-% We attempt to forecast 50 events ahead in N-MNIST using a two layer LSTM with 256 units per layer and a two layer LS-LSTM with 234 units per layer.
-The LS-LSTM layer size was selected so that it had slightly fewer parameters than the LSTM.
+% We attempt to forecast 50 events ahead in N-MNIST using a two layer LSTM with 256 units per layer and a two layer GILR-LSTM with 234 units per layer.
+The GILR-LSTM layer size was selected so that it had slightly fewer parameters than the LSTM.
 Each model transformed the incoming event position into a 40-dimensional embedding vector which was then combined with the polarity to produce a 41 dimensional input.
@@
 
-% The LSTM and LS-LSTM offer differ latency and throughput tradeoffs.
+% The LSTM and GILR-LSTM offer differ latency and throughput tradeoffs.
 LSTM throughput depends only on minibatch size but LSTM latency depends on both minibatch size and sequence length.
@@
 
-% The throughput and latency of the LS-LSTM are both influenced by batch size and sequence length.
-Several LS-LSTMs were trained with batch size 16 and sequence length 1024.
+% The throughput and latency of the GILR-LSTM are both influenced by batch size and sequence length.
+Several GILR-LSTMs were trained with batch size 16 and sequence length 1024.
 This combination was chosen because of the nearly maximum throughput, the low latency, and the ease of building a minibatch given the distribution of sequence lengths in the data set.
-The 234 unit LS-LSTM reaches a better training loss in roughly 4.5 hours than any of the LSTMs could reach in 18 hours.
-This experimental evidence indicates the LS-LSTM is as powerful of a model as the LSTM and can be trained in a fraction of the time.
+The 234 unit GILR-LSTM reaches a better training loss in roughly 4.5 hours than any of the LSTMs could reach in 18 hours.
+This experimental evidence indicates the GILR-LSTM is as powerful of a model as the LSTM and can be trained in a fraction of the time.
 
@@
 % latency of optimization steps and the higher final loss is explained by the
-% inability to learn long dependencies. The LS-LSTM has 234 units per hidden
+% inability to learn long dependencies. The GILR-LSTM has 234 units per hidden
 % layer to have the same number of parameters as the 256 unit LSTMs.}
@@
 %     \midrule
-%     LS-LSTM 			&  \textbf{4.567} & 4.856           & 10.51\% & 36.70\%   \\
+%     GILR-LSTM 			&  \textbf{4.567} & 4.856           & 10.51\% & 36.70\%   \\
 %     LSTM, seqlen=128    & 4.664 & \textbf{4.846}  & \textbf{10.62\%} & \textbf{37.08\%}   \\
@@
 Our focus was the fast training of powerful models, and we leave regularizing parallel linear recurrences and LS-RNNs to future work.
-Although not tested, it is possible that the LS-LSTM generalized better than the LSTMs trained on long sequences due to the much smaller minibatch size of LS-LSTM leading to a wider minima.
+Although not tested, it is possible that the GILR-LSTM generalized better than the LSTMs trained on long sequences due to the much smaller minibatch size of GILR-LSTM leading to a wider minima.
 
@@
 In order to demonstrate how PLR may be used to speed up tasks that the LSTM is
-well-suited for, we tackle the pathological example 2 from cite Hochreiter and
-Schmidhtbuer which involves storing a number for a very long sequence of time
-steps. We have an alphabet of size \(p\) with each character represented as a
-particular one-hot vector in this space. We input a sequence of \(n\) of these
-vectors, chosen at random (with replacement). The first vector in the sequence
-is always the same, up to the sign of the component (i.e. it is \(\pm p_0
-\).
+well-suited for, we tackle a pathological inference problem. Originally
+introduced as example 2b from Hochreiter and Schmidhtbuer's introduction of the
+LSTM, \cite{hochreiter1997long}, it involves storing a number for a very long
+sequence of time steps. We have an alphabet of size \(p\) with each character
+represented as a particular one-hot vector in this space. We input a sequence of
+\(n\) of these vectors, chosen at random (with replacement). The first vector in
+the sequence is always the same, up to the sign of the component (i.e. it is \(\pm
+\boldsymbol{p_0} \).
 
-The two signs on the first component separate the sequences into two
-sets. The whole sequence is fed into the LSTM and we aim to learn to classify
-the sequences. This requires remembering the first element over the length of
-the sequence. Due to the large throughput of the LS-LSTM, we would expect that
-it would do better when the sequence length is large. In the original
-formulation of the problem, \(p\) is set equal to \(n\). Since this would make
-the size of the input data grow impracticaly large as \(\mathcal{O}(n^2)\) for
-long sequences, we fix \(p = 128\) and vary \(n\).
+The two signs on the first component separate the sequences into two sets. The
+whole sequence is fed into the LSTM and we aim to learn to classify the
+sequences. This requires remembering the first element over the length of the
+sequence. Due to the large throughput of the GILR-LSTM, we would expect that it
+would do better when the sequence length is large. In the original formulation
+of the problem (dealing in the regime with around one hundred timesteps)
+, \(p\) is set equal to \(n\). Since this would make the size of the input data
+grow impracticaly large as \(\mathcal{O}(n^2)\) for long sequences, we fix \(p =
+128\) and vary \(n\).
 
 We generated three sets of data: for \(n\) equal to 1,024, 8,192, and
-1,048,576. For each of these we compared a two-layer LS-LSTM with 512 hidden
+1,048,576. For each of these we compared a two-layer GILR-LSTM with 512 hidden
 units to a canonical LSTM network with roughly the same number of parameters. We
@@
 Initial experiments showed that the two-layer LSTM was much quicker to converge
-than the one-layer LSTM, so here we compare the LS-LSTM to the two-layer
+than the one-layer LSTM, so here we compare the GILR-LSTM to the two-layer
 CudnnLSTM.  We ran all experiments on a NVIDIA K80 GPU, choosing the largest
@@
 minibatches giving 100\% accuracy. As can be seen from the learning curves
-in figure \ref{fig:synthetic_training}, this was a reasonable criterion.
+in figure \ref{fig:synthetic_training}, this was a reasonable criterion. For the
+longest sequence-length, we didn't observe the CuDNN-LSTM converging.
 \begin{figure}
@@
   \end{minipage}
-  \caption{Training curves for LS-LSTM and CUDNNLSTM architectures for various
-  sequence lengths.}
+  \caption{Training curves for GILR-LSTM and CUDNNLSTM architectures for various
+  sequence lengths. (Clockwise from top-left: 1024, 8192, 1048576)}
     \label{fig:synthetic_training}
@@
 \centering
-\caption{Performance of the LS-LSTM compared to CUDA-optimised CudnnLSTM
-  implementation on problem 2b from (cite '97). We see that the LS-LSTM has a clear
+\caption{Performance of the GILR-LSTM compared to CUDA-optimised CudnnLSTM
+  implementation on problem 2b from (cite '97). We see that the GILR-LSTM has a clear
   advantage in training speed as the sequence length increases.
+  We did not observe the Cudnn-LSTM converging for the longest task, even after several days.
   {\scriptsize *For the longest sequence length, the number of hidden units was decreased to
@@
 \multicolumn{2}{c}{\textbf{1,024}} & \multicolumn{2}{c}{\textbf{8,192}} &
-\multicolumn{2}{c}{\textbf{1,048,576\(^*\)}} \\ \midrule & CudnnLSTM & LS-LSTM &
-CudnnLSTM & LS-LSTM & CudnnLSTM & LS-LSTM \\ \cmidrule(l){2-7} Iterations (1000s) &
+\multicolumn{2}{c}{\textbf{1,048,576\(^*\)}} \\ \midrule &CuDNN& G &
+C &GILR&CuDNN&GILR\\ \cmidrule(l){2-7} Iterations (1000s) &
                      1.0 %bs=8
@@
                                                                  \(\pm\) 0.16
-                                                               & & 14 \(\pm\) 3 \\
+                                                               &-& 14 \(\pm\) 3 \\
 \begin{tabular}[c]{@{}l@{}}Wall Clock time (hours)\end{tabular} &
-0.278 \(\pm\) 0.083 & 0.0314 \(\pm\) 0.0019 & 0.58 \(\pm\) 0.06 & 0.10 \(\pm\) 0.03
-& & 9.7 \(\pm\) 1.7 \\ \bottomrule
+0.28 \(\pm\) 0.08 & 0.031 \(\pm\) 0.002 & 0.58 \(\pm\) 0.06 & 0.10 \(\pm\) 0.03
+&-& 9.7 \(\pm\) 1.7 \\ \bottomrule
 \end{tabular}
@@
 \subsection{Medical Example}
-A common application of RNNs is in medicine. In order to illustrate the increased
-performance that the plr algorithm allows, we implement an LS-LSTM on the problem
-presented in \texttt{CITE}, the 2016 PhysioNet/Computing in Cardiology Challenge 2016.
-The aim of the Classification of Normal/Abnormal Heart Sound Recordings challenge
-is to determine which of a set of EEGs are produced by people with abnormal hearts.
+A common application of RNNs is in medicine. In order to illustrate the
+increased performance that the plr algorithm allows, we implement an GILR-LSTM
+on the problem presented in \Chris{CITE}, the 2016 PhysioNet/Computing in
+Cardiology Challenge 2016.  The aim of the Classification of Normal/Abnormal
+Heart Sound Recordings challenge is to determine which of a set of EEGs are
+produced by people with abnormal hearts.  The data is recorded from a variety of
+sources and split into a training and testing set. The testing set is not
+publicly available, but can be evaluated through a web portal. The top results
+cover a large range of techniques, but generally rely on featurization of the
+raw EEG records. We want to see if we are able to leverage our advantage at high
+sequence-lenghts to train directly on the raw EEG. The EEGs are sampled at 2,000
+Hz and range in length from several seconds to more than one hundred seconds,
+giving a max sequence length of over 200,000. Since the testing server is not
+equipped with a GPU, we cannot run PLR in the testing step. We resample the
+records to 250 Hz so that we are able to evaluate the test results in a
+reasonable amount of time, padding all records that are shorter than the longest.
 
-The training set of data is unbalanced, with 75\% of the data corresponding to abnormal
-hearts.
+The training set of data is unbalanced, with 75\% of the data corresponding to
+normal hearts. We rebalance the data by oversampling the abnormal data by a factor
+of three, including each abnormal heart three times.
 
@@
 \begin{itemize}
-\item{With the fast training of the LS-LSTM / PLR, we can train quickly on very
+\item{With the fast training of the GILR-LSTM / PLR, we can train quickly on very
     long sequence-lengths}
@@
 \section{Conclusion}
-Parallel linear recurrence is an extremely powerful algorithm and the LS-LSTM is just one
+Parallel linear recurrence is an extremely powerful algorithm and the GILR-LSTM is just one
 of many possible models that can be built with it.
@@
 Besides future research, existing models such as Quasi-RNNs and strongly typed RNNs that already contain linear recurrences can immediately benefit from parallel linear recurrence.
-This work demonstrates the LS-LSTM significantly accelerates the training of small to medium sized LSTMs.
+This work demonstrates the GILR-LSTM significantly accelerates the training of small to medium sized LSTMs.
 Although similar techniques have been used before, the now explicit concept of linear surrogacy provides a framework for future development and analysis of fast sequence models.
@@
 
+\bibitem[Lei and Zhang (2017)]{lei2017}
+T.~Lei, Y.~Zhang,
+\newblock Training RNNs as fast as CNNs.
+\newblock \emph{arXiv preprint arXiv:1709.02755 }, 2017.
+
 \end{thebibliography}

### 2017-10-27 paper/main.tex
Added more to synthetic example section

@@
 
-\begin{figure}[t]
+\begin{table}[t]
 \centering
@@
 \label{fig:tp_perf}
-\end{figure}
+\end{table}
 
@@
 longest sequence-length, we didn't observe the CuDNN-LSTM converging.
+
+The results illustrate that the GILR-LSTM is able to robustly outperform the
+CuDNN-LSTM, which is currently the fastest LSTM implementation available. This
+is somewhat surprising, as we might expect that the LSTM would do particularly
+well at an example that was specifically constructed for it. We can conclude that
+the nonlinearities in the LSTM's design are not vital for its performance given
+that the linear GILR-LSTM is able to converge, and that if these nonlinearities are
+not completely necessary, they entail a large performance cost by diallowing the
+plr algorithm from being used.
+
 \begin{figure}
@@
 
-\subsection{Medical Example}
+\subsection{The Medical Setting}
 A common application of RNNs is in medicine. In order to illustrate the
@@
 
-NEED TO GET THOSE RESULTS!
+NEED TO GET THOSE RESULTS!\Chris{Quite likely that the results won't be available
+  for tomorrow, or that they won't be particularly impressive. We may have to
+  scratch this entire section unfortunately}
 
 \section{Conclusion}
-Parallel linear recurrence is an extremely powerful algorithm and the GILR-LSTM is just one
-of many possible models that can be built with it.
-Future applications of parallel linear recurrence could include sequences orders of magnitude longer than N-MNIST, the development of parallel computable differentiable memory modules, and the combination of linear recurrence with convolutional sequence models.
-Besides future research, existing models such as Quasi-RNNs and strongly typed RNNs that already contain linear recurrences can immediately benefit from parallel linear recurrence.
-This work demonstrates the GILR-LSTM significantly accelerates the training of small to medium sized LSTMs.
-Although similar techniques have been used before, the now explicit concept of linear surrogacy provides a framework for future development and analysis of fast sequence models.
+Parallel linear recurrence is an extremely powerful algorithm and the GILR-LSTM
+is just one of many possible models that can be built with it. While we were
+previously forced to use large minibatches in order to obtain reasonable
+training speed, with PLR in many cases we can train rapidly on small minibatch
+sizes, freeing us to utilise extremely long sequence lengths and more complex
+models. Our CUDA kernel is up to 40 times faster than a serial kernel on
+sequence lengths of 65,536, and this benefit transfers to implemented models,
+speeding up the SRU \cite{lei2017} by a factor of 9. The impressive results in
+\cite{lei2017} and other recent work comprehensively show that
+linearly-recurrent RNNs are, in many respects, equally as powerful as nonlinear
+LSTMs. Our results on the synthetic example continue this trend, showing that
+the GILR-LSTM is able to handle sequence length regimes that are longer than any
+we have seen seriously suggested in the literature. We feel that the framework
+of describing linear RNNs as linear surrogates for nonlinear RNNs is a fruitful
+approach, and that this, paired with the PLR algorithm, will substantially
+advance the ability to train RNNs, and to understand timeseries data.
 
-We intend to expand upon the experiments section and open-source the parallel linear recurrence kernel in the near future.
+%Future work can the development of parallel computable
+%differentiable memory modules, and the combination of linear recurrence with
+%convolutional sequence models.
+
+% Besides future research, existing models such as
+% Quasi-RNNs and strongly typed RNNs that already contain linear recurrences can
+% immediately benefit from parallel linear recurrence. This work demonstrates the
+% GILR-LSTM significantly accelerates the training of small to medium sized
+% LSTMs. Although similar techniques have been used before, the now explicit
+% concept of linear surrogacy provides a framework for future development and
+% analysis of fast sequence models.
+
+We intend to expand upon the experiments section and open-source the parallel
+linear recurrence kernel in the near future.
 

### 2017-10-27 paper/main.tex
Added more synthetic example explanation

@@
 not completely necessary, they entail a large performance cost by diallowing the
-plr algorithm from being used.
+plr algorithm from being used. Furthermore, the sequence length of over one million
+is the largest sequence length we have seen seriously discussed. The GILR
+handles it straightforwardly.
 
@@
 \multicolumn{2}{c}{\textbf{1,024}} & \multicolumn{2}{c}{\textbf{8,192}} &
-\multicolumn{2}{c}{\textbf{1,048,576\(^*\)}} \\ \midrule &CuDNN& G &
-C &GILR&CuDNN&GILR\\ \cmidrule(l){2-7} Iterations (1000s) &
+\multicolumn{2}{c}{\textbf{1,048,576\(^*\)}} \\ \midrule &CuDNN& GILR &
+CuDNN &GILR&CuDNN&GILR\\ \cmidrule(l){2-7} Iterations (1000s) &
                      1.0 %bs=8

### 2017-10-27 paper/main.tex
Medical section good to be included, we just need the final test score

@@
 
-% Experiments were performed on the N-MNIST \cite{orchard2015converting} dataset. N-MNIST captures the MNIST digit dataset \cite{lecun1998mnist} by panning
-% an event driven camera over each digit. Each example in N-MNIST is a sequence of single pixel events (x, y, polarity, timestamp) where
-% x and y each indicate a pixel position in [0, 33], polarity indicates whether the pixel was switching on or off, and timestamp is the time of the event in microseconds.
-Videos produced from the event data show that positive polarity events are often located on the leading edge of the digit motion and negative polarity events on the trailing edge.
-Between the 60,000 digits in the training set, N-MNIST contains approximately 250 million pixel events with sequence lengths ranging from 500 to 8000 and averaging 4000 events.
+% % Experiments were performed on the N-MNIST \cite{orchard2015converting}
+% dataset. N-MNIST captures the MNIST digit dataset \cite{lecun1998mnist} by
+% panning % an event driven camera over each digit. Each example in N-MNIST is a
+% sequence of single pixel events (x, y, polarity, timestamp) where % x and y each
+% indicate a pixel position in [0, 33], polarity indicates whether the pixel was
+% switching on or off, and timestamp is the time of the event in
+% microseconds. Videos produced from the event data show that positive polarity
+% events are often located on the leading edge of the digit motion and negative
+% polarity events on the trailing edge. Between the 60,000 digits in the training
+% set, N-MNIST contains approximately 250 million pixel events with sequence
+% lengths ranging from 500 to 8000 and averaging 4000 events.
 
-% We attempt to forecast 50 events ahead in N-MNIST using a two layer LSTM with 256 units per layer and a two layer GILR-LSTM with 234 units per layer.
-The GILR-LSTM layer size was selected so that it had slightly fewer parameters than the LSTM.
-Each model transformed the incoming event position into a 40-dimensional embedding vector which was then combined with the polarity to produce a 41 dimensional input.
-Both models output a $2 \mathbf{x} 34^2$ matrix containing two future event location probability distributions conditioned on the polarity of the future event.
-The training algorithm only considers the probability distribution of the true future polarity and uses the cross entropy loss function.
-The Adam \cite{kingma2014adam} optimization algorithm and Glorot \cite{glorot2010understanding} initialization scheme were used.
-Training on a minibatch of size $b$ with sequence length $T$ consisted of uniformly sampling $b$ N-MNIST sequences and then extracting a single random $T$ element subsequence (and its 50 element ahead forecast) from the sequence.
-Sequences less than $T$ elements were padded out to length $T$, and there was a "burn-in" of 30 events at the start of each sequence where no predictions were made.
-An epoch was defined as a pass over as many pixel events as there are in the full training set.
+% % We attempt to forecast 50 events ahead in N-MNIST using a two layer LSTM with
+% 256 units per layer and a two layer GILR-LSTM with 234 units per layer. The
+% GILR-LSTM layer size was selected so that it had slightly fewer parameters than
+% the LSTM. Each model transformed the incoming event position into a
+% 40-dimensional embedding vector which was then combined with the polarity to
+% produce a 41 dimensional input. Both models output a $2 \mathbf{x} 34^2$ matrix
+% containing two future event location probability distributions conditioned on
+% the polarity of the future event. The training algorithm only considers the
+% probability distribution of the true future polarity and uses the cross entropy
+% loss function. The Adam \cite{kingma2014adam} optimization algorithm and Glorot
+% \cite{glorot2010understanding} initialization scheme were used. Training on a
+% minibatch of size $b$ with sequence length $T$ consisted of uniformly sampling
+% $b$ N-MNIST sequences and then extracting a single random $T$ element
+% subsequence (and its 50 element ahead forecast) from the sequence. Sequences
+% less than $T$ elements were padded out to length $T$, and there was a "burn-in"
+% of 30 events at the start of each sequence where no predictions were made. An
+% epoch was defined as a pass over as many pixel events as there are in the full
+% training set.
 
-% All experiments were performed using TensorFlow 1.0 \cite{abadi2016tensorflow} on a single Nvidia K80 GPU running for up to 18 hours.
-The LSTM model was computed using TensorFlow's dynamic\_rnn and BasicLSTMCell routines which are slower than but algorithmically similar to the cuDNN LSTM implementation.
+% % All experiments were performed using TensorFlow 1.0 \cite{abadi2016tensorflow}
+% on a single Nvidia K80 GPU running for up to 18 hours. The LSTM model was
+% computed using TensorFlow's dynamic\_rnn and BasicLSTMCell routines which are
+% slower than but algorithmically similar to the cuDNN LSTM implementation.
 
-% \subsection{Training performance}
-% The speed of neural net training is not solely determined by the training throughput but also by the frequency of optimization steps.
-As an example, a batch method may process inputs at the same rate as a minibatch method, but the minibatch method will generally converge much faster on large datasets.
-On an infinite dataset, the batch method never takes a single optimization step regardless of its throughput.
-With this example in mind, it is clear that achieving fast training is a balancing act between training data throughput and optimization step latency and that training performance should be evaluated with training curves and not just throughput numbers such as time per epoch.
+% % \subsection{Training performance} % The speed of neural net training is not
+% solely determined by the training throughput but also by the frequency of
+% optimization steps. As an example, a batch method may process inputs at the same
+% rate as a minibatch method, but the minibatch method will generally converge
+% much faster on large datasets. On an infinite dataset, the batch method never
+% takes a single optimization step regardless of its throughput. With this example
+% in mind, it is clear that achieving fast training is a balancing act between
+% training data throughput and optimization step latency and that training
+% performance should be evaluated with training curves and not just throughput
+% numbers such as time per epoch.
 
-% The LSTM and GILR-LSTM offer differ latency and throughput tradeoffs.
-LSTM throughput depends only on minibatch size but LSTM latency depends on both minibatch size and sequence length.
-Experiments were conducted training the LSTM models with minibatch size 256 and sequence lengths 128, 256, 512, and 1024.
-Figure \ref{fig:learning_curves} shows the smaller sequence lengths led to faster initial learning but inferior final performance.
+% % The LSTM and GILR-LSTM offer differ latency and throughput tradeoffs. LSTM
+% throughput depends only on minibatch size but LSTM latency depends on both
+% minibatch size and sequence length. Experiments were conducted training the LSTM
+% models with minibatch size 256 and sequence lengths 128, 256, 512, and
+% 1024. Figure \ref{fig:learning_curves} shows the smaller sequence lengths led to
+% faster initial learning but inferior final performance.
 
-% The throughput and latency of the GILR-LSTM are both influenced by batch size and sequence length.
-Several GILR-LSTMs were trained with batch size 16 and sequence length 1024.
-This combination was chosen because of the nearly maximum throughput, the low latency, and the ease of building a minibatch given the distribution of sequence lengths in the data set.
-The 234 unit GILR-LSTM reaches a better training loss in roughly 4.5 hours than any of the LSTMs could reach in 18 hours.
-This experimental evidence indicates the GILR-LSTM is as powerful of a model as the LSTM and can be trained in a fraction of the time.
+% % The throughput and latency of the GILR-LSTM are both influenced by batch size
+% and sequence length. Several GILR-LSTMs were trained with batch size 16 and
+% sequence length 1024. This combination was chosen because of the nearly maximum
+% throughput, the low latency, and the ease of building a minibatch given the
+% distribution of sequence lengths in the data set. The 234 unit GILR-LSTM reaches
+% a better training loss in roughly 4.5 hours than any of the LSTMs could reach in
+% 18 hours. This experimental evidence indicates the GILR-LSTM is as powerful of a
+% model as the LSTM and can be trained in a fraction of the time.
 
@@
 increased performance that the plr algorithm allows, we implement an GILR-LSTM
-on the problem presented in \Chris{CITE}, the 2016 PhysioNet/Computing in
+on the problem presented in , the 2016 PhysioNet/Computing in
 Cardiology Challenge 2016.  The aim of the Classification of Normal/Abnormal
@@
 giving a max sequence length of over 200,000. Since the testing server is not
-equipped with a GPU, we cannot run PLR in the testing step. We resample the
-records to 250 Hz so that we are able to evaluate the test results in a
+equipped with a GPU, we cannot run PLR in the inference. We resample the
+records to 500 Hz so that we are able to evaluate the test results in a
 reasonable amount of time, padding all records that are shorter than the longest.
@@
 normal hearts. We rebalance the data by oversampling the abnormal data by a factor
-of three, including each abnormal heart three times.
+of three, including each abnormal heart three times. We train a network consisting
+of three stacked GIRL-LSTM layers, each with 16 hidden units. Although reasonably
+small, since the input is only one-dimensional 16 hidden units allows us to
+develop representations. PLR allows us to train the network very rapidly over a
+long sequence length, so we don't have to downsample and lose information or
+spend lots of time developing methods to form features.
+We were able to fit exactly to the training set after
+a few hours of training, and on the unseen test data we obtained a score of
+PLACEHOLDER.
 
-Key points are:
-\begin{itemize}
-\item{With the fast training of the GILR-LSTM / PLR, we can train quickly on very
-    long sequence-lengths}
-\item{Allows us to get state-of-the-art results, even when performing very quickly
-    and with little hyperparameter tuning}
-\end{itemize}
-
-NEED TO GET THOSE RESULTS!\Chris{Quite likely that the results won't be available
-  for tomorrow, or that they won't be particularly impressive. We may have to
-  scratch this entire section unfortunately}
+\Chris{If we get good results (doesn't have to be best ever, but can be on par),
+  then keep}
+\Chris{If we don't get good results, or don't have results, we will need to kick
+  this section}
 
@@
 
+\bibitem[Goldberger et~al.(2000) Amaral, Glass, Hausdorff, Ivanov, Mark, Mietus,
+    Moody, Peng, Stanley]{physiobank} Goldberger et~al. 
+    \newblock PhysioBank, PhysioToolkit, and PhysioNet:
+Components of a New Research Resource for Complex Physiologic
+Signals.
+\newblock In \emph{Circulation} 101(23):e215-e220; 2000 (June 13). PMID:
+10851218; doi: 10.1161/01.CIR.101.23.e215
 \end{thebibliography}

### 2017-10-27 paper/main.tex
Fix gitignore and image name
Some small changes

@@
 pass and reduces the rate of optimization steps. Finally, training with larger
-minibatches damages generalization ability \cite{keskar2017large}. It is clearly
-important to investigate if it is possible to parallelise RNN training without
-introducing minibatches. Persistent RNNs \cite{diamos2016persistent} use a novel
+minibatches damages generalization ability \cite{keskar2017large}. Given these
+effects, it is desirable to obtain high training throughput with small minibatches.
+Persistent RNNs \cite{diamos2016persistent} use a novel
 implementation that can achieve high GPU utilization with very small minibatch
@@
 sequence length while attention mechanisms are expensive to apply on long
-sequences in online inference use cases.  One dimensional convolution can be
-viewed as a learnable linear finite impulse response (FIR) filter with a
-parallel evaluation algorithm, while linear recurrence is a learnable linear
-infinite impulse response (IIR).\Chris{Seems like an aside - take out?}
+sequences in online inference use cases.
 
-  This work parallelizes evaluation of linear recurrences through application of
-the parallel scan algorithm.
-
+A linear recurrence is a specific instance of a general form of computation known as a scan.
 Scans and reductions are computations involving repeated application of a binary
@@
 associative operation of vector addition
-($\boldsymbol{x} \oplus \boldsymbol{y} = \boldsymbol{x} + \boldsymbol{y}$), the
+($x \oplus y = x + y$), the
 semiassociative operation of matrix-vector multiplication
-($A \otimes \boldsymbol{x} = A\boldsymbol{x}$) and the associative operation of
+($A \otimes x = Ax$) and the associative operation of
 matrix-matrix multiplication ($A \odot B=AB$) satisfy Blelloch's three
@@
 \begin{itemize}
-\item{We classify the RNNs which satisfy the conditions above, and
-    show that many of the RNNs used in practice such as the Quasi-RNN are
+\item{We classify RNNs which satisfy the conditions above, and
+    show that many RNNs used in practice such as Quasi-RNNs (QRNNs) \cite{bradbury2017quasi} are
     contained in this class}
@@
     algorithm as a CUDA kernel, and show that it speeds up training of
-    the SRU and QRNN architectures by factors of up to 9x}
+    QRNN and Simple Recurrent Unit (SRU) \cite{lei2017} architectures by factors of up to 9x}
 \item{We describe how several recent linear RNNs can be described as
-    linear surrogates for nonlinear architectures. We introduce a linear
+    linear surrogates for non-linear architectures. We introduce a linear
     surrogate for the LSTM and show that we are able to train it
@@
 $C_\text{sscan} \in \mathcal{O}((C_\otimes + C_\oplus)T)$, compared to the parallel scan cost
-$C_\text{pscan} \in \mathcal{O}(2(C_\odot + C_\otimes + C_\oplus)(T/p + \lg p))$ \cite{blelloch1990prefix},
-where \(p\) is the number of available parallel cores. 
+$C_\text{pscan} \in \mathcal{O}(2(C_\odot + C_\otimes + C_\oplus)(T/p + \lg p))$ \cite{blelloch1990prefix} on $p$ processors.
 If $h_t$ is a vector of dimension $n$ then
@@
 proportional to $n$ and $C_\text{pscan}\in \mathcal{O}(6n(T/p + \lg p))$ and
-$C_\text{sscan} \in \mathcal{O}(2nT)$. Assuming $p \ll T$, then $C_\text{pscan} \le
-C_\text{sscan}$ when $p \ge 3$.\Chris{Can include the actual speedup
-  factor, \(\alpha = \frac{PT}{3(T + \lg(p))}\).}
+$C_\text{sscan} \in \mathcal{O}(2nT)$. This gives a parallel speedup factor of
+$\frac{pT}{3(T+\lg p)}$. Assuming $p \ll T$, then $C_\text{pscan} \le
+C_\text{sscan}$ when $p \ge 3$.
+
 As we are only considering diagonal matrices,
@@
 network models use separate recurrent matrices for each sequence element and using these
-separate matrices would require potentially prohibitive $n^2T$ memory.\Chris{I can't
-seem to find anyone explicitly saying this before from a quick google.}
+separate matrices would require potentially prohibitive $n^2T$ memory.
 Applying
@@
 \subsection{Implementation}
-In 2017, GPUs commonly used in deep learning consist of between 640 and 3200
-concurrently executing warps. Each warp operates on 32 single precision floating
+GPUs commonly used for deep learning in 2017 consist of between 640 and 3200 parallel
+processors known as warps. Each warp operates on 32 single precision floating
 point numbers in parallel.
@@
 into the TensorFlow \cite{abadi2016tensorflow} framework. Each warp acts as a
-processor, which means the algorithmic parallelizability $p$ is up to 3200 and the theoretical
+processor, which means the algorithmic $p$ is up to 3200 and the theoretical
 parallelization speedup factor is up to several hundred.  The 32 lanes of each
@@
 \subsubsection{Impact on effective "batch size"}
-Consider evaluating an RNN with recurrence $h_t = \sigma(Uh_{t-1} + Vx_t + b)$,
+Consider evaluating an RNN with recurrence $h_t = \sigma(Uh_{t-1} + Vx_t + b)$
 from $m$ inputs to $n$ hidden units on a sequence of length $T$ with minibatch
@@
 recurrence forces even the better approach to perform $T$ potentially small $(b,
-m) \times (m, n)$ matrix multiplications in serial which makes performance
-heavily dependent on minibatch size.
+m) \times (m, n)$ matrix multiplications in serial. This makes serial RNN
+performance heavily dependent on minibatch size.
 
@@
 two large matrix multiplications and a parallel linear recurrence. GILR performance
-is much less dependent on batch size as the matrix multiplies see an "effective
+is much less dependent on batch size as the matrix multiplication kernel sees an "effective
 batch size" of $bT$ and $T$ is typically large.
@@
 prevents application of the parallel linear recurrence algorithm and forces slow
-serial evaluation. To work around this inefficiency, note that $s_t$ serves a
-dual purpose. In $s_t = f(s_{t-1}, x_t)$, $s_{t-1}$ serves as an input to $f$
+serial evaluation. To work around this inefficiency, note that $s_t$ serves
+dual purposes. In $s_t = f(s_{t-1}, x_t)$, $s_{t-1}$ serves as an input to $f$
 summarizing the previous inputs while $s_t$ serves as the output of $f$ to be
@@
 computable.  We refer to this class of model as a linear surrogate RNN
-(LS-RNN). Quasi-RNNs \cite{bradbury2017quasi} are LS-RNNs using $\tilde{h}_{t-1}
+(LS-RNN). QRNNs \cite{bradbury2017quasi} are LS-RNNs using $\tilde{h}_{t-1}
 = W_k x_{t-k} + ... W_1 x_{t-1}$ and strongly typed
@@
 
-As an example LS-RNN, consider an LSTM:
-\begin{align*} f_t, i_t, o_t &= \sigma(U_{f,i,o} h_{t-1} + V_{f,i,o} x_t +
-b_{f,i,o}) \\ z_t &= \tau(U_z h_{t-1} + V_z x_t + b_z) \\ c_t &= f_t \odot
-c_{t-1} + i_t \odot z_t \\ h_t &= o_t \odot c_t
-\end{align*} An LSTM has state $s_t = (h_t, c_t)$. Since $c_t$ depends only
+Consider an LSTM:
+\begin{align*}
+  f_t, i_t, o_t &= \sigma(U_{f,i,o} h_{t-1} + V_{f,i,o} x_t + b_{f,i,o}) \\
+  z_t &= \tau(U_z h_{t-1} + V_z x_t + b_z) \\
+  c_t &= f_t \odot c_{t-1} + i_t \odot z_t \\
+  h_t &= o_t \odot c_t
+\end{align*}
+An LSTM has state $s_t = (h_t, c_t)$. Since $c_t$ depends only
 linearly on $c_{t-1}$, no surrogate is needed for $c_t$. $h_t$ has a non-linear
@@
 the surrogate, we obtain the GILR-LSTM:
-\begin{align*} f_t, i_t, o_t &= \sigma(U_{f,i,o} \tilde{h}_{t-1} + V_{f,i,o} x_t
-+ b_{f,i,o}) \\ z_t &= \tau(U_z \tilde{h}_{t-1} + V_z x_t + b_z) \\ c_t &= f_t
-\odot c_{t-1} + i_t \odot z_t \\ h_t &= o_t \odot c_t \\ g_t &= \sigma(V_g x_t +
-b_g) \\ \tilde{h}_t &= g_t \odot \tilde{h}_{t-1} + (1-g_t)\odot \tau(Wx_t + b_h)
+\begin{align*}
+  g_t &= \sigma(V_g x_t + b_g) \\
+  j_t &= \tau(V_j x_t + b_j) \\
+  \tilde{h}_t &= g_t \odot \tilde{h}_{t-1} + (1-g_t)\odot j_t \\
+  f_t, i_t, o_t &= \sigma(U_{f,i,o} \tilde{h}_{t-1} + V_{f,i,o} x_t + b_{f,i,o}) \\
+  z_t &= \tau(U_z \tilde{h}_{t-1} + V_z x_t + b_z) \\
+  c_t &= f_t \odot c_{t-1} + i_t \odot z_t \\
+  h_t &= o_t \odot c_t \\
 \end{align*}
 
-For $m$ inputs and hidden size $n$, the GILR-LSTM contains $2n(n+m)$ more
+For $m$ inputs and hidden size $n$, a GILR-LSTM contains $2n(n+m)$ more
 parameters than the equivalently sized LSTM to handle the mapping from $x$ to
@@
 parallel linear recurrence algorithm is able to achieve higher throughput than a
-similarly implemented serial version of the algorithm. We show that the plr kernel
-is up to 40 times faster than the serial equivalent for long sequence lengths, and
-that this speedup translates to implementations of LS-RNNs such as the QRNN (where
-we can get a speedup of up to 9 times).
+similarly implemented serial version of the algorithm. We show that the parallel kernel
+is up to 40 times faster than the serial equivalent for long sequence lengths and
+that this speedup translates to implementations of LS-RNNs such as QRNN (where
+we can get a speedup of up to 9x).
 
-In order to illustrate that the linearisation does not necessarily come at the
-cost of expressibility, we show that the GILR-LSTM architecture, trained with
-the PLR algorithm, is able to outperform the optimized CuDNN LSTM implementation
+In order to illustrate that the linearization does not necessarily come at the
+cost of expressibility, we show that the GILR-LSTM architecture, computed with
+the parallel linear recurrence algorithm, is able to outperform the optimized CuDNN LSTM
+implementation
 on a pathological example from the original LSTM paper
@@
 evaluates the parallel linear recurrence described in algorithm \ref{alg:plr},
-and one which evaluates the linear recurrence using the straightforward serial
-approach. This comparison is performed directly at the kernel level, avoiding
-any overhead due to calling Tensorflow. We find that at long sequence lengths,
-the PLR kernel has a distinct advantage, as shown in table
-\ref{table:kernel-throughput}, with a speedup factor of 40. At short sequence
-lengths, the parallel kernel does not perform well due to the extra overhead
-involved in marshalling the parallel algorithm. \Chris{Important:
-  What's the batch size here? Is it just one sequence? Need to say that}
+and one which evaluates the linear recurrence using the standard serial
+approach (with parallelization over features but not sequence length).
+This comparison is performed directly at the kernel level, avoiding
+any overhead from TensorFlow. We find that the parallel kernel has
+a distinct advantage at long sequence lengths with a speedup factor of up to 40x,
+as shown in table \ref{table:kernel-throughput}.
+At short sequence lengths, the parallel kernel does not perform well due to the
+overhead of multiple passes over data and communication between processors.
 \begin{table}[]
   \centering
-\caption{Speedup of the PLR kernel compared to the SLR kernel, without any
-  overhead from Tensorflow, for various sequence lengths and input size \(m\).}      
+  \caption{
+    Speedup of the parallel kernel compared to the serial kernel at various
+    values of $(\text{minibatch size } \mathbf{x} \text{ features}) = m$.
+    Performance was directly measured without any overhead from TensorFlow.
+  }
 \label{table:kernel-throughput}
@@
 
-Many of the RNNs that have been recently introduced can be implemented with a
-parallel linear recurrence. Table \ref{table:rnn-throughput} shows the
+Many recently introduced RNNs can be accelerated with the parallel linear recurrence
+algorithm.
+Table \ref{table:rnn-throughput} shows the
 throughput advantage by using a parallel linear recurrence when compared to a
@@
   \begin{minipage}{0.6\textwidth}
-    \includegraphics[width=1.0\textwidth]{./1M_synthetic.png}      
+    \includegraphics[width=1.0\textwidth]{./1m_synthetic.png}
   \end{minipage}

### 2017-10-27 paper/main.tex
More paper edits

@@
 
-\subsection{Throughput Benchmarks}
+\subsection{Throughput benchmarks}
+
+\subsubsection{Kernel performance}
 We first illustrate the throughput advantage of the parallel scan algorithm for
@@
 evaluates the parallel linear recurrence described in algorithm \ref{alg:plr},
-and one which evaluates the linear recurrence using the standard serial
+and one which evaluates the linear recurrence on GPU using the standard serial
 approach (with parallelization over features but not sequence length).
@@
 overhead of multiple passes over data and communication between processors.
-\begin{table}[]
-  \centering
-  \caption{
-    Speedup of the parallel kernel compared to the serial kernel at various
-    values of $(\text{minibatch size } \mathbf{x} \text{ features}) = m$.
-    Performance was directly measured without any overhead from TensorFlow.
-  }
+\begin{table}
+\begin{center}
+  \begin{tabular}{@{}llll@{}}
     \label{table:kernel-throughput}
-\begin{tabular}{@{}llll@{}}
 Sequence Length & \(m=4\)  & \(m=32\) & \(m=128\) \\ \midrule
@@
   \end{tabular}
+\end{center}
+\caption{
+  Speedup of the parallel kernel compared to the serial kernel at various
+  values of $(\text{minibatch size } \mathbf{x} \text{ features}) = m$.
+  Performance was directly measured without any overhead from TensorFlow.
+}
 \end{table}
+
+\subsubsection{Accelerating existing RNN architectures}
 
@@
 Table \ref{table:rnn-throughput} shows the
-throughput advantage by using a parallel linear recurrence when compared to a
-serial linear recurrence. Since PLR and SLR are simply different methods for
-computing the recurrence, there is no penalty incurred for this speedup - we
-obtain exactly the same output.
+throughput advantage from using parallel linear recurrence compared to
+serial linear recurrence. Both methods compute an identical recurrence,
+so using parallel rather than serial causes no numerical changes.
 
-Typically experiments will aim to maximize the memory
-utilisation of the GPU, so we simulate this by choosing the product of the batch
-size and the sequence length to be constant. We see that the architectures which
-have fewer terms (for which the linear recurrence is a higher proportion of the
-computational burden) are more effected by the switch to PLR.  This is
+We controlled for GPU memory usage by choosing minibatch size such that the product
+of minibatch size and sequence length is constant. We see that the simpler architectures
+(for which the linear recurrence is a higher proportion of the
+computational burden) are more affected by the switch to the parallel kernel.  This is
 particularly clear in the case of the QRNN, where including a wider
-convolutional window results in less time spent on the recurrence and less to
-gain from the parallelization.
-\begin{figure}
+convolutional window results in more time spent outside of the linear recurrence and
+therefore less to gain from the parallelization.
+\begin{table}
 \begin{center}
@@
 \end{center}
-\caption{Speedup obtained when using parallel linear recurrence compared to
-  serial linear recurrence. We analyse this for several recent architectures:
-  the Simple Recurrent Unit from \cite{lei2017}, the Quasi-RNN from
-  \cite{bradbury2017quasi}, with a convolutional filter of size 2 and 10, and
-  the GILR-LSTM which we introduce in section \ref{sec:ls-rnns}. To give a fair
-  model of a typical use-case maximising the memory usage of the GPU, we set
-  batch size \(b\) such that \(Tb = 65536\) for a sequence length \(T\).}
-\end{figure}
+\caption{
+  Speedup obtained from using parallel linear recurrence rather than
+  serial linear recurrence. We analyse this for
+  SRUs \cite{lei2017} and QRNNs \cite{bradbury2017quasi} with a convolutional
+  filters of size 2 and 10, and GILR-LSTM which we introduce in section \ref{sec:ls-rnns}.
+  We controlled for GPU memory by setting minibatch size $b$ such that
+  $bT = 65536$ for sequence length $T$.
+}
+\end{table}
 
-Finally, we compare the throughput of the linear surrogate LSTM to the current
-fastest available implementation of the LSTM, the CuDNN-LSTM. Figure
-\ref{fig:tp_perf} shows that our GILR-LSTM has higher throughput for most of the
-low-batch-size regime, with a 30x speedup at a sequence length of 65,536. Recent
-work \cite{keskar2017large} has suggested that small mini-batches may aid
-generalization by encouraging the discovery of shallow minima. The GILR-LSTM
-allows us to achieve high throughput for all choices of minibatch size and
+\Chris{For table \ref{table:rnn-throughput} how many layers and what was hidden size?}
+
+\subsubsection{GILR-LSTM throughput}
+
+Finally, we compare the throughput of the GILR-LSTM to the current fastest
+LSTM implementation (which comes from the CuDNN library). Figure
+\ref{fig:tp_perf} shows that the GILR-LSTM has higher throughput for most of the
+small minibatch regime, with a 30x speedup at a sequence length of 65,536.
+The GILR-LSTM allows us to achieve high throughput for all choices of minibatch size and
 sequence length.
@@
 
-\begin{table}[t]
+\Chris{Put title on Figure \ref{fig:tp_perf}}
+
+\begin{figure}[t]
 \centering
@@
 \label{fig:tp_perf}
-\end{table}
+\end{figure}
 
@@
 
-\subsection{Synthetic Example}
+\subsection{Synthetic Experiment}
 In order to demonstrate how PLR may be used to speed up tasks that the LSTM is
-well-suited for, we tackle a pathological inference problem. Originally
-introduced as example 2b from Hochreiter and Schmidhtbuer's introduction of the
-LSTM, \cite{hochreiter1997long}, it involves storing a number for a very long
-sequence of time steps. We have an alphabet of size \(p\) with each character
-represented as a particular one-hot vector in this space. We input a sequence of
-\(n\) of these vectors, chosen at random (with replacement). The first vector in
-the sequence is always the same, up to the sign of the component (i.e. it is \(\pm
-\boldsymbol{p_0} \).
-
-The two signs on the first component separate the sequences into two sets. The
-whole sequence is fed into the LSTM and we aim to learn to classify the
-sequences. This requires remembering the first element over the length of the
-sequence. Due to the large throughput of the GILR-LSTM, we would expect that it
+well-suited for, we tackle a pathological inference problem first introduced
+in Hochreiter and Schmidhuber's initial LSTM paper \cite{hochreiter1997long}.
+The input consists of sequences of length $n$ with each sequence element a randomly
+chosen one-hot vector in $p$ dimensional space.
+The first vector in each sequence is always the same, up to the sign of the component (i.e. it is $\pm
+\boldsymbol{p_0}$). The sequential model must read in an entire sequence and then output
+the sign of the first sequence element.
+This sequence classification problem requires remembering the first element over the length of the
+sequence and is a canonical example of the long-term dependency problem.
+Due to the large throughput of the GILR-LSTM, we would expect that it
 would do better when the sequence length is large. In the original formulation
 of the problem (dealing in the regime with around one hundred timesteps)
-, \(p\) is set equal to \(n\). Since this would make the size of the input data
+, $p$ is set equal to $n$. As this would make the size of the input data
 grow impracticaly large as \(\mathcal{O}(n^2)\) for long sequences, we fix \(p =
@@
 
-We generated three sets of data: for \(n\) equal to 1,024, 8,192, and
-1,048,576. For each of these we compared a two-layer GILR-LSTM with 512 hidden
-units to a canonical LSTM network with roughly the same number of parameters. We
-could imagine using a two-layer LSTM with 512 hidden units in each layer, or a
-one-layer LSTM with 1024 hidden units.
+We generated sequences for $n$ equal to 1,024, 8,192, and
+1,048,576. For each of these we compared a two layer GILR-LSTM with 512 hidden
+units to a two layer LSTM with 512 hidden units per layer implemented by CuDNN.
 
-Initial experiments showed that the two-layer LSTM was much quicker to converge
-than the one-layer LSTM, so here we compare the GILR-LSTM to the two-layer
-CudnnLSTM.  We ran all experiments on a NVIDIA K80 GPU, choosing the largest
-minibatch size which fit into the GPU memory, with five runs per configuration
+We ran all experiments on a NVIDIA K80 GPU, choosing the largest
+minibatch size which fit into the GPU memory with five runs per configuration
 allowing us to find the average and standard deviation of the time and number of
-iterations to convergence. For all CUDA runs, a brief search over learning rate
+iterations to convergence. A brief search over learning rate
 and batch size was carried out to find the parameters which allow the network to
-converge most rapidly. The criterion for convergence was five consecutive
-minibatches giving 100\% accuracy. As can be seen from the learning curves
-in figure \ref{fig:synthetic_training}, this was a reasonable criterion. For the
-longest sequence-length, we didn't observe the CuDNN-LSTM converging.
+converge most rapidly for all runs. The criterion for convergence was five consecutive
+minibatches giving 100\% accuracy. The learning curves
+in figure \ref{fig:synthetic_training} give support to this being a reasonable
+convergence criteria.
+For the longest sequence length, we could not run long enough to observe the LSTM converge.
 
-The results illustrate that the GILR-LSTM is able to robustly outperform the
-CuDNN-LSTM, which is currently the fastest LSTM implementation available. This
-is somewhat surprising, as we might expect that the LSTM would do particularly
-well at an example that was specifically constructed for it. We can conclude that
-the nonlinearities in the LSTM's design are not vital for its performance given
-that the linear GILR-LSTM is able to converge, and that if these nonlinearities are
-not completely necessary, they entail a large performance cost by diallowing the
-plr algorithm from being used. Furthermore, the sequence length of over one million
-is the largest sequence length we have seen seriously discussed. The GILR
-handles it straightforwardly.
+The results illustrate that the GILR-LSTM is able to converge much more quickly than
+the fastest available LSTM implementation.
+This is somewhat surprising given the LSTM was specifically constructed for problems
+of this sort. This GILR-LSTM convergence
+leads to the conclusion that the non-linearities present in
+LSTM are not necessary for solving this instance of the long-term dependency problem.
+The time to convergence further leads to the conclusion that inclusion of a non-linearity
+at every step incurs a significant training time slowdown.
+Furthermore, the GILR-LSTM is able to learn to carry dependencies over a one million
+element sequence. As far as we know, this one million step sequence experiment
+is the longest sequential learning problem to be handled by neural networks to date.
 
@@
 \subsection{The Medical Setting}
-A common application of RNNs is in medicine. In order to illustrate the
-increased performance that the plr algorithm allows, we implement an GILR-LSTM
-on the problem presented in , the 2016 PhysioNet/Computing in
+Medical waveform data consists of long sequences and is a reasonable application area for RNNs.
+We applied the GILR-LSTM
+on the problem presented in the 2016 PhysioNet/Computing in
 Cardiology Challenge 2016.  The aim of the Classification of Normal/Abnormal
-Heart Sound Recordings challenge is to determine which of a set of EEGs are
-produced by people with abnormal hearts.  The data is recorded from a variety of
-sources and split into a training and testing set. The testing set is not
+Heart Sound Recordings challenge is classify electroencephalogram (EEG) recordings
+as coming from a normal or abnormal heart.
+The waveforms are recorded on several sets of sensors
+and split into a training and testing set. The testing set is not
 publicly available, but can be evaluated through a web portal. The top results
-cover a large range of techniques, but generally rely on featurization of the
-raw EEG records. We want to see if we are able to leverage our advantage at high
-sequence-lenghts to train directly on the raw EEG. The EEGs are sampled at 2,000
-Hz and range in length from several seconds to more than one hundred seconds,
+cover a large range of techniques, but generally rely on manual featurization of the
+raw EEG records. The GILR-LSTM's ability to handle long sequences makes it a
+candidate for training directly on the raw EEG signal.
+The EEGs are sampled at 2,000 Hz
+and range in length from several seconds to more than one hundred seconds,
 giving a max sequence length of over 200,000. Since the testing server is not
@@
 normal hearts. We rebalance the data by oversampling the abnormal data by a factor
-of three, including each abnormal heart three times. We train a network consisting
-of three stacked GIRL-LSTM layers, each with 16 hidden units. Although reasonably
-small, since the input is only one-dimensional 16 hidden units allows us to
-develop representations. PLR allows us to train the network very rapidly over a
-long sequence length, so we don't have to downsample and lose information or
-spend lots of time developing methods to form features.
-We were able to fit exactly to the training set after
-a few hours of training, and on the unseen test data we obtained a score of
-PLACEHOLDER.
+of three, including each abnormal heart three times. We train a network feeding
+the one dimensional input signal into three stacked GILR-LSTM layers of 16 hidden units each.
+Although reasonably small, 16 hidden units allows developing
+develop representations. Parallel linear recurrence allows us to rapidly train the network
+over the entire long sequence length, so we don't have to work on truncated samples,
+downsampled sequenes, or
+spend lots of time manually developing EEG specific features.
+The model obtained perfect classification on the train dataset after
+a few hours \Chris{how many?} of training.
+On the unseen test data we obtained a score of PLACEHOLDER.
 
@@
 \section{Conclusion}
-Parallel linear recurrence is an extremely powerful algorithm and the GILR-LSTM
-is just one of many possible models that can be built with it. While we were
-previously forced to use large minibatches in order to obtain reasonable
-training speed, with PLR in many cases we can train rapidly on small minibatch
-sizes, freeing us to utilise extremely long sequence lengths and more complex
-models. Our CUDA kernel is up to 40 times faster than a serial kernel on
+Parallel linear recurrence can accelerate existing RNNs and be used in the design
+of new architectures such as the GILR-LSTM. Parallel linear recurrence enables
+rapid training with small minibatches, which further enables training models on
+extremely long sequence lengths.
+Our CUDA kernel is up to 40 times faster than a serial kernel on
 sequence lengths of 65,536, and this benefit transfers to implemented models,
@@
 \cite{lei2017} and other recent work comprehensively show that
-linearly-recurrent RNNs are, in many respects, equally as powerful as nonlinear
+linear RNNs are, in many respects, equally as powerful as nonlinear
 LSTMs. Our results on the synthetic example continue this trend, showing that

### 2017-10-27 paper/main.tex
Small change to conclusion

@@
 extremely long sequence lengths.
-Our CUDA kernel is up to 40 times faster than a serial kernel on
-sequence lengths of 65,536, and this benefit transfers to implemented models,
-speeding up the SRU \cite{lei2017} by a factor of 9. The impressive results in
+The impressive results in
 \cite{lei2017} and other recent work comprehensively show that
-linear RNNs are, in many respects, equally as powerful as nonlinear
-LSTMs. Our results on the synthetic example continue this trend, showing that
+linear RNNs can often be used in place of traditional non-linear RNNs.
+Our results on the synthetic example continue this trend, showing that
 the GILR-LSTM is able to handle sequence length regimes that are longer than any

### 2017-10-27 paper/main.tex
Delete comments

@@
 In order to illustrate that the linearization does not necessarily come at the
-cost of expressibility, we show that the GILR-LSTM architecture, computed with
+cost of expressibility, we show that the GILR-LSTM architecture computed with
 the parallel linear recurrence algorithm, is able to outperform the optimized CuDNN LSTM
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% The computational performance of the LSTM and GILR-LSTM models were compared       %
-% across a wide range of minibatch sizes and sequence lengths. Although the 234    %
-% unit GILR-LSTM has a similar number of parameters to the 256 unit LSTM, we         %
-% measured the computational performance of a 256 unit GILR-LSTM to observe any      %
-% impact of the linear surrogate calculation. The serial evaluation of the LSTM    %
-% causes its throughput to be independent of the sequence length as a doubling of  %
-% sequence length causes a doubling of runtime. Notably, the GILR-LSTM can achieve a %
-% higher throughput by running on one 8192 event sequence at a time than an LSTM   %
-% running with 256 sequence minibatches. Similar speedups were found for           %
-% inference.                                                                       %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 Unlike the previous case, these speedups do come at a cost. The GILR-LSTM
@@
 \end{figure}
-
-% % Experiments were performed on the N-MNIST \cite{orchard2015converting}
-% dataset. N-MNIST captures the MNIST digit dataset \cite{lecun1998mnist} by
-% panning % an event driven camera over each digit. Each example in N-MNIST is a
-% sequence of single pixel events (x, y, polarity, timestamp) where % x and y each
-% indicate a pixel position in [0, 33], polarity indicates whether the pixel was
-% switching on or off, and timestamp is the time of the event in
-% microseconds. Videos produced from the event data show that positive polarity
-% events are often located on the leading edge of the digit motion and negative
-% polarity events on the trailing edge. Between the 60,000 digits in the training
-% set, N-MNIST contains approximately 250 million pixel events with sequence
-% lengths ranging from 500 to 8000 and averaging 4000 events.
-
-% % We attempt to forecast 50 events ahead in N-MNIST using a two layer LSTM with
-% 256 units per layer and a two layer GILR-LSTM with 234 units per layer. The
-% GILR-LSTM layer size was selected so that it had slightly fewer parameters than
-% the LSTM. Each model transformed the incoming event position into a
-% 40-dimensional embedding vector which was then combined with the polarity to
-% produce a 41 dimensional input. Both models output a $2 \mathbf{x} 34^2$ matrix
-% containing two future event location probability distributions conditioned on
-% the polarity of the future event. The training algorithm only considers the
-% probability distribution of the true future polarity and uses the cross entropy
-% loss function. The Adam \cite{kingma2014adam} optimization algorithm and Glorot
-% \cite{glorot2010understanding} initialization scheme were used. Training on a
-% minibatch of size $b$ with sequence length $T$ consisted of uniformly sampling
-% $b$ N-MNIST sequences and then extracting a single random $T$ element
-% subsequence (and its 50 element ahead forecast) from the sequence. Sequences
-% less than $T$ elements were padded out to length $T$, and there was a "burn-in"
-% of 30 events at the start of each sequence where no predictions were made. An
-% epoch was defined as a pass over as many pixel events as there are in the full
-% training set.
-
-% % All experiments were performed using TensorFlow 1.0 \cite{abadi2016tensorflow}
-% on a single Nvidia K80 GPU running for up to 18 hours. The LSTM model was
-% computed using TensorFlow's dynamic\_rnn and BasicLSTMCell routines which are
-% slower than but algorithmically similar to the cuDNN LSTM implementation.
-
-% % \subsection{Training performance} % The speed of neural net training is not
-% solely determined by the training throughput but also by the frequency of
-% optimization steps. As an example, a batch method may process inputs at the same
-% rate as a minibatch method, but the minibatch method will generally converge
-% much faster on large datasets. On an infinite dataset, the batch method never
-% takes a single optimization step regardless of its throughput. With this example
-% in mind, it is clear that achieving fast training is a balancing act between
-% training data throughput and optimization step latency and that training
-% performance should be evaluated with training curves and not just throughput
-% numbers such as time per epoch.
-
-% % The LSTM and GILR-LSTM offer differ latency and throughput tradeoffs. LSTM
-% throughput depends only on minibatch size but LSTM latency depends on both
-% minibatch size and sequence length. Experiments were conducted training the LSTM
-% models with minibatch size 256 and sequence lengths 128, 256, 512, and
-% 1024. Figure \ref{fig:learning_curves} shows the smaller sequence lengths led to
-% faster initial learning but inferior final performance.
-
-% % The throughput and latency of the GILR-LSTM are both influenced by batch size
-% and sequence length. Several GILR-LSTMs were trained with batch size 16 and
-% sequence length 1024. This combination was chosen because of the nearly maximum
-% throughput, the low latency, and the ease of building a minibatch given the
-% distribution of sequence lengths in the data set. The 234 unit GILR-LSTM reaches
-% a better training loss in roughly 4.5 hours than any of the LSTMs could reach in
-% 18 hours. This experimental evidence indicates the GILR-LSTM is as powerful of a
-% model as the LSTM and can be trained in a fraction of the time.
-
-% \begin{figure}[t]
-% \centering
-% \includegraphics[width=12cm]{lc2.png}
-% \caption{The LSTM models had a direct relationship between sequence length and training curve: the shorter the sequence length, the faster the initial learning and
-% the higher the final loss. The faster initial learning is explained by the decreased
-% latency of optimization steps and the higher final loss is explained by the
-% inability to learn long dependencies. The GILR-LSTM has 234 units per hidden
-% layer to have the same number of parameters as the 256 unit LSTMs.}
-% \label{fig:learning_curves}
-% \end{figure}
-
-% \subsection{Test performance}
-% \begin{table}[h]
-%   \caption{N-MNIST test results}
-%   \label{test-results}
-%   \centering
-%   \begin{tabular}{lllll}
-%     \toprule
-%     & Train Cross Entropy     & Test Cross Entropy     & Top-5 Accuracy & Top-20 Accuracy \\
-%     \midrule
-%     GILR-LSTM 			&  \textbf{4.567} & 4.856           & 10.51\% & 36.70\%   \\
-%     LSTM, seqlen=128    & 4.664 & \textbf{4.846}  & \textbf{10.62\%} & \textbf{37.08\%}   \\
-%     LSTM, seqlen=256    & 4.620                & 4.906           & 10.11\% & 35.40\%   \\
-%     LSTM, seqlen=512    & 4.602                & 4.883           & 10.45\% & 36.32\%   \\
-%     LSTM, seqlen=1024   & 4.611                & 4.880           & 10.32\% & 35.92\%   \\
-%     \bottomrule
-%   \end{tabular}
-% \end{table}
-% Beyond cross entropy on train and test, we also evaluated top-$k$ accuracy for $k=5, 20$. A probability distribution is top-$k$ correct if it assigns the realized location one of the
-% $k$ largest probabilities. N-MNIST contains 1156 pixel positions, so top-5 and top-20 accuracy are equivalent to localizing the future to 0.43\% and 1.7\% of pixels.
-
-% No regularization was attempted and all of the models overfit, as indicated by the best test performance from the model with the worst train performance.
-Table \ref{test-results} contains the full test results.
-Our focus was the fast training of powerful models, and we leave regularizing parallel linear recurrences and LS-RNNs to future work.
-Although not tested, it is possible that the GILR-LSTM generalized better than the LSTMs trained on long sequences due to the much smaller minibatch size of GILR-LSTM leading to a wider minima.
 

### 2017-10-27 paper/main.tex
Clarify some comments

@@
 
-\Chris{For table \ref{table:rnn-throughput} how many layers and what was hidden size?}
+\Eric{For table \ref{table:rnn-throughput} how many layers and what was hidden size?}
 
 \subsubsection{GILR-LSTM throughput}
+
+\Eric{I think we should delete this section. Already over length limit, doesn't show
+anything great}
 
@@
 
-\Chris{Put title on Figure \ref{fig:tp_perf}}
+\Eric{Chris: Put title on Figure \ref{fig:tp_perf}}
 
@@
 
-\subsection{The Medical Setting}
+\subsection{Abnormal heart sound detection}
 Medical waveform data consists of long sequences and is a reasonable application area for RNNs.
@@
 The model obtained perfect classification on the train dataset after
-a few hours \Chris{how many?} of training.
+a few hours \Eric{how many?} of training.
 On the unseen test data we obtained a score of PLACEHOLDER.

### 2017-10-27 paper/main.tex
Stage 1 reformat to ICLR format

@@
 \documentclass{article}
-
-% if you need to pass options to natbib, use, e.g.:
-\PassOptionsToPackage{numbers, compress}{natbib}
-% before loading nips_2017
-%
-% to avoid loading the natbib package, add option nonatbib:
-% \usepackage[nonatbib]{nips_2017}
-
-%\usepackage{nips_2017}
-\usepackage[final]{nips_2017}
+\usepackage{iclr2018_conference,times}
 \bibliographystyle{abbrvnat}
 
-% to compile a camera-ready version, add the [final] option, e.g.:
-%\usepackage[final]{nips_2017}
+%\iclrfinalcopy % Uncomment for camera-ready version, but NOT for submission.
 
@@
 \providecommand{\reduce}{\text{REDUCE}}
+
 % declaration of the new block
@@
 
-% The \author macro works with any number of authors. There are two
-% commands used to separate the names and addresses of multiple
-% authors: \And and \AND.
-%
-% Using \And between authors leaves it to LaTeX to determine where to
-% break the lines. Using \AND forces a line break at that point. So,
-% if LaTeX puts 3 of 4 authors names on the first line, and the last
-% on the second line, try using \AND instead of \And before the third
-% author name.
-
-\author{
-  Eric Martin \\%\thanks{Use footnote for providing further
-    %information about author (webpage, alternative
-    %address)---\emph{not} for acknowledging funding agencies.} \\
-  %Some Affiliation\\
-  \texttt{eric@ericmart.in}
-  %% examples of more authors
+\author{Eric Martin \\
+\texttt{eric@ericmart.in} \\
 \And
-  Chris Cundy \\
-  \texttt{c.cundy@berkeley.edu}
-  %%\texttt{some@email.com}
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
-  %% \AND
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
-  %% \And
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
-  %% \And
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
+Chris Cundy
+Department of XX \\
+University of XX \\
+City, Country \\
+\texttt{c.cundy@berkeley.edu} \\
 }
@@
 \begin{document}
-% \nipsfinalcopy is no longer used
-
 \maketitle
@@
 overhead of multiple passes over data and communication between processors.
-\begin{table}
+\begin{table}[t]
+\caption{Parallel kernel speedup on $m$ features}
 \begin{center}
@@
 \end{center}
-\caption{
-  Speedup of the parallel kernel compared to the serial kernel at various
-  values of $(\text{minibatch size } \mathbf{x} \text{ features}) = m$.
-  Performance was directly measured without any overhead from TensorFlow.
-}
 \end{table}
@@
 
-%Future work can the development of parallel computable
-%differentiable memory modules, and the combination of linear recurrence with
-%convolutional sequence models.
-
-% Besides future research, existing models such as
-% Quasi-RNNs and strongly typed RNNs that already contain linear recurrences can
-% immediately benefit from parallel linear recurrence. This work demonstrates the
-% GILR-LSTM significantly accelerates the training of small to medium sized
-% LSTMs. Although similar techniques have been used before, the now explicit
-% concept of linear surrogacy provides a framework for future development and
-% analysis of fast sequence models.
-
-We intend to expand upon the experiments section and open-source the parallel
-linear recurrence kernel in the near future.
-
 \subsubsection*{Acknowledgments}

### 2017-10-27 paper/main.tex
Lots of rewriting

@@
 sequential dependencies can be parallelized over the sequence length using the
-parallel scan algorithm, leading to rapid training on long sequences with small
-minibatch size.\Chris{Thought: we could rephrase this as allowing us to break
-the necessity of using large batches to train fast} We develop a
-parallel linear recurrence CUDA kernel and show that it can be applied to
-immediately speed up several RNN architectures by up to a factor of 9. We abstract
-recent work on linear RNNs into a new framework of linear surrogate RNNs and develop
-a linear surrogate for the long short-term memory unit, the GILR-LSTM, which is
-able to achieve throughput 30 times greater than the currently fastest-available LSTM
-implementation. We show that the GILR-LSTM is able to perform well on real-world
-problems,
+parallel scan algorithm, leading to rapid training on long sequences even with small
+minibatch size. We develop a parallel linear recurrence CUDA kernel and show that
+it can be applied to immediately speed up training and inference of
+several RNN architectures by up to 9x.
+We abstract recent work on linear RNNs into a new framework of linear surrogate RNNs
+and develop
+a linear surrogate model for the long short-term memory unit, the GILR-LSTM, that
+utilizies parallel linear recurrence.
+We push the limits of sequence learning by successfully training a GILR-LSTM
+on a synthetic sequence classification task with a one million timestep dependency.
+SOMETHING MEDICAL?
 \end{abstract}
@@
 next state, with \(s_t = f(\tilde{s}_{t-1}, x_t)\). We are still able to choose a
-nonlinear \(f\), our only limitation being that \(\tilde{s}_t\) must be linearly
+non-linear \(f\), our only limitation being that \(\tilde{s}_t\) must be linearly
 computable.  We refer to this class of model as a linear surrogate RNN
@@
 \section{Experiments}
-We perform several experiments, first confirming that our implementation of the
-parallel linear recurrence algorithm is able to achieve higher throughput than a
-similarly implemented serial version of the algorithm. We show that the parallel kernel
-is up to 40 times faster than the serial equivalent for long sequence lengths and
-that this speedup translates to implementations of LS-RNNs such as QRNN (where
-we can get a speedup of up to 9x).
+We perform several experiments. First we find that our parallel linear recurrence
+kernel is able to able to achieve up to 40x higher throughput than a serial implementation
+when applied to long sequences. Secondly, we confirm that this kernel speedup
+translates to up to a 9x speedup to LS-RNNs such as QRNNs.
 
@@
 cost of expressibility, we show that the GILR-LSTM architecture computed with
-the parallel linear recurrence algorithm, is able to outperform the optimized CuDNN LSTM
-implementation
-on a pathological example from the original LSTM paper
-\cite{hochreiter1997long}. Finally we show that our GILR-LSTM is able to
+the parallel linear recurrence algorithm is able to train significantly faster
+than an optimized LSTM implementation on a pathological long-term dependency
+problem on a pathological example from the original LSTM paper \cite{hochreiter1997long}.
+Finally we show that the GILR-LSTM is able to
 quickly achieve state of the art results on an existing medical data
@@
 of length \(T\), we define the number of events as \(bT\) and the throughput as
-the number of events per second. We implement two CUDA kernels, one which
+the number of events processed per second. We implement two CUDA kernels, one which
 evaluates the parallel linear recurrence described in algorithm \ref{alg:plr},
-and one which evaluates the linear recurrence on GPU using the standard serial
-approach (with parallelization over features but not sequence length).
-This comparison is performed directly at the kernel level, avoiding
-any overhead from TensorFlow. We find that the parallel kernel has
+and one which evaluates the same linear recurrence on GPU in serial over sequence length
+and in parallel over features and minibatch. The performance of each kernel depends on
+two factors: the sequence length and the product of number of features and minibatch size.
+The performance measurements for this experiment are made directly at the kernel level,
+avoiding any overhead from TensorFlow. We find that the parallel kernel has
 a distinct advantage at long sequence lengths with a speedup factor of up to 40x,
 as shown in table \ref{table:kernel-throughput}.
-At short sequence lengths, the parallel kernel does not perform well due to the
+The parallel kernel does not perform well at short sequence lengths due to the
 overhead of multiple passes over data and communication between processors.
 \begin{table}[t]
-\caption{Parallel kernel speedup on $m$ features}
+\caption{Parallel kernel speedup on $m$ features (minibatch size $= 1$)}
 \begin{center}
@@
 \subsubsection{Accelerating existing RNN architectures}
+Several recently introduced LS-RNNs can be accelerated with the parallel linear recurrence
+algorithm. We implemented SRUs, QRNNs (with filter width 2 and 10), and GILR-LSTMs
+that can be computed with either the standard
+serial linear recurrence algorithm or parallel linear recurrence. Both methods compute
+an identical recurrence, so switching from a serial to parallel implementation
+does not cause any numerical changes and takes only a single line of code changes.
+Notably, both SRUs and QRNNs
+claim an order of magnitude speedup compared to CuDNN LSTM when implemented with serial linear
+recurrence. Any further speedup from parallel linear recurrence applies on top of the existing
+speedup.
+We timed train throughput (forwards and backwards propagation), but the linear time
+of each pass also makes the results applicable to forwards (inference) performance. However,
+parallel linear recurrence can only accelerate inference in scenarios where the entire
+input sequence is known at the start of the inference phase.
+We controlled for GPU memory usage within these experiments by fixing $bT = 65536$ for
+minibatch size $b$ and sequence length $T$.
 
-Many recently introduced RNNs can be accelerated with the parallel linear recurrence
-algorithm.
-Table \ref{table:rnn-throughput} shows the
+Table \ref{table:rnn-throughput} shows that
 throughput advantage from using parallel linear recurrence compared to
-serial linear recurrence. Both methods compute an identical recurrence,
-so using parallel rather than serial causes no numerical changes.
-
-We controlled for GPU memory usage by choosing minibatch size such that the product
-of minibatch size and sequence length is constant. We see that the simpler architectures
+serial linear recurrence reaches up to 9x. Simpler architectures
 (for which the linear recurrence is a higher proportion of the
-computational burden) are more affected by the switch to the parallel kernel.  This is
-particularly clear in the case of the QRNN, where including a wider
-convolutional window results in more time spent outside of the linear recurrence and
-therefore less to gain from the parallelization.
-\begin{table}
+total computational load) are more affected by the switch to the parallel kernel.
+This is
+particularly clear in the case of the QRNN, where including wider
+convolutional filters results in more time spent outside of the linear recurrence and
+therefore speedup from linear recurrence parallelization.
+\begin{table}[t]
+\caption{Parallel kernel speedup for a variety of LS-RNNs}
 \begin{center}
@@
 \label{table:rnn-throughput}
-Sequence Length & GILR-LSTM & SRU & QRNN(2) & QRNN(10)\\ \midrule
-16 & 0.61 & 0.28 & 0.38 & 0.78\\
-256 & 0.91 & 0.84 & 0.86 & 0.99\\
-4,096 & 0.98 & 1.38 & 1.18 & 1.05\\
-65,536 & 1.41 & 9.21 & 6.68 & 2.05\\ \bottomrule
+Sequence Length & SRU & QRNN (filter size 2) & QRNN (filter size 10) & GILR-LSTM\\ \midrule
+16 & 0.28 & 0.38 & 0.78 & 0.61\\
+256 & 0.84 & 0.86 & 0.99 & 0.91\\
+4,096 & 1.38 & 1.18 & 1.05 & 0.98\\
+65,536 & 9.21 & 6.68 & 2.05 & 1.41\\ \bottomrule
 \end{tabular}
 \end{center}
-\caption{
-  Speedup obtained from using parallel linear recurrence rather than
-  serial linear recurrence. We analyse this for
-  SRUs \cite{lei2017} and QRNNs \cite{bradbury2017quasi} with a convolutional
-  filters of size 2 and 10, and GILR-LSTM which we introduce in section \ref{sec:ls-rnns}.
-  We controlled for GPU memory by setting minibatch size $b$ such that
-  $bT = 65536$ for sequence length $T$.
-}
 \end{table}
@@
 anything great}
+
+Eric: I think we should delete this. We're over length, results aren't great.
 
@@
 \section{Conclusion}
-Parallel linear recurrence can accelerate existing RNNs and be used in the design
-of new architectures such as the GILR-LSTM. Parallel linear recurrence enables
-rapid training with small minibatches, which further enables training models on
-extremely long sequence lengths.
-The impressive results in
-\cite{lei2017} and other recent work comprehensively show that
-linear RNNs can often be used in place of traditional non-linear RNNs.
-Our results on the synthetic example continue this trend, showing that
-the GILR-LSTM is able to handle sequence length regimes that are longer than any
-we have seen seriously suggested in the literature. We feel that the framework
-of describing linear RNNs as linear surrogates for nonlinear RNNs is a fruitful
-approach, and that this, paired with the PLR algorithm, will substantially
-advance the ability to train RNNs, and to understand timeseries data.
+A significant portion of the success of deep learning can be attributed to massive
+amounts of computation. Most of this computation is accessed through two
+building blocks with highly efficient parallel implementations:
+matrix multiplication and convolution. Recent research has demonstrated that
+linear RNNs can achieve similar prediction accuracy to non-linear RNNs on a wide
+variety of tasks in a fraction of the training time. We propose the framework of
+LS-RNNs as a way to tame the growing zoo of sequential neural nets. We identify linear
+recurrence as another parallelizable building block for current and future sequential models
+and we use it to obtain significant speedups on already fast models.
+With the power of parallel
+linear recurrence we are able to solve a sequential dependency problem multiple
+orders of magnitude larger than anything done prior.
+Future applications of parallel linear recurrence within neural nets could include
+parallel training of memory augmented models or providing a new sort of image filter
+on very high resolution images.
+We hope that parallel linear recurrence can be to large scale sequence
+modelling what fast convolution algorithms are to image recognition.
 
 \subsubsection*{Acknowledgments}
-We would like to acknowledge Kevin Bowers, Alex Meiburg, JD Co-Reyes, Carson
-McNeil, Andy Palan, and several others for fruitful conversations and guidance.
+Acknowledgements removed for anonymous submission.
+%We would like to acknowledge Kevin Bowers, Alex Meiburg, JD Co-Reyes, Carson
+%McNeil, Andy Palan, and several others for fruitful conversations and guidance.
 

### 2017-10-27 paper/main.tex
Assorted small changes

--- 2017-10-27 paper/main.tex
@@
 pass and reduces the rate of optimization steps. Finally, training with larger
-minibatches damages generalization ability \cite{keskar2017large}. Given these
-effects, it is desirable to obtain high training throughput with small minibatches.
-Persistent RNNs \cite{diamos2016persistent} use a novel
+minibatches damages generalization ability \cite{keskar2017large}. It is clearly
+important to investigate if it is possible to parallelise RNN training without
+introducing minibatches. Persistent RNNs \cite{diamos2016persistent} use a novel
 implementation that can achieve high GPU utilization with very small minibatch
@@
 sequence length while attention mechanisms are expensive to apply on long
-sequences in online inference use cases.
+sequences in online inference use cases.  One dimensional convolution can be
+viewed as a learnable linear finite impulse response (FIR) filter with a
+parallel evaluation algorithm, while linear recurrence is a learnable linear
+infinite impulse response (IIR).\Chris{Seems like an aside - take out?}
 
-A linear recurrence is a specific instance of a general form of computation known as a scan.
+  This work parallelizes evaluation of linear recurrences through application of
+the parallel scan algorithm.
+
 Scans and reductions are computations involving repeated application of a binary
@@
 associative operation of vector addition
-($x \oplus y = x + y$), the
+($\boldsymbol{x} \oplus \boldsymbol{y} = \boldsymbol{x} + \boldsymbol{y}$), the
 semiassociative operation of matrix-vector multiplication
-($A \otimes x = Ax$) and the associative operation of
+($A \otimes \boldsymbol{x} = A\boldsymbol{x}$) and the associative operation of
 matrix-matrix multiplication ($A \odot B=AB$) satisfy Blelloch's three
@@
 \begin{itemize}
-\item{We classify RNNs which satisfy the conditions above, and
-    show that many RNNs used in practice such as Quasi-RNNs (QRNNs) \cite{bradbury2017quasi} are
+\item{We classify the RNNs which satisfy the conditions above, and
+    show that many of the RNNs used in practice such as the Quasi-RNN are
     contained in this class}
@@
     algorithm as a CUDA kernel, and show that it speeds up training of
-    QRNN and Simple Recurrent Unit (SRU) \cite{lei2017} architectures by factors of up to 9x}
+    the SRU and QRNN architectures by factors of up to 9x}
 \item{We describe how several recent linear RNNs can be described as
-    linear surrogates for non-linear architectures. We introduce a linear
+    linear surrogates for nonlinear architectures. We introduce a linear
     surrogate for the LSTM and show that we are able to train it
@@
 $C_\text{sscan} \in \mathcal{O}((C_\otimes + C_\oplus)T)$, compared to the parallel scan cost
-$C_\text{pscan} \in \mathcal{O}(2(C_\odot + C_\otimes + C_\oplus)(T/p + \lg p))$ \cite{blelloch1990prefix} on $p$ processors.
+$C_\text{pscan} \in \mathcal{O}(2(C_\odot + C_\otimes + C_\oplus)(T/p + \lg p))$ \cite{blelloch1990prefix},
+where \(p\) is the number of available parallel cores. 
 If $h_t$ is a vector of dimension $n$ then 
@@
 proportional to $n$ and $C_\text{pscan}\in \mathcal{O}(6n(T/p + \lg p))$ and
-$C_\text{sscan} \in \mathcal{O}(2nT)$. This gives a parallel speedup factor of
-$\frac{pT}{3(T+\lg p)}$. Assuming $p \ll T$, then $C_\text{pscan} \le
-C_\text{sscan}$ when $p \ge 3$.
-
+$C_\text{sscan} \in \mathcal{O}(2nT)$. Assuming $p \ll T$, then $C_\text{pscan} \le
+C_\text{sscan}$ when $p \ge 3$.\Chris{Can include the actual speedup
+  factor, \(\alpha = \frac{PT}{3(T + \lg(p))}\).}
 As we are only considering diagonal matrices,
@@
 network models use separate recurrent matrices for each sequence element and using these
-separate matrices would require potentially prohibitive $n^2T$ memory.
+separate matrices would require potentially prohibitive $n^2T$ memory.\Chris{I can't
+seem to find anyone explicitly saying this before from a quick google.}
 Applying
@@
 \subsection{Implementation}
-GPUs commonly used for deep learning in 2017 consist of between 640 and 3200 parallel
-processors known as warps. Each warp operates on 32 single precision floating
+In 2017, GPUs commonly used in deep learning consist of between 640 and 3200
+concurrently executing warps. Each warp operates on 32 single precision floating
 point numbers in parallel.
@@
 into the TensorFlow \cite{abadi2016tensorflow} framework. Each warp acts as a
-processor, which means the algorithmic $p$ is up to 3200 and the theoretical
+processor, which means the algorithmic parallelizability $p$ is up to 3200 and the theoretical
 parallelization speedup factor is up to several hundred.  The 32 lanes of each
@@
 \subsubsection{Impact on effective "batch size"}
-Consider evaluating an RNN with recurrence $h_t = \sigma(Uh_{t-1} + Vx_t + b)$
+Consider evaluating an RNN with recurrence $h_t = \sigma(Uh_{t-1} + Vx_t + b)$,
 from $m$ inputs to $n$ hidden units on a sequence of length $T$ with minibatch
@@
 recurrence forces even the better approach to perform $T$ potentially small $(b,
-m) \times (m, n)$ matrix multiplications in serial. This makes serial RNN
-performance heavily dependent on minibatch size.
+m) \times (m, n)$ matrix multiplications in serial which makes performance
+heavily dependent on minibatch size.
 
@@
 two large matrix multiplications and a parallel linear recurrence. GILR performance
-is much less dependent on batch size as the matrix multiplication kernel sees an "effective
+is much less dependent on batch size as the matrix multiplies see an "effective
 batch size" of $bT$ and $T$ is typically large.
@@
 prevents application of the parallel linear recurrence algorithm and forces slow
-serial evaluation. To work around this inefficiency, note that $s_t$ serves
-dual purposes. In $s_t = f(s_{t-1}, x_t)$, $s_{t-1}$ serves as an input to $f$
+serial evaluation. To work around this inefficiency, note that $s_t$ serves a
+dual purpose. In $s_t = f(s_{t-1}, x_t)$, $s_{t-1}$ serves as an input to $f$
 summarizing the previous inputs while $s_t$ serves as the output of $f$ to be
@@
 computable.  We refer to this class of model as a linear surrogate RNN
-(LS-RNN). QRNNs \cite{bradbury2017quasi} are LS-RNNs using $\tilde{h}_{t-1}
+(LS-RNN). Quasi-RNNs \cite{bradbury2017quasi} are LS-RNNs using $\tilde{h}_{t-1}
 = W_k x_{t-k} + ... W_1 x_{t-1}$ and strongly typed
@@
 
-Consider an LSTM:
-\begin{align*}
-  f_t, i_t, o_t &= \sigma(U_{f,i,o} h_{t-1} + V_{f,i,o} x_t + b_{f,i,o}) \\
-  z_t &= \tau(U_z h_{t-1} + V_z x_t + b_z) \\
-  c_t &= f_t \odot c_{t-1} + i_t \odot z_t \\
-  h_t &= o_t \odot c_t
-\end{align*}
-An LSTM has state $s_t = (h_t, c_t)$. Since $c_t$ depends only
+As an example LS-RNN, consider an LSTM:
+\begin{align*} f_t, i_t, o_t &= \sigma(U_{f,i,o} h_{t-1} + V_{f,i,o} x_t +
+b_{f,i,o}) \\ z_t &= \tau(U_z h_{t-1} + V_z x_t + b_z) \\ c_t &= f_t \odot
+c_{t-1} + i_t \odot z_t \\ h_t &= o_t \odot c_t
+\end{align*} An LSTM has state $s_t = (h_t, c_t)$. Since $c_t$ depends only
 linearly on $c_{t-1}$, no surrogate is needed for $c_t$. $h_t$ has a non-linear
@@
 the surrogate, we obtain the GILR-LSTM:
-\begin{align*}
-  g_t &= \sigma(V_g x_t + b_g) \\
-  j_t &= \tau(V_j x_t + b_j) \\
-  \tilde{h}_t &= g_t \odot \tilde{h}_{t-1} + (1-g_t)\odot j_t \\
-  f_t, i_t, o_t &= \sigma(U_{f,i,o} \tilde{h}_{t-1} + V_{f,i,o} x_t + b_{f,i,o}) \\
-  z_t &= \tau(U_z \tilde{h}_{t-1} + V_z x_t + b_z) \\
-  c_t &= f_t \odot c_{t-1} + i_t \odot z_t \\
-  h_t &= o_t \odot c_t \\
+\begin{align*} f_t, i_t, o_t &= \sigma(U_{f,i,o} \tilde{h}_{t-1} + V_{f,i,o} x_t
++ b_{f,i,o}) \\ z_t &= \tau(U_z \tilde{h}_{t-1} + V_z x_t + b_z) \\ c_t &= f_t
+\odot c_{t-1} + i_t \odot z_t \\ h_t &= o_t \odot c_t \\ g_t &= \sigma(V_g x_t +
+b_g) \\ \tilde{h}_t &= g_t \odot \tilde{h}_{t-1} + (1-g_t)\odot \tau(Wx_t + b_h)
 \end{align*}
 
-For $m$ inputs and hidden size $n$, a GILR-LSTM contains $2n(n+m)$ more
+For $m$ inputs and hidden size $n$, the GILR-LSTM contains $2n(n+m)$ more
 parameters than the equivalently sized LSTM to handle the mapping from $x$ to
@@
 parallel linear recurrence algorithm is able to achieve higher throughput than a
-similarly implemented serial version of the algorithm. We show that the parallel kernel
-is up to 40 times faster than the serial equivalent for long sequence lengths and
-that this speedup translates to implementations of LS-RNNs such as QRNN (where
-we can get a speedup of up to 9x).
+similarly implemented serial version of the algorithm. We show that the plr kernel
+is up to 40 times faster than the serial equivalent for long sequence lengths, and
+that this speedup translates to implementations of LS-RNNs such as the QRNN (where
+we can get a speedup of up to 9 times).
 
-In order to illustrate that the linearization does not necessarily come at the
-cost of expressibility, we show that the GILR-LSTM architecture computed with
-the parallel linear recurrence algorithm, is able to outperform the optimized CuDNN LSTM
-implementation
+In order to illustrate that the linearisation does not necessarily come at the
+cost of expressibility, we show that the GILR-LSTM architecture, trained with
+the PLR algorithm, is able to outperform the optimized CuDNN LSTM implementation
 on a pathological example from the original LSTM paper
@@
 
-\subsection{Throughput benchmarks}
-
-\subsubsection{Kernel performance}
+\subsection{Throughput Benchmarks}
 We first illustrate the throughput advantage of the parallel scan algorithm for
@@
 evaluates the parallel linear recurrence described in algorithm \ref{alg:plr},
-and one which evaluates the linear recurrence on GPU using the standard serial
-approach (with parallelization over features but not sequence length).
-This comparison is performed directly at the kernel level, avoiding
-any overhead from TensorFlow. We find that the parallel kernel has
-a distinct advantage at long sequence lengths with a speedup factor of up to 40x,
-as shown in table \ref{table:kernel-throughput}.
-At short sequence lengths, the parallel kernel does not perform well due to the
-overhead of multiple passes over data and communication between processors.
-\begin{table}
-\begin{center}
+and one which evaluates the linear recurrence using the straightforward serial
+approach. This comparison is performed directly at the kernel level, avoiding
+any overhead due to calling Tensorflow. We find that at long sequence lengths,
+the PLR kernel has a distinct advantage, as shown in table
+\ref{table:kernel-throughput}, with a speedup factor of 40. At short sequence
+lengths, the parallel kernel does not perform well due to the extra overhead
+involved in marshalling the parallel algorithm. \Chris{Important:
+  What's the batch size here? Is it just one sequence? Need to say that}
+\begin{table}[]
+  \centering
+\caption{Speedup of the PLR kernel compared to the SLR kernel, without any
+  overhead from Tensorflow, for various sequence lengths and input size \(m\).}      
+\label{table:kernel-throughput}
 \begin{tabular}{@{}llll@{}}
-    \label{table:kernel-throughput}
 Sequence Length & \(m=4\)  & \(m=32\) & \(m=128\) \\ \midrule
@@
 \end{tabular}
-\end{center}
-\caption{
-  Speedup of the parallel kernel compared to the serial kernel at various
-  values of $(\text{minibatch size } \mathbf{x} \text{ features}) = m$.
-  Performance was directly measured without any overhead from TensorFlow.
-}
 \end{table}
 
-\subsubsection{Accelerating existing RNN architectures}
+Many of the RNNs that have been recently introduced can be implemented with a
+parallel linear recurrence. Table \ref{table:rnn-throughput} shows the
+throughput advantage by using a parallel linear recurrence when compared to a
+serial linear recurrence. Since PLR and SLR are simply different methods for
+computing the recurrence, there is no penalty incurred for this speedup - we
+obtain exactly the same output.
 
-Many recently introduced RNNs can be accelerated with the parallel linear recurrence
-algorithm.
-Table \ref{table:rnn-throughput} shows the
-throughput advantage from using parallel linear recurrence compared to
-serial linear recurrence. Both methods compute an identical recurrence,
-so using parallel rather than serial causes no numerical changes.
-
-We controlled for GPU memory usage by choosing minibatch size such that the product
-of minibatch size and sequence length is constant. We see that the simpler architectures
-(for which the linear recurrence is a higher proportion of the
-computational burden) are more affected by the switch to the parallel kernel.  This is
+Typically experiments will aim to maximize the memory
+utilisation of the GPU, so we simulate this by choosing the product of the batch
+size and the sequence length to be constant. We see that the architectures which
+have fewer terms (for which the linear recurrence is a higher proportion of the
+computational burden) are more effected by the switch to PLR.  This is
 particularly clear in the case of the QRNN, where including a wider
-convolutional window results in more time spent outside of the linear recurrence and
-therefore less to gain from the parallelization.
-\begin{table}
+convolutional window results in less time spent on the recurrence and less to
+gain from the parallelization.
+\begin{figure}
 \begin{center}
@@
 \end{center}
-\caption{
-  Speedup obtained from using parallel linear recurrence rather than
-  serial linear recurrence. We analyse this for
-  SRUs \cite{lei2017} and QRNNs \cite{bradbury2017quasi} with a convolutional
-  filters of size 2 and 10, and GILR-LSTM which we introduce in section \ref{sec:ls-rnns}.
-  We controlled for GPU memory by setting minibatch size $b$ such that
-  $bT = 65536$ for sequence length $T$.
-}
-\end{table}
+\caption{Speedup obtained when using parallel linear recurrence compared to
+  serial linear recurrence. We analyse this for several recent architectures:
+  the Simple Recurrent Unit from \cite{lei2017}, the Quasi-RNN from
+  \cite{bradbury2017quasi}, with a convolutional filter of size 2 and 10, and
+  the GILR-LSTM which we introduce in section \ref{sec:ls-rnns}. To give a fair
+  model of a typical use-case maximising the memory usage of the GPU, we set
+  batch size \(b\) such that \(Tb = 65536\) for a sequence length \(T\).}
+\end{figure}
 
-\Eric{For table \ref{table:rnn-throughput} how many layers and what was hidden size?}
+Finally, we compare the throughput of the linear surrogate LSTM to the current
+fastest available implementation of the LSTM, the CuDNN-LSTM. Figure
+\ref{fig:tp_perf} shows that our GILR-LSTM has higher throughput for most of the
+low-batch-size regime, with a 30x speedup at a sequence length of 65,536. Recent
+work \cite{keskar2017large} has suggested that small mini-batches may aid
+generalization by encouraging the discovery of shallow minima. The GILR-LSTM
+allows us to achieve high throughput for all choices of minibatch size and
+sequence length.
 
-\subsubsection{GILR-LSTM throughput}
-
-\Eric{I think we should delete this section. Already over length limit, doesn't show
-anything great}
-
-Finally, we compare the throughput of the GILR-LSTM to the current fastest
-LSTM implementation (which comes from the CuDNN library). Figure
-\ref{fig:tp_perf} shows that the GILR-LSTM has higher throughput for most of the
-small minibatch regime, with a 30x speedup at a sequence length of 65,536.
-The GILR-LSTM allows us to achieve high throughput for all choices of minibatch size and
-sequence length.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% The computational performance of the LSTM and GILR-LSTM models were compared       %
+% across a wide range of minibatch sizes and sequence lengths. Although the 234    %
+% unit GILR-LSTM has a similar number of parameters to the 256 unit LSTM, we         %
+% measured the computational performance of a 256 unit GILR-LSTM to observe any      %
+% impact of the linear surrogate calculation. The serial evaluation of the LSTM    %
+% causes its throughput to be independent of the sequence length as a doubling of  %
+% sequence length causes a doubling of runtime. Notably, the GILR-LSTM can achieve a %
+% higher throughput by running on one 8192 event sequence at a time than an LSTM   %
+% running with 256 sequence minibatches. Similar speedups were found for           %
+% inference.                                                                       %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-\Eric{Chris: Put title on Figure \ref{fig:tp_perf}}
-
-\begin{figure}[t]
+\begin{table}[t]
 \centering
@@
 \label{fig:tp_perf}
-\end{figure}
+\end{table}
 
-\subsection{Synthetic Experiment}
+% % Experiments were performed on the N-MNIST \cite{orchard2015converting}
+% dataset. N-MNIST captures the MNIST digit dataset \cite{lecun1998mnist} by
+% panning % an event driven camera over each digit. Each example in N-MNIST is a
+% sequence of single pixel events (x, y, polarity, timestamp) where % x and y each
+% indicate a pixel position in [0, 33], polarity indicates whether the pixel was
+% switching on or off, and timestamp is the time of the event in
+% microseconds. Videos produced from the event data show that positive polarity
+% events are often located on the leading edge of the digit motion and negative
+% polarity events on the trailing edge. Between the 60,000 digits in the training
+% set, N-MNIST contains approximately 250 million pixel events with sequence
+% lengths ranging from 500 to 8000 and averaging 4000 events.
+
+% % We attempt to forecast 50 events ahead in N-MNIST using a two layer LSTM with
+% 256 units per layer and a two layer GILR-LSTM with 234 units per layer. The
+% GILR-LSTM layer size was selected so that it had slightly fewer parameters than
+% the LSTM. Each model transformed the incoming event position into a
+% 40-dimensional embedding vector which was then combined with the polarity to
+% produce a 41 dimensional input. Both models output a $2 \mathbf{x} 34^2$ matrix
+% containing two future event location probability distributions conditioned on
+% the polarity of the future event. The training algorithm only considers the
+% probability distribution of the true future polarity and uses the cross entropy
+% loss function. The Adam \cite{kingma2014adam} optimization algorithm and Glorot
+% \cite{glorot2010understanding} initialization scheme were used. Training on a
+% minibatch of size $b$ with sequence length $T$ consisted of uniformly sampling
+% $b$ N-MNIST sequences and then extracting a single random $T$ element
+% subsequence (and its 50 element ahead forecast) from the sequence. Sequences
+% less than $T$ elements were padded out to length $T$, and there was a "burn-in"
+% of 30 events at the start of each sequence where no predictions were made. An
+% epoch was defined as a pass over as many pixel events as there are in the full
+% training set.
+
+% % All experiments were performed using TensorFlow 1.0 \cite{abadi2016tensorflow}
+% on a single Nvidia K80 GPU running for up to 18 hours. The LSTM model was
+% computed using TensorFlow's dynamic\_rnn and BasicLSTMCell routines which are
+% slower than but algorithmically similar to the cuDNN LSTM implementation.
+
+% % \subsection{Training performance} % The speed of neural net training is not
+% solely determined by the training throughput but also by the frequency of
+% optimization steps. As an example, a batch method may process inputs at the same
+% rate as a minibatch method, but the minibatch method will generally converge
+% much faster on large datasets. On an infinite dataset, the batch method never
+% takes a single optimization step regardless of its throughput. With this example
+% in mind, it is clear that achieving fast training is a balancing act between
+% training data throughput and optimization step latency and that training
+% performance should be evaluated with training curves and not just throughput
+% numbers such as time per epoch.
+
+% % The LSTM and GILR-LSTM offer differ latency and throughput tradeoffs. LSTM
+% throughput depends only on minibatch size but LSTM latency depends on both
+% minibatch size and sequence length. Experiments were conducted training the LSTM
+% models with minibatch size 256 and sequence lengths 128, 256, 512, and
+% 1024. Figure \ref{fig:learning_curves} shows the smaller sequence lengths led to
+% faster initial learning but inferior final performance.
+
+% % The throughput and latency of the GILR-LSTM are both influenced by batch size
+% and sequence length. Several GILR-LSTMs were trained with batch size 16 and
+% sequence length 1024. This combination was chosen because of the nearly maximum
+% throughput, the low latency, and the ease of building a minibatch given the
+% distribution of sequence lengths in the data set. The 234 unit GILR-LSTM reaches
+% a better training loss in roughly 4.5 hours than any of the LSTMs could reach in
+% 18 hours. This experimental evidence indicates the GILR-LSTM is as powerful of a
+% model as the LSTM and can be trained in a fraction of the time.
+
+% \begin{figure}[t]
+% \centering
+% \includegraphics[width=12cm]{lc2.png}
+% \caption{The LSTM models had a direct relationship between sequence length and training curve: the shorter the sequence length, the faster the initial learning and
+% the higher the final loss. The faster initial learning is explained by the decreased
+% latency of optimization steps and the higher final loss is explained by the
+% inability to learn long dependencies. The GILR-LSTM has 234 units per hidden
+% layer to have the same number of parameters as the 256 unit LSTMs.}
+% \label{fig:learning_curves}
+% \end{figure}
+
+% \subsection{Test performance}
+% \begin{table}[h]
+%   \caption{N-MNIST test results}
+%   \label{test-results}
+%   \centering
+%   \begin{tabular}{lllll}
+%     \toprule
+%     & Train Cross Entropy     & Test Cross Entropy     & Top-5 Accuracy & Top-20 Accuracy \\
+%     \midrule
+%     GILR-LSTM 			&  \textbf{4.567} & 4.856           & 10.51\% & 36.70\%   \\
+%     LSTM, seqlen=128    & 4.664 & \textbf{4.846}  & \textbf{10.62\%} & \textbf{37.08\%}   \\
+%     LSTM, seqlen=256    & 4.620                & 4.906           & 10.11\% & 35.40\%   \\
+%     LSTM, seqlen=512    & 4.602                & 4.883           & 10.45\% & 36.32\%   \\
+%     LSTM, seqlen=1024   & 4.611                & 4.880           & 10.32\% & 35.92\%   \\
+%     \bottomrule
+%   \end{tabular}
+% \end{table}
+% Beyond cross entropy on train and test, we also evaluated top-$k$ accuracy for $k=5, 20$. A probability distribution is top-$k$ correct if it assigns the realized location one of the
+% $k$ largest probabilities. N-MNIST contains 1156 pixel positions, so top-5 and top-20 accuracy are equivalent to localizing the future to 0.43\% and 1.7\% of pixels.
+
+% No regularization was attempted and all of the models overfit, as indicated by the best test performance from the model with the worst train performance.
+Table \ref{test-results} contains the full test results.
+Our focus was the fast training of powerful models, and we leave regularizing parallel linear recurrences and LS-RNNs to future work.
+Although not tested, it is possible that the GILR-LSTM generalized better than the LSTMs trained on long sequences due to the much smaller minibatch size of GILR-LSTM leading to a wider minima.
+
+\subsection{Synthetic Example}
 In order to demonstrate how PLR may be used to speed up tasks that the LSTM is
-well-suited for, we tackle a pathological inference problem first introduced
-in Hochreiter and Schmidhuber's initial LSTM paper \cite{hochreiter1997long}.
-The input consists of sequences of length $n$ with each sequence element a randomly
-chosen one-hot vector in $p$ dimensional space.
-The first vector in each sequence is always the same, up to the sign of the component (i.e. it is $\pm
-\boldsymbol{p_0}$). The sequential model must read in an entire sequence and then output
-the sign of the first sequence element.
-This sequence classification problem requires remembering the first element over the length of the
-sequence and is a canonical example of the long-term dependency problem.
-Due to the large throughput of the GILR-LSTM, we would expect that it
+well-suited for, we tackle a pathological inference problem. Originally
+introduced as example 2b from Hochreiter and Schmidhtbuer's introduction of the
+LSTM, \cite{hochreiter1997long}, it involves storing a number for a very long
+sequence of time steps. We have an alphabet of size \(p\) with each character
+represented as a particular one-hot vector in this space. We input a sequence of
+\(n\) of these vectors, chosen at random (with replacement). The first vector in
+the sequence is always the same, up to the sign of the component (i.e. it is \(\pm
+\boldsymbol{p_0} \).
+
+The two signs on the first component separate the sequences into two sets. The
+whole sequence is fed into the LSTM and we aim to learn to classify the
+sequences. This requires remembering the first element over the length of the
+sequence. Due to the large throughput of the GILR-LSTM, we would expect that it
 would do better when the sequence length is large. In the original formulation
 of the problem (dealing in the regime with around one hundred timesteps)
-, $p$ is set equal to $n$. As this would make the size of the input data
-grow impracticaly large as \(\mathcal{O}(n^2)\) for long sequences, we fix \(p =
+, \(p\) is set equal to \(n\). Since this would make the size of the input data
+grow impractically large as \(\mathcal{O}(n^2)\) for long sequences, we fix \(p =
 128\) and vary \(n\).
 
-We generated sequences for $n$ equal to 1,024, 8,192, and
-1,048,576. For each of these we compared a two layer GILR-LSTM with 512 hidden
-units to a two layer LSTM with 512 hidden units per layer implemented by CuDNN.
+We generated three sets of data: for \(n\) equal to 1,024, 8,192, and
+1,048,576. For each of these we compared a two-layer GILR-LSTM with 512 hidden
+units to a canonical LSTM network with roughly the same number of parameters. We
+could imagine using a two-layer LSTM with 512 hidden units in each layer, or a
+one-layer LSTM with 1024 hidden units.
 
-We ran all experiments on a NVIDIA K80 GPU, choosing the largest
-minibatch size which fit into the GPU memory with five runs per configuration
+Initial experiments showed that the two-layer LSTM was much quicker to converge
+than the one-layer LSTM, so here we compare the GILR-LSTM to the two-layer
+CudnnLSTM.  We ran all experiments on a NVIDIA K80 GPU, choosing the largest
+minibatch size which fit into the GPU memory, with five runs per configuration
 allowing us to find the average and standard deviation of the time and number of
-iterations to convergence. A brief search over learning rate
+iterations to convergence. For all CUDA runs, a brief search over learning rate
 and batch size was carried out to find the parameters which allow the network to
-converge most rapidly for all runs. The criterion for convergence was five consecutive
-minibatches giving 100\% accuracy. The learning curves
-in figure \ref{fig:synthetic_training} give support to this being a reasonable
-convergence criteria.
-For the longest sequence length, we could not run long enough to observe the LSTM converge.
+converge most rapidly. The criterion for convergence was five consecutive
+minibatches giving 100\% accuracy. As can be seen from the learning curves
+in figure \ref{fig:synthetic_training}, this was a reasonable criterion. For the
+longest sequence-length, we didn't observe the CuDNN-LSTM converging.
 
-The results illustrate that the GILR-LSTM is able to converge much more quickly than
-the fastest available LSTM implementation.
-This is somewhat surprising given the LSTM was specifically constructed for problems
-of this sort. This GILR-LSTM convergence
-leads to the conclusion that the non-linearities present in
-LSTM are not necessary for solving this instance of the long-term dependency problem.
-The time to convergence further leads to the conclusion that inclusion of a non-linearity
-at every step incurs a significant training time slowdown.
-Furthermore, the GILR-LSTM is able to learn to carry dependencies over a one million
-element sequence. As far as we know, this one million step sequence experiment
-is the longest sequential learning problem to be handled by neural networks to date.
+The results illustrate that the GILR-LSTM is able to robustly outperform the
+CuDNN-LSTM, which is currently the fastest LSTM implementation available. This
+is somewhat surprising, as we might expect that the LSTM would do particularly
+well at an example that was specifically constructed for it. We can conclude that
+the nonlinearities in the LSTM's design are not vital for its performance given
+that the linear GILR-LSTM is able to converge, and that if these nonlinearities are
+not completely necessary, they entail a large performance cost by diallowing the
+plr algorithm from being used. Furthermore, the sequence length of over one million
+is the largest sequence length we have seen seriously discussed. The GILR
+handles it straightforwardly.
 
@@
   \begin{minipage}{0.6\textwidth}
-    \includegraphics[width=1.0\textwidth]{./1m_synthetic.png}
+    \includegraphics[width=1.0\textwidth]{./1M_synthetic.png}      
   \end{minipage}
@@
 
-\subsection{Abnormal heart sound detection}
-Medical waveform data consists of long sequences and is a reasonable application area for RNNs.
-We applied the GILR-LSTM
-on the problem presented in the 2016 PhysioNet/Computing in
+\subsection{The Medical Setting}
+A common application of RNNs is in medicine. In order to illustrate the
+increased performance that the plr algorithm allows, we implement an GILR-LSTM
+on the problem presented in , the 2016 PhysioNet/Computing in
 Cardiology Challenge 2016.  The aim of the Classification of Normal/Abnormal
-Heart Sound Recordings challenge is classify electroencephalogram (EEG) recordings
-as coming from a normal or abnormal heart.
-The waveforms are recorded on several sets of sensors
-and split into a training and testing set. The testing set is not
+Heart Sound Recordings challenge is to determine which of a set of EEGs are
+produced by people with abnormal hearts.  The data is recorded from a variety of
+sources and split into a training and testing set. The testing set is not
 publicly available, but can be evaluated through a web portal. The top results
-cover a large range of techniques, but generally rely on manual featurization of the
-raw EEG records. The GILR-LSTM's ability to handle long sequences makes it a
-candidate for training directly on the raw EEG signal.
-The EEGs are sampled at 2,000 Hz
-and range in length from several seconds to more than one hundred seconds,
+cover a large range of techniques, but generally rely on featurization of the
+raw EEG records. We want to see if we are able to leverage our advantage at high
+sequence-lenghts to train directly on the raw EEG. The EEGs are sampled at 2,000
+Hz and range in length from several seconds to more than one hundred seconds,
 giving a max sequence length of over 200,000. Since the testing server is not
@@
 normal hearts. We rebalance the data by oversampling the abnormal data by a factor
-of three, including each abnormal heart three times. We train a network feeding
-the one dimensional input signal into three stacked GILR-LSTM layers of 16 hidden units each.
-Although reasonably small, 16 hidden units allows developing
-develop representations. Parallel linear recurrence allows us to rapidly train the network
-over the entire long sequence length, so we don't have to work on truncated samples,
-downsampled sequenes, or
-spend lots of time manually developing EEG specific features.
-The model obtained perfect classification on the train dataset after
-a few hours \Eric{how many?} of training.
-On the unseen test data we obtained a score of PLACEHOLDER.
+of three, including each abnormal heart three times. We train a network consisting
+of three stacked GIRL-LSTM layers, each with 16 hidden units. Although reasonably
+small, since the input is only one-dimensional 16 hidden units allows us to
+develop representations. PLR allows us to train the network very rapidly over a
+long sequence length, so we don't have to downsample and lose information or
+spend lots of time developing methods to form features.
+We were able to fit exactly to the training set after
+a few hours of training, and on the unseen test data we obtained a score of
+PLACEHOLDER.
 
@@
 \section{Conclusion}
-Parallel linear recurrence can accelerate existing RNNs and be used in the design
-of new architectures such as the GILR-LSTM. Parallel linear recurrence enables
-rapid training with small minibatches, which further enables training models on
-extremely long sequence lengths.
-The impressive results in
+Parallel linear recurrence is an extremely powerful algorithm and the GILR-LSTM
+is just one of many possible models that can be built with it. While we were
+previously forced to use large minibatches in order to obtain reasonable
+training speed, with PLR in many cases we can train rapidly on small minibatch
+sizes, freeing us to utilise extremely long sequence lengths and more complex
+models. Our CUDA kernel is up to 40 times faster than a serial kernel on
+sequence lengths of 65,536, and this benefit transfers to implemented models,
+speeding up the SRU \cite{lei2017} by a factor of 9. The impressive results in
 \cite{lei2017} and other recent work comprehensively show that
-linear RNNs can often be used in place of traditional non-linear RNNs.
-Our results on the synthetic example continue this trend, showing that
+linearly-recurrent RNNs are, in many respects, equally as powerful as nonlinear
+LSTMs. Our results on the synthetic example continue this trend, showing that
 the GILR-LSTM is able to handle sequence length regimes that are longer than any

### 2017-10-27 paper/main.tex

--- 2017-10-27 paper/main.tex
@@
 of the problem (dealing in the regime with around one hundred timesteps)
-, $p$ is set equal to $n$. As this would make the size of the input data
-grow impracticaly large as \(\mathcal{O}(n^2)\) for long sequences, we fix \(p =
-128\) and vary \(n\).
+
+, \(p\) is set equal to \(n\). Since this would make the size of the
+input data grow impractically large as \(\mathcal{O}(n^2)\) for long
+sequences, we fix \(p = 128\) and vary \(n\).
 

### 2017-10-27 paper/main.tex
Cleaned up plots, added extra diagram

--- 2017-10-27 paper/main.tex
@@
 \usepackage{iclr2018_conference,times}
-\bibliographystyle{abbrvnat}
+\bibliographystyle{natbib}
 
@@
 sequential dependencies can be parallelized over the sequence length using the
-parallel scan algorithm, leading to rapid training on long sequences even with small
-minibatch size. We develop a parallel linear recurrence CUDA kernel and show that
-it can be applied to immediately speed up training and inference of
-several RNN architectures by up to 9x.
-We abstract recent work on linear RNNs into a new framework of linear surrogate RNNs
-and develop
-a linear surrogate model for the long short-term memory unit, the GILR-LSTM, that
-utilizies parallel linear recurrence.
-We push the limits of sequence learning by successfully training a GILR-LSTM
-on a synthetic sequence classification task with a one million timestep dependency.
-SOMETHING MEDICAL?
+parallel scan algorithm, leading to rapid training on long sequences even with
+small minibatch size. We develop a parallel linear recurrence CUDA kernel and
+show that it can be applied to immediately speed up training and inference of
+several state of the art RNN architectures by up to 9x.  We abstract recent work
+on linear RNNs into a new framework of linear surrogate RNNs and develop a
+linear surrogate model for the long short-term memory unit, the GILR-LSTM, that
+utilizies parallel linear recurrence.  We extend sequence learning to new
+extremely-long-timestep regimes that were previously out of reach by
+successfully training a GILR-LSTM on a synthetic sequence classification task
+with a one million timestep dependency.
 \end{abstract}
@@
 Recurrent neural networks (RNNs) are widely used for sequence modelling tasks in
-domains such as natural language processing \cite{sutskever2014sequence}, speech
-recognition \cite{amodei2015deep}, and reinforcement learning
-\cite{hausknecht2015deep}. Most RNNs, including popular variants such as long
-short-term memories (LSTMs) \cite{hochreiter1997long} and gated recurrent units
-(GRUs) \cite{cho2014learning}, contain a non-linear dependency between
-sequential inputs. These non-linear dependencies create a very flexible class of
-models but limit the feasibility of training RNNs on long sequences as each
-sequence element must be processed sequentially.  Modelling sequences of
+domains such as natural language processing \citep{sutskever2014sequence},
+speech recognition \citep{amodei2015deep}, and reinforcement learning
+\citep{hausknecht2015deep}. Most RNNs, including popular variants such as long
+short-term memories (LSTMs), introduced by \citet{hochreiter1997long}, and gated
+recurrent units (GRUs), introduced by \citet{cho2014learning}, contain a non-linear dependency
+between sequential inputs. These non-linear dependencies create a very flexible
+class of models but limit the feasibility of training RNNs on long sequences as
+each sequence element must be processed sequentially.  Modelling sequences of
 thousands to millions of elements is important to domains such as robotics,
@@
 pass and reduces the rate of optimization steps. Finally, training with larger
-minibatches damages generalization ability \cite{keskar2017large}. Given these
+minibatches damages generalization ability \citep{keskar2017large}. Given these
 effects, it is desirable to obtain high training throughput with small minibatches.
-Persistent RNNs \cite{diamos2016persistent} use a novel
+Persistent RNNs \citep{diamos2016persistent} use a novel
 implementation that can achieve high GPU utilization with very small minibatch
@@
 \begin{align*}
-\scan(\oplus, [a_1, a_2, ..., a_n], b) = [(a_1 \oplus b), (a_2 \oplus a_1 \oplus b), ..., (a_n \oplus a_{n-1} ... \oplus a_1 \oplus b)]
+\scan(\oplus, [a_1, a_2, ..., a_n], b) = [(a_1 \oplus b), (a_2 \oplus a_1 \oplus b), ..., (a_n \oplus a_{n-1} ... \oplus a_1 \oplus b)].
 \end{align*}
@@
 Despite their dependent computation graph, algorithms exist to parallelize scans
-and reductions when $\oplus$ is associative \cite{ladner1980parallel}.
+and reductions when $\oplus$ is associative \citep{ladner1980parallel}.
 
@@
 \begin{itemize}
-\item{We classify RNNs which satisfy the conditions above, and
-    show that many RNNs used in practice such as Quasi-RNNs (QRNNs) \cite{bradbury2017quasi} are
-    contained in this class}
-\item{We provide an implementation of the parallel linear recurrence
-    algorithm as a CUDA kernel, and show that it speeds up training of
-    QRNN and Simple Recurrent Unit (SRU) \cite{lei2017} architectures by factors of up to 9x}
+\item{We classify RNNs which satisfy the conditions above, and show that many
+    RNNs used in practice such as the Quasi-RNNs (QRNNs) introduced by
+    \citet{bradbury2017quasi} are contained in this class.}
+\item{We provide an implementation of the parallel linear recurrence algorithm
+    as a CUDA kernel, and show that it speeds up training of QRNN and
+    \citet{lei2017}'s Simple Recurrent Unit (SRU) architectures by factors of up
+    to 9x.}
 \item{We describe how several recent linear RNNs can be described as
@@
 $C_\text{sscan} \in \mathcal{O}((C_\otimes + C_\oplus)T)$, compared to the parallel scan cost
-$C_\text{pscan} \in \mathcal{O}(2(C_\odot + C_\otimes + C_\oplus)(T/p + \lg p))$ \cite{blelloch1990prefix} on $p$ processors.
+$C_\text{pscan} \in \mathcal{O}(2(C_\odot + C_\otimes + C_\oplus)(T/p + \lg p))$ on $p$ processors \citep{blelloch1990prefix}.
 If $h_t$ is a vector of dimension $n$ then
@@
 $C_\text{sscan} \in \mathcal{O}(2nT)$. This gives a parallel speedup factor of
-$\frac{pT}{3(T+\lg p)}$. Assuming $p \ll T$, then $C_\text{pscan} \le
+$pT/3(T+\lg p)$. Assuming $p \ll T$, then $C_\text{pscan} \le
 C_\text{sscan}$ when $p \ge 3$.
@@
 Limiting $\Lambda_t$ to be diagonal may seem like a severe constraint but there are
-several reasons to do so beyond the unfavorable parallelization performance. Relatively few neural
+several reasons to do so beyond the favorable parallelization performance. Relatively few neural
 network models use separate recurrent matrices for each sequence element and using these
@@
 This work implemented parallel linear recurrence as a CUDA kernel with bindings
-into the TensorFlow \cite{abadi2016tensorflow} framework. Each warp acts as a
+into the TensorFlow \citep{abadi2016tensorflow} framework. Each warp acts as a
 processor, which means the algorithmic $p$ is up to 3200 and the theoretical
@@
 computable.  We refer to this class of model as a linear surrogate RNN
-(LS-RNN). QRNNs \cite{bradbury2017quasi} are LS-RNNs using $\tilde{h}_{t-1}
+(LS-RNN). QRNNs \citep{bradbury2017quasi} are LS-RNNs using $\tilde{h}_{t-1}
 = W_k x_{t-k} + ... W_1 x_{t-1}$ and strongly typed
-RNNs\cite{balduzzi2016strongly} are LS-RNNs with $\tilde{h}_t=x_{t-1}$. Although
+RNNs \citep{balduzzi2016strongly} are LS-RNNs with $\tilde{h}_t=x_{t-1}$. Although
 not a rule, LS-RNNs can often be parallelized over sequence length with either
@@
 the surrogate, we obtain the GILR-LSTM:
+
 \begin{align*}
@@
 than an optimized LSTM implementation on a pathological long-term dependency
-problem on a pathological example from the original LSTM paper \cite{hochreiter1997long}.
+problem from the original LSTM paper \citep{hochreiter1997long}.
 Finally we show that the GILR-LSTM is able to
@@
 \subsubsection{Accelerating existing RNN architectures}
-Several recently introduced LS-RNNs can be accelerated with the parallel linear recurrence
-algorithm. We implemented SRUs, QRNNs (with filter width 2 and 10), and GILR-LSTMs
-that can be computed with either the standard
-serial linear recurrence algorithm or parallel linear recurrence. Both methods compute
-an identical recurrence, so switching from a serial to parallel implementation
-does not cause any numerical changes and takes only a single line of code changes.
-Notably, both SRUs and QRNNs
-claim an order of magnitude speedup compared to CuDNN LSTM when implemented with serial linear
-recurrence. Any further speedup from parallel linear recurrence applies on top of the existing
-speedup.
-We timed train throughput (forwards and backwards propagation), but the linear time
-of each pass also makes the results applicable to forwards (inference) performance. However,
-parallel linear recurrence can only accelerate inference in scenarios where the entire
-input sequence is known at the start of the inference phase.
-We controlled for GPU memory usage within these experiments by fixing $bT = 65536$ for
-minibatch size $b$ and sequence length $T$.
+Several recently introduced LS-RNNs can be accelerated with the parallel linear
+recurrence algorithm. We implemented SRUs, QRNNs (with filter width 2 and 10),
+and GILR-LSTMs that can be computed with either the standard serial linear
+recurrence algorithm or parallel linear recurrence. Both methods compute an
+identical recurrence, so switching from a serial to parallel implementation does
+not cause any numerical changes and takes only a single line of code changes.
+Notably, both SRUs and QRNNs claim an order of magnitude speedup compared to
+CuDNN LSTM when implemented with serial linear recurrence. Any further speedup
+from parallel linear recurrence applies on top of the existing speedup.  We
+timed train throughput (forwards and backwards propagation), but the linear time
+of each pass also makes the results applicable to forwards (inference)
+performance. However, parallel linear recurrence can only accelerate inference
+in scenarios where the entire input sequence is known at the start of the
+inference phase.  We controlled for GPU memory usage within these experiments by
+fixing $bT = 65,536$ for minibatch size $b$ and sequence length $T$, and chose a
+popular architecture consisting of two stacked RNN layers with 256 hidden units
+and an input size of 4.
 
-Table \ref{table:rnn-throughput} shows that
+Table \ref{table:rnn-throughput} shows that the
 throughput advantage from using parallel linear recurrence compared to
@@
 convolutional filters results in more time spent outside of the linear recurrence and
-therefore speedup from linear recurrence parallelization.
+therefore reduces the speedup from linear recurrence parallelization.
 \begin{table}[t]
-\caption{Parallel kernel speedup for a variety of LS-RNNs}
+  \caption{Parallel kernel speedup for a variety of LS-RNNs, implemented as two
+    stacked RNN layers with 256 hidden units. We keep the GPU memory usage constant
+  by fixing $bT = 65,536$ for minibatch size $b$ and sequence length $T$}
 \begin{center}
@@
 
-\Eric{For table \ref{table:rnn-throughput} how many layers and what was hidden size?}
+% \subsubsection{GILR-LSTM throughput}
 
-\subsubsection{GILR-LSTM throughput}
+% \Eric{I think we should delete this section. Already over length limit, doesn't show
+% anything great}
 
-\Eric{I think we should delete this section. Already over length limit, doesn't show
-anything great}
+% Eric: I think we should delete this. We're over length, results aren't great.
 
-Eric: I think we should delete this. We're over length, results aren't great.
+% Finally, we compare the throughput of the GILR-LSTM to the current fastest
+% LSTM implementation (which comes from the CuDNN library). Figure
+% \ref{fig:tp_perf} shows that the GILR-LSTM has higher throughput for most of the
+% small minibatch regime, with a 30x speedup at a sequence length of 65,536.
+% The GILR-LSTM allows us to achieve high throughput for all choices of minibatch size and
+% sequence length.
 
-Finally, we compare the throughput of the GILR-LSTM to the current fastest
-LSTM implementation (which comes from the CuDNN library). Figure
-\ref{fig:tp_perf} shows that the GILR-LSTM has higher throughput for most of the
-small minibatch regime, with a 30x speedup at a sequence length of 65,536.
-The GILR-LSTM allows us to achieve high throughput for all choices of minibatch size and
-sequence length.
+% Unlike the previous case, these speedups do come at a cost. The GILR-LSTM
+% architecture is different to the canonical LSTM, and it's possible that the loss
+% of model capacity outweighs the speedup in training. The following experiment
+% shows that the GILR-LSTM is able to outperform the LSTM on an example originally
+% designed to illustrate the advantages of the LSTM.
 
-Unlike the previous case, these speedups do come at a cost. The GILR-LSTM
-architecture is different to the canonical LSTM, and it's possible that the loss
-of model capacity outweighs the speedup in training. The following experiment
-shows that the GILR-LSTM is able to outperform the LSTM on an example originally
-designed to illustrate the advantages of the LSTM.
+% \Eric{Chris: Put title on Figure \ref{fig:tp_perf}}
 
-\Eric{Chris: Put title on Figure \ref{fig:tp_perf}}
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=12cm]{cudnn_heatmap.png}
-\caption{Throughput comparison between LSTM-256-256 with GILR-LSTM-256-256, with a
-  32-dimensional input and a 2-dimensional output, for various batch sizes and
-  sequence lengths.The LSTM only has a single row of data because its throughput
-  is independent of sequence length. Entries are missing from the GILR-LSTM table
-  because there was not enough memory on the GPU to handle such large batch
-  sizes and sequences.}
-\label{fig:tp_perf}
-\end{figure}
+% \begin{figure}[t]
+% \centering
+% \includegraphics[width=12cm]{cudnn_heatmap.png}
+% \caption{Throughput comparison between LSTM-256-256 with GILR-LSTM-256-256, with a
+%   32-dimensional input and a 2-dimensional output, for various batch sizes and
+%   sequence lengths.The LSTM only has a single row of data because its throughput
+%   is independent of sequence length. Entries are missing from the GILR-LSTM table
+%   because there was not enough memory on the GPU to handle such large batch
+%   sizes and sequences.}
+% \label{fig:tp_perf}
+% \end{figure}
 
 \subsection{Synthetic Experiment}
-In order to demonstrate how PLR may be used to speed up tasks that the LSTM is
-well-suited for, we tackle a pathological inference problem first introduced
-in Hochreiter and Schmidhuber's initial LSTM paper \cite{hochreiter1997long}.
-The input consists of sequences of length $n$ with each sequence element a randomly
-chosen one-hot vector in $p$ dimensional space.
-The first vector in each sequence is always the same, up to the sign of the component (i.e. it is $\pm
-\boldsymbol{p_0}$). The sequential model must read in an entire sequence and then output
-the sign of the first sequence element.
-This sequence classification problem requires remembering the first element over the length of the
-sequence and is a canonical example of the long-term dependency problem.
-Due to the large throughput of the GILR-LSTM, we would expect that it
-would do better when the sequence length is large. In the original formulation
-of the problem (dealing in the regime with around one hundred timesteps)
-, $p$ is set equal to $n$. As this would make the size of the input data
-grow impracticaly large as \(\mathcal{O}(n^2)\) for long sequences, we fix \(p =
-128\) and vary \(n\).
+One of the key strengths of the LSTM is that it is capable of dealing with
+long-term dependencies. In order to demonstrate that the GILR-LSTM is also able
+to handle long-term dependencies we tackle a canonical example of inference over
+many time steps from \citet{hochreiter1997long}.  We show that in fact the
+GILR-LSTM is able to outperform the CuDNN-LSTM and extend to sequence lengths
+orders of magnitude longer than dealt with previously. The input consists of
+sequences of length $n$ where for $n > 0$ each element is a randomly chosen
+one-hot vector \(x\) in $p$-dimensional space.  The first vector in each
+sequence, \(x_0\), is always either \((1, 0, \ldots, 0)\) or \((-1, 0, \ldots,
+0)\). The sequential model must read in an entire sequence and then output the
+sign of the first sequence element.  This sequence classification problem
+requires remembering the first element over the length of the sequence, and
+early RNNs struggled with this for \(p\) as small as a few dozen. In the
+original formulation of the problem (dealing in the regime with around one
+hundred timesteps), the dimensionality of the input \(p\) is set equal to
+\(n\). Since this would make the size of the input data grow impractically large
+as \(\mathcal{O}(n^2)\) for long sequences, we fix \(p = 128\) as we vary \(n\).
+\begin{figure}[] \centering
+\includegraphics[width=12cm]{synthetic_diagram.pdf}
+\caption{The structure of the synthetic example and the GILR-LSTM architecture
+we used to tackle it. We feed in one-hot unit vectors \(x\) which are chosen
+uniformly at random (with replacement). The class is determined by the very
+first vector \(x_0\), which has a fixed direction. The sign of \(x_0\)
+determines the class. In the diagram, each rounded block indicates a cell of the
+RNN, whilst the square indicates a linear unit.}
+\label{fig:synthetic_diagram}
+\end{figure} We generated sequences for $n$ equal to 1,024, 8,192, and
+1,048,576. For each of these we compared a two layer GILR-LSTM with 512 hidden
+units to a two layer LSTM with 512 hidden units\footnote{For the longest
+sequence length, the number of hidden units was decreased to 64 for both
+architectures so that the net could fit in memory.} per layer implemented by
+CuDNN.
 
-We generated sequences for $n$ equal to 1,024, 8,192, and
-1,048,576. For each of these we compared a two layer GILR-LSTM with 512 hidden
-units to a two layer LSTM with 512 hidden units per layer implemented by CuDNN.
-
-We ran all experiments on a NVIDIA K80 GPU, choosing the largest
-minibatch size which fit into the GPU memory with five runs per configuration
+We ran all experiments on a NVIDIA K80 GPU, with five runs per configuration
 allowing us to find the average and standard deviation of the time and number of
-iterations to convergence. A brief search over learning rate
+iterations to convergence. We continually generated random sequences to serve
+as input data.  A brief search over learning rate
 and batch size was carried out to find the parameters which allow the network to
@@
 convergence criteria.
-For the longest sequence length, we could not run long enough to observe the LSTM converge.
-
-The results illustrate that the GILR-LSTM is able to converge much more quickly than
-the fastest available LSTM implementation.
-This is somewhat surprising given the LSTM was specifically constructed for problems
-of this sort. This GILR-LSTM convergence
-leads to the conclusion that the non-linearities present in
-LSTM are not necessary for solving this instance of the long-term dependency problem.
-The time to convergence further leads to the conclusion that inclusion of a non-linearity
-at every step incurs a significant training time slowdown.
-Furthermore, the GILR-LSTM is able to learn to carry dependencies over a one million
-element sequence. As far as we know, this one million step sequence experiment
-is the longest sequential learning problem to be handled by neural networks to date.
-
-\begin{figure}
-  \centering
-  \begin{minipage}{0.5\textwidth}
-    \includegraphics[width=1.0\textwidth]{./1k_synthetic.png}
-  \end{minipage}%
-    \begin{minipage}{0.5\textwidth}
-    \includegraphics[width=1.0\textwidth]{./8k_synthetic.png}
-  \end{minipage}
-  \begin{minipage}{0.6\textwidth}
-    \includegraphics[width=1.0\textwidth]{./1m_synthetic.png}
-  \end{minipage}
-  \caption{Training curves for GILR-LSTM and CUDNNLSTM architectures for various
-  sequence lengths. (Clockwise from top-left: 1024, 8192, 1048576)}
-    \label{fig:synthetic_training}
-  \end{figure}
-
+For the longest sequence length, we did not observe the CuDDN converging, even after
+several days' training.
 \begin{table}[]
 \centering
-\caption{Performance of the GILR-LSTM compared to CUDA-optimised CudnnLSTM
-  implementation on problem 2b from (cite '97). We see that the GILR-LSTM has a clear
-  advantage in training speed as the sequence length increases.
-  We did not observe the Cudnn-LSTM converging for the longest task, even after several days.
-  {\scriptsize *For the longest sequence length, the number of hidden units was decreased to
-    64 for both architectures so that the net could fit in memory. }    }
+\caption{Performance of the GILR-LSTM compared to the CuDNN-LSTM
+  on problem 2b from \citet{hochreiter1997long}. We see that the GILR-LSTM has a
+  clear advantage in training speed as the sequence length increases.  We did
+  not observe the Cudnn-LSTM converging for the longest task, even after several
+  days.}
 \label{my-label}
@@
 \multicolumn{2}{c}{\textbf{1,024}} & \multicolumn{2}{c}{\textbf{8,192}} &
-\multicolumn{2}{c}{\textbf{1,048,576\(^*\)}} \\ \midrule &CuDNN& GILR &
+\multicolumn{2}{c}{\textbf{1,048,576}} \\ \midrule &CuDNN& GILR &
 CuDNN &GILR&CuDNN&GILR\\ \cmidrule(l){2-7} Iterations (1000s) &
@@
                                                                &-& 14 \(\pm\) 3 \\
-\begin{tabular}[c]{@{}l@{}}Wall Clock time (hours)\end{tabular} &
+\begin{tabular}[c]{@{}l@{}}Wall time (hours)\end{tabular} &
 0.28 \(\pm\) 0.08 & 0.031 \(\pm\) 0.002 & 0.58 \(\pm\) 0.06 & 0.10 \(\pm\) 0.03
@@
 
-\subsection{Abnormal heart sound detection}
-Medical waveform data consists of long sequences and is a reasonable application area for RNNs.
-We applied the GILR-LSTM
-on the problem presented in the 2016 PhysioNet/Computing in
-Cardiology Challenge 2016.  The aim of the Classification of Normal/Abnormal
-Heart Sound Recordings challenge is classify electroencephalogram (EEG) recordings
-as coming from a normal or abnormal heart.
-The waveforms are recorded on several sets of sensors
-and split into a training and testing set. The testing set is not
-publicly available, but can be evaluated through a web portal. The top results
-cover a large range of techniques, but generally rely on manual featurization of the
-raw EEG records. The GILR-LSTM's ability to handle long sequences makes it a
-candidate for training directly on the raw EEG signal.
-The EEGs are sampled at 2,000 Hz
-and range in length from several seconds to more than one hundred seconds,
-giving a max sequence length of over 200,000. Since the testing server is not
-equipped with a GPU, we cannot run PLR in the inference. We resample the
-records to 500 Hz so that we are able to evaluate the test results in a
-reasonable amount of time, padding all records that are shorter than the longest.
+The results illustrate that the GILR-LSTM is able to converge between 6 and 10
+times faster than the CuDNN-LSTM. This is somewhat surprising
+given the LSTM was specifically constructed for problems of this sort, and the
+CuDNN-LSTM is a highly-optimised implementation of the LSTM (to the extent that
+the monolithic interface it exposes is difficult to understand or extend). The
+GILR-LSTM is written almost entirely in native tensorflow, with only the call to
+the linear recurrence kernel activating an outside hook. This GILR-LSTM
+convergence leads to the conclusion that the non-linearities present in LSTM are
+not necessary for solving this instance of the long-term dependency problem.
+The time to convergence further leads to the conclusion that inclusion of a
+non-linearity at every step incurs a significant training time slowdown.
+Furthermore, the GILR-LSTM is able to learn to carry dependencies over a one
+million element sequence. As far as we know, this one million step sequence
+experiment is the longest sequential learning problem to be handled by neural
+networks to date.
 
-The training set of data is unbalanced, with 75\% of the data corresponding to
-normal hearts. We rebalance the data by oversampling the abnormal data by a factor
-of three, including each abnormal heart three times. We train a network feeding
-the one dimensional input signal into three stacked GILR-LSTM layers of 16 hidden units each.
-Although reasonably small, 16 hidden units allows developing
-develop representations. Parallel linear recurrence allows us to rapidly train the network
-over the entire long sequence length, so we don't have to work on truncated samples,
-downsampled sequenes, or
-spend lots of time manually developing EEG specific features.
-The model obtained perfect classification on the train dataset after
-a few hours \Eric{how many?} of training.
-On the unseen test data we obtained a score of PLACEHOLDER.
+\begin{figure}
+  \centering
+  \begin{minipage}{0.5\textwidth}
+    \includegraphics[width=1.0\textwidth]{./1k_synthetic_new.png}
+  \end{minipage}%
+    \begin{minipage}{0.5\textwidth}
+    \includegraphics[width=1.0\textwidth]{./8k_synthetic_new.png}
+  \end{minipage}
+  \begin{minipage}{0.5\textwidth}
+    \includegraphics[width=1.0\textwidth]{./1m_synthetic_new.png}
+  \end{minipage}
+  \caption{Training curves for GILR-LSTM and CuDNN-LSTM architectures for various
+    sequence lengths. Since we used
+    small batch sizes to train and evaluate accuracy, the accuracy fluctuated
+    (between 0 and 1 for the longest sequence). We show the mean and standard deviation
+    in a sliding window around each learning curve to illustrate this.}
+    \label{fig:synthetic_training}
+  \end{figure}
 
-\Chris{If we get good results (doesn't have to be best ever, but can be on par),
-  then keep}
-\Chris{If we don't get good results, or don't have results, we will need to kick
-  this section}
+% \subsection{Abnormal heart sound detection - will probably be cut :'(}
+% Medical waveform data consists of long sequences and is a reasonable application area for RNNs.
+% We applied the GILR-LSTM
+% on the problem presented in the 2016 PhysioNet/Computing in
+% Cardiology Challenge 2016.  The aim of the Classification of Normal/Abnormal
+% Heart Sound Recordings challenge is classify electroencephalogram (EEG) recordings
+% as coming from a normal or abnormal heart.
+% The waveforms are recorded on several sets of sensors
+% and split into a training and testing set. The testing set is not
+% publicly available, but can be evaluated through a web portal. The top results
+% cover a large range of techniques, but generally rely on manual featurization of the
+% raw EEG records. The GILR-LSTM's ability to handle long sequences makes it a
+% candidate for training directly on the raw EEG signal.
+% The EEGs are sampled at 2,000 Hz
+% and range in length from several seconds to more than one hundred seconds,
+% giving a max sequence length of over 200,000. Since the testing server is not
+% equipped with a GPU, we cannot run PLR in the inference. We resample the
+% records to 500 Hz so that we are able to evaluate the test results in a
+% reasonable amount of time, padding all records that are shorter than the longest.
+
+% The training set of data is unbalanced, with 75\% of the data corresponding to
+% normal hearts. We rebalance the data by oversampling the abnormal data by a factor
+% of three, including each abnormal heart three times. We train a network feeding
+% the one dimensional input signal into three stacked GILR-LSTM layers of 16 hidden units each.
+% Although reasonably small, 16 hidden units allows developing
+% develop representations. Parallel linear recurrence allows us to rapidly train the network
+% over the entire long sequence length, so we don't have to work on truncated samples,
+% downsampled sequenes, or
+% spend lots of time manually developing EEG specific features.
+% The model obtained perfect classification on the train dataset after
+% a few hours \Eric{how many?} of training.
+% On the unseen test data we obtained a score of PLACEHOLDER.
+
+% \Chris{If we get good results (doesn't have to be best ever, but can be on par),
+%   then keep}
+% \Chris{If we don't get good results, or don't have results, we will need to kick
+%   this section}
 
@@
 
-\bibitem[Lei and Zhang (2017)]{lei2017}
+\bibitem[Lei and Zhang(2017)]{lei2017}
 T.~Lei, Y.~Zhang,

### 2017-10-27 paper/main.tex
Hid final comment from me
Fixing stuff

@@
 linear surrogate model for the long short-term memory unit, the GILR-LSTM, that
-utilizies parallel linear recurrence.  We extend sequence learning to new
-extremely-long-timestep regimes that were previously out of reach by
+utilizes parallel linear recurrence.  We extend sequence learning to new
+extremely long sequence regimes that were previously out of reach by
 successfully training a GILR-LSTM on a synthetic sequence classification task
@@
 We perform several experiments. First we find that our parallel linear recurrence
-kernel is able to able to achieve up to 40x higher throughput than a serial implementation
+kernel is able to achieve up to 40x higher throughput than a serial implementation
 when applied to long sequences. Secondly, we confirm that this kernel speedup
@@
 problem from the original LSTM paper \citep{hochreiter1997long}.
-Finally we show that the GILR-LSTM is able to
-quickly achieve state of the art results on an existing medical data
-set. \Chris{I hope so\ldots}
 
@@
   days.}
-\label{my-label}
+\label{table:synth-table}
 \begin{tabular}{@{}lllllll@{}} \toprule Sequence Length &
@@
 
-The results illustrate that the GILR-LSTM is able to converge between 6 and 10
+The results as show in table \ref{synth-table} illustrate that the GILR-LSTM is able to converge between 6 and 10
 times faster than the CuDNN-LSTM. This is somewhat surprising
 given the LSTM was specifically constructed for problems of this sort, and the
-CuDNN-LSTM is a highly-optimised implementation of the LSTM (to the extent that
-the monolithic interface it exposes is difficult to understand or extend). The
-GILR-LSTM is written almost entirely in native tensorflow, with only the call to
-the linear recurrence kernel activating an outside hook. This GILR-LSTM
-convergence leads to the conclusion that the non-linearities present in LSTM are
+CuDNN LSTM implementation is highly optimized (to the extent that
+the monolithic interface it exposes is difficult to modify or extend). The
+GILR-LSTM is implemented entirely in standard TensorFlow with the exception of
+the call to new linear recurrence kernel in place of a TensorFlow symbolic
+loop.
+Convergence of the GILR-LSTM models
+leads to the conclusion that the non-linearities present in LSTM are
 not necessary for solving this instance of the long-term dependency problem.
@@
 \section{Conclusion}
-A significant portion of the success of deep learning can be attributed to massive
+A significant portion of the success of deep learning can be attributed to access to massive
 amounts of computation. Most of this computation is accessed through two
-building blocks with highly efficient parallel implementations:
+highly efficient and parallelizable building blocks:
 matrix multiplication and convolution. Recent research has demonstrated that

### 2017-10-27 paper/main.tex
Fixing stuff

@@
 convergence criteria.
-For the longest sequence length, we did not observe the CuDDN converging, even after
+For the longest sequence length, we did not observe the CuDNN LSTM converging, even after
 several days' training.
+
 \begin{table}[]
 \centering
-\caption{Performance of the GILR-LSTM compared to the CuDNN-LSTM
-  on problem 2b from \citet{hochreiter1997long}. We see that the GILR-LSTM has a
-  clear advantage in training speed as the sequence length increases.  We did
-  not observe the Cudnn-LSTM converging for the longest task, even after several
-  days.}
+\caption{Performance of the GILR-LSTM compared to the CuDNN LSTM
+  on problem 2b from \citet{hochreiter1997long}. }
 \label{table:synth-table}

### 2017-10-27 paper/main.tex
Final submission material

@@
     surrogate for the LSTM and show that we are able to train it
-    with a speedup of 5-10x compared to the CuDNN-LSTM when we use the
+    with a speedup of 5-10x compared to the CuDNN LSTM when we use the
     parallel linear recurrence algorithm.}
@@
 many time steps from \citet{hochreiter1997long}.  We show that in fact the
-GILR-LSTM is able to outperform the CuDNN-LSTM and extend to sequence lengths
+GILR-LSTM is able to outperform the CuDNN LSTM and extend to sequence lengths
 orders of magnitude longer than dealt with previously. The input consists of
@@
 \begin{table}[]
+\label{table:synth-table}
 \centering
@@
   on problem 2b from \citet{hochreiter1997long}. }
-\label{table:synth-table}
 \begin{tabular}{@{}lllllll@{}} \toprule Sequence Length &
@@
 
-The results as show in table \ref{synth-table} illustrate that the GILR-LSTM is able to converge between 6 and 10
-times faster than the CuDNN-LSTM. This is somewhat surprising
+The results as show in table \ref{table:synth-table} illustrate that the GILR-LSTM is able to converge between 6 and 10
+times faster than the CuDNN LSTM. This is somewhat surprising
 given the LSTM was specifically constructed for problems of this sort, and the
@@
 GILR-LSTM is implemented entirely in standard TensorFlow with the exception of
-the call to new linear recurrence kernel in place of a TensorFlow symbolic
-loop.
+using the new linear recurrence op instead of a TensorFlow symbolic loop.
 Convergence of the GILR-LSTM models
@@
   \end{minipage}
-  \caption{Training curves for GILR-LSTM and CuDNN-LSTM architectures for various
-    sequence lengths. Since we used
-    small batch sizes to train and evaluate accuracy, the accuracy fluctuated
-    (between 0 and 1 for the longest sequence). We show the mean and standard deviation
-    in a sliding window around each learning curve to illustrate this.}
+  \caption{Learning curves for GILR-LSTM and CuDNN LSTM architectures for various
+    sequence lengths. Each plot includes learning curves from multiple independent
+    runs. The shaded area around each learning curve shows the rolling variance
+    of the loss over minibatches within the run.
+}
     \label{fig:synthetic_training}

### 2018-01-02 paper/main.tex
Update paper to use new plots

@@
   \begin{minipage}{0.5\textwidth}
-    \includegraphics[width=1.0\textwidth]{./1k_synthetic_new.png}
+    \includegraphics[width=1.0\textwidth]{./between_chain_plots/1k_20_smoothing.png}
   \end{minipage}%
     \begin{minipage}{0.5\textwidth}
-    \includegraphics[width=1.0\textwidth]{./8k_synthetic_new.png}
+    \includegraphics[width=1.0\textwidth]{./between_chain_plots/8k_5_smoothing.png}
   \end{minipage}
   \begin{minipage}{0.5\textwidth}
-    \includegraphics[width=1.0\textwidth]{./1m_synthetic_new.png}
+    \includegraphics[width=1.0\textwidth]{./between_chain_plots/1m_20_smoothing.png}
   \end{minipage}
   \caption{Learning curves for GILR-LSTM and CuDNN LSTM architectures for various
-    sequence lengths. Each plot includes learning curves from multiple independent
-    runs. The shaded area around each learning curve shows the rolling variance
-    of the loss over minibatches within the run.
-}
+    sequence lengths. Each plot shows the moving mean and standard deviation of
+    classification accuracy over five training runs, with the exception of a single
+    run for CuDNN LSTM on 1 million sequence length.}
     \label{fig:synthetic_training}

### 2018-02-20 paper/main.tex
Preparing for camera ready

@@
 
-%\iclrfinalcopy % Uncomment for camera-ready version, but NOT for submission.
+\iclrfinalcopy % Uncomment for camera-ready version, but NOT for submission.
 
@@
 of at least several thousand steps on at least a 32 element vector.
+
+The parallel linear recurrence CUDA kernel and TensorFlow bindings are
+available at \url{https://github.com/eamartin/parallelizing_linear_rnns} .
 
@@
 \subsubsection*{Acknowledgments}
-Acknowledgements removed for anonymous submission.
-%We would like to acknowledge Kevin Bowers, Alex Meiburg, JD Co-Reyes, Carson
-%McNeil, Andy Palan, and several others for fruitful conversations and guidance.
+We would like to acknowledge Kevin Bowers, Alex Meiburg, JD Co-Reyes, Carson
+McNeil, Andy Palan, and several others for fruitful conversations and guidance.
 

### 2018-02-21 paper/main.tex
Added Chris contact details/acknowledgments

@@
 \And
-Chris Cundy
-Department of XX \\
-University of XX \\
-City, Country \\
+Chris Cundy \\
+Department of Computer Science\\
+University of California, Berkeley\\
+Berkeley, CA 94720, USA\thanks{Currently at the Future of Humanity Institute, University of Oxford, Oxford, UK} \\
 \texttt{c.cundy@berkeley.edu} \\
@@
 We would like to acknowledge Kevin Bowers, Alex Meiburg, JD Co-Reyes, Carson
-McNeil, Andy Palan, and several others for fruitful conversations and guidance.
+McNeil, Andy Palan, Sören Mindermann, and several others for fruitful conversations and guidance.
 

### 2018-04-21 poster/main.tex
add poster template

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Jacobs Landscape Poster
% LaTeX Template
% Version 1.1 (14/06/14)
%
% Created by:
% Computational Physics and Biophysics Group, Jacobs University
% https://teamwork.jacobs-university.de:8443/confluence/display/CoPandBiG/LaTeX+Poster
%
% Further modified by:
% Nathaniel Johnston (nathaniel@njohnston.ca)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[final]{beamer}

\usepackage[scale=1.24]{beamerposter} % Use the beamerposter package for laying out the poster

\usetheme{confposter} % Use the confposter theme supplied with this template

\setbeamercolor{block title}{fg=ngreen,bg=white} % Colors of the block titles
\setbeamercolor{block body}{fg=black,bg=white} % Colors of the body of blocks
\setbeamercolor{block alerted title}{fg=white,bg=dblue!70} % Colors of the highlighted block titles
\setbeamercolor{block alerted body}{fg=black,bg=dblue!10} % Colors of the body of highlighted blocks
% Many more colors are available for use in beamerthemeconfposter.sty

%-----------------------------------------------------------
% Define the column widths and overall poster size
% To set effective sepwid, onecolwid and twocolwid values, first choose how many columns you want and how much separation you want between columns
% In this template, the separation width chosen is 0.024 of the paper width and a 4-column layout
% onecolwid should therefore be (1-(# of columns+1)*sepwid)/# of columns e.g. (1-(4+1)*0.024)/4 = 0.22
% Set twocolwid to be (2*onecolwid)+sepwid = 0.464
% Set threecolwid to be (3*onecolwid)+2*sepwid = 0.708

\newlength{\sepwid}
\newlength{\onecolwid}
\newlength{\twocolwid}
\newlength{\threecolwid}
\setlength{\paperwidth}{48in} % A0 width: 46.8in
\setlength{\paperheight}{36in} % A0 height: 33.1in
\setlength{\sepwid}{0.024\paperwidth} % Separation width (white space) between columns
\setlength{\onecolwid}{0.22\paperwidth} % Width of one column
\setlength{\twocolwid}{0.464\paperwidth} % Width of two columns
\setlength{\threecolwid}{0.708\paperwidth} % Width of three columns
\setlength{\topmargin}{-0.5in} % Reduce the top margin size
%-----------------------------------------------------------

\usepackage{graphicx}  % Required for including images

\usepackage{booktabs} % Top and bottom rules for tables

%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\title{Parallelizing Linear Recurrent Neural Nets Over Sequence Length} % Poster title

\author{Eric Martin and Chris Cundy} % Author(s)

\institute{Fill in institutes} % Institution(s)

%----------------------------------------------------------------------------------------

\begin{document}

\addtobeamertemplate{block end}{}{\vspace*{2ex}} % White space under blocks
\addtobeamertemplate{block alerted end}{}{\vspace*{2ex}} % White space under highlighted (alert) blocks

\setlength{\belowcaptionskip}{2ex} % White space under figures
\setlength\belowdisplayshortskip{2ex} % White space under equations

\begin{frame}[t] % The whole poster is enclosed in one beamer frame

\begin{columns}[t] % The whole poster consists of three major columns, the second of which is split into two columns twice - the [t] option aligns each column's content to the top

\begin{column}{\sepwid}\end{column} % Empty spacer column

\begin{column}{\onecolwid} % The first column

%----------------------------------------------------------------------------------------
%	OBJECTIVES
%----------------------------------------------------------------------------------------

\begin{alertblock}{Objectives}

Lorem ipsum dolor sit amet, consectetur, nunc tellus pulvinar tortor, commodo eleifend risus arcu sed odio:
\begin{itemize}
\item Mollis dignissim, magna augue tincidunt dolor, interdum vestibulum urna
\item Sed aliquet luctus lectus, eget aliquet leo ullamcorper consequat. Vivamus eros sem, iaculis ut euismod non, sollicitudin vel orci.
\item Nascetur ridiculus mus.
\item Euismod non erat. Nam ultricies pellentesque nunc, ultrices volutpat nisl ultrices a.
\end{itemize}

\end{alertblock}

%----------------------------------------------------------------------------------------
%	INTRODUCTION
%----------------------------------------------------------------------------------------

\begin{block}{Introduction}

Lorem ipsum dolor \textbf{sit amet}, consectetur adipiscing elit. Sed commodo molestie porta. Sed ultrices scelerisque sapien ac commodo. Donec ut volutpat elit. Sed laoreet accumsan mattis. Integer sapien tellus, auctor ac blandit eget, sollicitudin vitae lorem. Praesent dictum tempor pulvinar. Suspendisse potenti. Sed tincidunt varius ipsum, et porta nulla suscipit et. Etiam congue bibendum felis, ac dictum augue cursus a. \textbf{Donec} magna eros, iaculis sit amet placerat quis, laoreet id est. In ut orci purus, interdum ornare nibh. Pellentesque pulvinar, nibh ac malesuada accumsan, urna nunc convallis tortor, ac vehicula nulla tellus eget nulla. Nullam lectus tortor, \textit{consequat tempor hendrerit} quis, vestibulum in diam. Maecenas sed diam augue.

This statement requires citation \cite{Smith:2012qr}.

\end{block}

%------------------------------------------------

\begin{figure}
\includegraphics[width=0.8\linewidth]{placeholder.jpg}
\caption{Figure caption}
\end{figure}

%----------------------------------------------------------------------------------------

\end{column} % End of the first column

\begin{column}{\sepwid}\end{column} % Empty spacer column

\begin{column}{\twocolwid} % Begin a column which is two columns wide (column 2)

\begin{columns}[t,totalwidth=\twocolwid] % Split up the two columns wide column

\begin{column}{\onecolwid}\vspace{-.6in} % The first column within column 2 (column 2.1)

%----------------------------------------------------------------------------------------
%	MATERIALS
%----------------------------------------------------------------------------------------

\begin{block}{Materials}

The following materials were required to complete the research:

\begin{itemize}
\item Curabitur pellentesque dignissim
\item Eu facilisis est tempus quis
\item Duis porta consequat lorem
\item Eu facilisis est tempus quis
\end{itemize}

The materials were prepared according to the steps outlined below:

\begin{enumerate}
\item Curabitur pellentesque dignissim
\item Eu facilisis est tempus quis
\item Duis porta consequat lorem
\item Curabitur pellentesque dignissim
\end{enumerate}

\end{block}

%----------------------------------------------------------------------------------------

\end{column} % End of column 2.1

\begin{column}{\onecolwid}\vspace{-.6in} % The second column within column 2 (column 2.2)

%----------------------------------------------------------------------------------------
%	METHODS
%----------------------------------------------------------------------------------------

\begin{block}{Methods}

Lorem ipsum dolor \textbf{sit amet}, consectetur adipiscing elit. Sed laoreet accumsan mattis. Integer sapien tellus, auctor ac blandit eget, sollicitudin vitae lorem. Praesent dictum tempor pulvinar. Suspendisse potenti. Sed tincidunt varius ipsum, et porta nulla suscipit et. Etiam congue bibendum felis, ac dictum augue cursus a. \textbf{Donec} magna eros, iaculis sit amet placerat quis, laoreet id est. In ut orci purus, interdum ornare nibh. Pellentesque pulvinar, nibh ac malesuada accumsan, urna nunc convallis tortor, ac vehicula nulla tellus eget nulla. Nullam lectus tortor, \textit{consequat tempor hendrerit} quis, vestibulum in diam. Maecenas sed diam augue.

\end{block}

%----------------------------------------------------------------------------------------

\end{column} % End of column 2.2

\end{columns} % End of the split of column 2 - any content after this will now take up 2 columns width

%----------------------------------------------------------------------------------------
%	IMPORTANT RESULT
%----------------------------------------------------------------------------------------

\begin{alertblock}{Important Result}

Lorem ipsum dolor \textbf{sit amet}, consectetur adipiscing elit. Sed commodo molestie porta. Sed ultrices scelerisque sapien ac commodo. Donec ut volutpat elit.

\end{alertblock}

%----------------------------------------------------------------------------------------

\begin{columns}[t,totalwidth=\twocolwid] % Split up the two columns wide column again

\begin{column}{\onecolwid} % The first column within column 2 (column 2.1)

%----------------------------------------------------------------------------------------
%	MATHEMATICAL SECTION
%----------------------------------------------------------------------------------------

\begin{block}{Mathematical Section}

Nam quis odio enim, in molestie libero. Vivamus cursus mi at nulla elementum sollicitudin. Nam quis odio enim, in molestie libero. Vivamus cursus mi at nulla elementum sollicitudin.

\begin{equation}
E = mc^{2}
\label{eqn:Einstein}
\end{equation}

Nam quis odio enim, in molestie libero. Vivamus cursus mi at nulla elementum sollicitudin. Nam quis odio enim, in molestie libero. Vivamus cursus mi at nulla elementum sollicitudin.

\begin{equation}
\cos^3 \theta =\frac{1}{4}\cos\theta+\frac{3}{4}\cos 3\theta
\label{eq:refname}
\end{equation}

Nam quis odio enim, in molestie libero. Vivamus cursus mi at nulla elementum sollicitudin. Nam quis odio enim, in molestie libero. Vivamus cursus mi at nulla elementum sollicitudin.

\begin{equation}
\kappa =\frac{\xi}{E_{\mathrm{max}}} %\mathbb{ZNR}
\end{equation}

\end{block}

%----------------------------------------------------------------------------------------

\end{column} % End of column 2.1

\begin{column}{\onecolwid} % The second column within column 2 (column 2.2)

%----------------------------------------------------------------------------------------
%	RESULTS
%----------------------------------------------------------------------------------------

\begin{block}{Results}

\begin{figure}
\includegraphics[width=0.8\linewidth]{placeholder.jpg}
\caption{Figure caption}
\end{figure}

Nunc tempus venenatis facilisis. Curabitur suscipit consequat eros non porttitor. Sed a massa dolor, id ornare enim:

\begin{table}
\vspace{2ex}
\begin{tabular}{l l l}
\toprule
\textbf{Treatments} & \textbf{Response 1} & \textbf{Response 2}\\
\midrule
Treatment 1 & 0.0003262 & 0.562 \\
Treatment 2 & 0.0015681 & 0.910 \\
Treatment 3 & 0.0009271 & 0.296 \\
\bottomrule
\end{tabular}
\caption{Table caption}
\end{table}

\end{block}

%----------------------------------------------------------------------------------------

\end{column} % End of column 2.2

\end{columns} % End of the split of column 2

\end{column} % End of the second column

\begin{column}{\sepwid}\end{column} % Empty spacer column

\begin{column}{\onecolwid} % The third column

%----------------------------------------------------------------------------------------
%	CONCLUSION
%----------------------------------------------------------------------------------------

\begin{block}{Conclusion}

Nunc tempus venenatis facilisis. \textbf{Curabitur suscipit} consequat eros non porttitor. Sed a massa dolor, id ornare enim. Fusce quis massa dictum tortor \textbf{tincidunt mattis}. Donec quam est, lobortis quis pretium at, laoreet scelerisque lacus. Nam quis odio enim, in molestie libero. Vivamus cursus mi at \textit{nulla elementum sollicitudin}.

\end{block}

%----------------------------------------------------------------------------------------
%	ADDITIONAL INFORMATION
%----------------------------------------------------------------------------------------

\begin{block}{Additional Information}

Maecenas ultricies feugiat velit non mattis. Fusce tempus arcu id ligula varius dictum.
\begin{itemize}
\item Curabitur pellentesque dignissim
\item Eu facilisis est tempus quis
\item Duis porta consequat lorem
\end{itemize}

\end{block}

%----------------------------------------------------------------------------------------
%	REFERENCES
%----------------------------------------------------------------------------------------

\begin{block}{References}

\nocite{*} % Insert publications even if they are not cited in the poster
\small{\bibliographystyle{unsrt}
\bibliography{sample}\vspace{0.75in}}

\end{block}

%----------------------------------------------------------------------------------------
%	ACKNOWLEDGEMENTS
%----------------------------------------------------------------------------------------

\setbeamercolor{block title}{fg=red,bg=white} % Change the block title color

\begin{block}{Acknowledgements}

\small{\rmfamily{Nam mollis tristique neque eu luctus. Suspendisse rutrum congue nisi sed convallis. Aenean id neque dolor. Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.}} \\

\end{block}

%----------------------------------------------------------------------------------------
%	CONTACT INFORMATION
%----------------------------------------------------------------------------------------

\setbeamercolor{block alerted title}{fg=black,bg=norange} % Change the alert block title colors
\setbeamercolor{block alerted body}{fg=black,bg=white} % Change the alert block body colors

\begin{alertblock}{Contact Information}

\begin{itemize}
\item Web: \href{http://www.university.edu/smithlab}{http://www.university.edu/smithlab}
\item Email: \href{mailto:john@smith.com}{john@smith.com}
\item Phone: +1 (000) 111 1111
\end{itemize}

\end{alertblock}

\begin{center}
\begin{tabular}{ccc}
\includegraphics[width=0.4\linewidth]{logo.png} & \hfill & \includegraphics[width=0.4\linewidth]{logo.png}
\end{tabular}
\end{center}

%----------------------------------------------------------------------------------------

\end{column} % End of the third column

\end{columns} % End of all the columns in the poster

\end{frame} % End of the enclosing frame

\end{document}

### 2018-04-22 poster/main.tex
Add cumsum example

@@
 
-\begin{alertblock}{Objectives}
-
-Lorem ipsum dolor sit amet, consectetur, nunc tellus pulvinar tortor, commodo eleifend risus arcu sed odio:
-\begin{itemize}
-\item Mollis dignissim, magna augue tincidunt dolor, interdum vestibulum urna
-\item Sed aliquet luctus lectus, eget aliquet leo ullamcorper consequat. Vivamus eros sem, iaculis ut euismod non, sollicitudin vel orci.
-\item Nascetur ridiculus mus.
-\item Euismod non erat. Nam ultricies pellentesque nunc, ultrices volutpat nisl ultrices a.
-\end{itemize}
-
+\begin{alertblock}{Abstract}
+RNN training and inference generally takes time linear in the sequence length because
+of non-linear sequential dependencies.
+We show the training and inference of RNNs with only linear
+sequential dependencies can be parallelized over the sequence length using the
+parallel scan algorithm, leading to rapid training on long sequences even with
+small minibatch size. We use this insight and a parallel linear recurrence CUDA
+kernel to accelerate several state of the art RNN architectures by up to 9x and
+to solve a synthetic sequence classification task with a one million timestep
+dependency.
 \end{alertblock}
@@
 
-\begin{block}{Introduction}
+\begin{block}{Background}
 
-Lorem ipsum dolor \textbf{sit amet}, consectetur adipiscing elit.
-Sed commodo molestie porta.
-Sed ultrices scelerisque sapien ac commodo.
-Donec ut volutpat elit.
-Sed laoreet accumsan mattis.
-Integer sapien tellus, auctor ac blandit eget, sollicitudin vitae lorem.
-Praesent dictum tempor pulvinar.
-Suspendisse potenti.
-Sed tincidunt varius ipsum, et porta nulla suscipit et.
-Etiam congue bibendum felis, ac dictum augue cursus a.
-\textbf{Donec} magna eros, iaculis sit amet placerat quis, laoreet id est.
-In ut orci purus, interdum ornare nibh.
-Pellentesque pulvinar, nibh ac malesuada accumsan, urna nunc convallis tortor, ac vehicula nulla tellus eget nulla.
-Nullam lectus tortor, \textit{consequat tempor hendrerit} quis, vestibulum in diam.
-Maecenas sed diam augue.
+Large minibatches are necessary for computational performance but create large memory
+requirements and may damage model generalization ability.
+
+Linear RNNs and convolutional models such as strongly typed RNNs, Wavenet, Bytenet,
+quasi-RNNs, and simple recurrent units
+have achieved state of the art results on many sequential tasks with rapid training times.
+
+Given $x_t$, $\lambda_t$ can compute $h_t=\lambda_t h_{t-1} + x_t$ for $t=1\ldots T$ on $p$
+processors in $O(T/p + \log(p))$ with the classic parallel scan algorithm.
 
@@
 \begin{figure}
-\includegraphics[width=0.8\linewidth]{placeholder.jpg}
-\caption{Figure caption}
+\includegraphics[width=1.0\linewidth]{cumsum.png}
+\caption{Parallel cumulative sum example}
 \end{figure}

### 2018-04-22 poster/main.tex
Add GILR-LSTM description

@@
 %----------------------------------------------------------------------------------------
-
 \begin{block}{Background}
-
 Large minibatches are necessary for computational performance but create large memory
@@
 processors in $O(T/p + \log(p))$ with the classic parallel scan algorithm.
+\end{block}
 
-This statement requires citation \cite{Smith:2012qr}.
+\begin{block}{Gated Impulse Linear Recurrence}
+We designed a recurrent layer that can take maximium advantage of parallel linear recurrence??
 
+\begin{align*}
+g_t &= \sigma(Ux_t + b_g) \\
+i_t &= \tau(Vx_t + b_z) \\
+h_t &= g_t \odot h_{t-1} + (1-g_t)\odot i_t
+\end{align*}
 \end{block}
@@
 
-\begin{block}{Materials}
+\begin{block}{Linear Surrogate RNNs}
 
-The following materials were required to complete the research:
+RNNs have a transition function $s_t = f(s_{t-1},x_t)$. $s_t$ serves dual roles as a
+summary of the past as well as the output of the unit. Non-linear $f$ in units such
+as vanilla RNN and LSTM prevents parallelization over sequence length.
 
-\begin{itemize}
-\item Curabitur pellentesque dignissim
-\item Eu facilisis est tempus quis
-\item Duis porta consequat lorem
-\item Eu facilisis est tempus quis
-\end{itemize}
+Replacing the summary of the past $s_{t-1}$ with a linear surrogate $\tilde{s}_{t-1}$
+allows the easy adaption of any existing RNN architecture for parallel computation.
 
-The materials were prepared according to the steps outlined below:
-
-\begin{enumerate}
-\item Curabitur pellentesque dignissim
-\item Eu facilisis est tempus quis
-\item Duis porta consequat lorem
-\item Curabitur pellentesque dignissim
-\end{enumerate}
+The state of an LSTM consists of $(c_t, h_t)$. $c_t$ is already computed
+by linear recurrence, so a linear surrogate LSTM must only compute a
+linear $\tilde{h}_t$. We let $h = \text{GILR}(x)$ and call it a GILR-LSTM.
 
@@
 
-\begin{block}{Methods}
+\begin{block}{Scan diagram and complexity??}
 
@@
 
-\begin{alertblock}{Important Result}
+\begin{alertblock}{Important Result (???)}
 

### 2018-04-22 poster/main.tex
Added first pass at poster results with modified graph

@@
 
-\begin{block}{Mathematical Section}
+\begin{block}{Beating CuDNN: Synthetic Data}
 
-Nam quis odio enim, in molestie libero. Vivamus cursus mi at nulla elementum sollicitudin. Nam quis odio enim, in molestie libero. Vivamus cursus mi at nulla elementum sollicitudin.
+  We tested the GILR-LSTM by comparing to the CuDNN LSTM implementation
+  on a synthetic memorisation task (problem 2b from %\citet{hochreiter1997long}. )
+  We compared the performance on variants with different values of
+  \(n\).  We obtained speedups of over 6x measured in wall clock time to
+  convergence.  Further, the GILR-LSTM attained convergence
+  when the time dependence of the problem had a length of \textbf{one  million time steps}.
 
-\begin{equation}
-E = mc^{2}
-\label{eqn:Einstein}
-\end{equation}
-
-Nam quis odio enim, in molestie libero. Vivamus cursus mi at nulla elementum sollicitudin. Nam quis odio enim, in molestie libero. Vivamus cursus mi at nulla elementum sollicitudin.
-
-\begin{equation}
-\cos^3 \theta =\frac{1}{4}\cos\theta+\frac{3}{4}\cos 3\theta
-\label{eq:refname}
-\end{equation}
-
-Nam quis odio enim, in molestie libero. Vivamus cursus mi at nulla elementum sollicitudin. Nam quis odio enim, in molestie libero. Vivamus cursus mi at nulla elementum sollicitudin.
-
-\begin{equation}
-\kappa =\frac{\xi}{E_{\mathrm{max}}} %\mathbb{ZNR}
-\end{equation}
+ \vspace{2em}
+\begin{figure}
+\includegraphics[width=1.0\linewidth]{8k_for_poster.png}
+\caption{Accuracy on the memorisation task with 8,192 sequence length}
+\end{figure}
 
@@
 
-\begin{block}{Results}
+\begin{block}{Accelerating existing linear RNNs}
+  We implemented several recent linear RNNs with the parallel linear
+  recurrence kernel, as well as a variant with purely serial evaluation.
+  We were able to obtain significant speedups with parallel evaluation
+  in all the architectures at longer sequence lengths.
 
-\begin{figure}
-\includegraphics[width=0.8\linewidth]{placeholder.jpg}
-\caption{Figure caption}
-\end{figure}
+  Since the parallel kernel computes the exact same result as the
+  serial kernel, we obtain this speedup with no tradeoff in terms of
+  reduced accuracy or precision.
 
-Nunc tempus venenatis facilisis. Curabitur suscipit consequat eros non porttitor. Sed a massa dolor, id ornare enim:
+  [Can be cut if no space]
+  We used two stacked RNN layers with 256 hidden
+  units, keeping the GPU memory usage constant by fixing $bT = 65,536$
+  for minibatch size $b$ and sequence length $T$. QRNN(2) refers to a
+  QRNN with filter size 2. %\citep{bradbury2017quasi}
 
-\begin{table}
+\begin{table}[t]
   \vspace{2ex}
-\begin{tabular}{l l l}
-\toprule
-\textbf{Treatments} & \textbf{Response 1} & \textbf{Response 2}\\
-\midrule
-Treatment 1 & 0.0003262 & 0.562 \\
-Treatment 2 & 0.0015681 & 0.910 \\
-Treatment 3 & 0.0009271 & 0.296 \\
-\bottomrule
+  \caption{Parallel kernel speedup for a variety of LS-RNNs.}
+\begin{center}
+\begin{tabular}{lrrrr}
+\label{table:rnn-throughput}
+\small{Seq.\ Len.} & \small{SRU} & \small{QRNN(2)}
+  & \small{QRNN(10)} & \small{GILR-LSTM}\\ \midrule
+16 & 0.28 & 0.38 & 0.78 & 0.61\\
+256 & 0.84 & 0.86 & 0.99 & 0.91\\
+4,096 & 1.38 & 1.18 & 1.05 & 0.98\\
+65,536 & 9.21 & 6.68 & 2.05 & 1.41\\ \bottomrule
 \end{tabular}
-\caption{Table caption}
+\end{center}
 \end{table}
+\end{block}
 
-\end{block}
+% \begin{table}
+% \vspace{2ex}
+% \begin{tabular}{l l l}
+% \toprule
+% \textbf{Treatments} & \textbf{Response 1} & \textbf{Response 2}\\
+% \midrule
+% Treatment 1 & 0.0003262 & 0.562 \\
+% Treatment 2 & 0.0015681 & 0.910 \\
+% Treatment 3 & 0.0009271 & 0.296 \\
+% \bottomrule
+% \end{tabular}
+% \caption{Table caption}
+% \end{table}
+
+% \end{block}
 

### 2018-04-24 poster/main.tex
Little rearrangement

@@
 
+%------------------------------------------------
+
+\begin{figure}
+\includegraphics[width=1.0\linewidth]{cumsum.png}
+\caption{Parallel cumulative sum example}
+\end{figure}
+
+%------------------------------------------------
+
 \begin{block}{Gated Impulse Linear Recurrence}
@@
 \end{block}
-
-%------------------------------------------------
-
-\begin{figure}
-\includegraphics[width=1.0\linewidth]{cumsum.png}
-\caption{Parallel cumulative sum example}
-\end{figure}
 
@@
 
-\begin{block}{Beating CuDNN: Synthetic Data}
-
-  We tested the GILR-LSTM by comparing to the CuDNN LSTM implementation
-  on a synthetic memorisation task (problem 2b from %\citet{hochreiter1997long}. )
-  We compared the performance on variants with different values of
-  \(n\).  We obtained speedups of over 6x measured in wall clock time to
-  convergence.  Further, the GILR-LSTM attained convergence
-  when the time dependence of the problem had a length of \textbf{one  million time steps}.
-
- \vspace{2em}
-\begin{figure}
-\includegraphics[width=1.0\linewidth]{8k_for_poster.png}
-\caption{Accuracy on the memorisation task with 8,192 sequence length}
-\end{figure}
-
-\end{block}
-
-%----------------------------------------------------------------------------------------
-
-\end{column} % End of column 2.1
-
-\begin{column}{\onecolwid} % The second column within column 2 (column 2.2)
-
-%----------------------------------------------------------------------------------------
-%	RESULTS
-%----------------------------------------------------------------------------------------
-
-\begin{block}{Accelerating existing linear RNNs}
+\begin{block}{Accelerating Existing Models}
   We implemented several recent linear RNNs with the parallel linear
@@
 \end{table}
+\end{block}
+
+%----------------------------------------------------------------------------------------
+
+\end{column} % End of column 2.1
+
+\begin{column}{\onecolwid} % The second column within column 2 (column 2.2)
+
+%----------------------------------------------------------------------------------------
+%	RESULTS
+%----------------------------------------------------------------------------------------
+
+\begin{block}{Synthetic Long-Term Dependency}
+
+  We tested the GILR-LSTM by comparing to the CuDNN LSTM implementation
+  on a synthetic memorisation task (problem 2b from %\citet{hochreiter1997long}. )
+  We compared the performance on variants with different values of
+  \(n\).  We obtained speedups of over 6x measured in wall clock time to
+  convergence.  Further, the GILR-LSTM attained convergence
+  when the time dependence of the problem had a length of \textbf{one  million time steps}.
+
+ \vspace{2em}
+\begin{figure}
+\includegraphics[width=1.0\linewidth]{8k_for_poster.png}
+\caption{Accuracy on the memorisation task with 8,192 sequence length}
+\end{figure}
+
 \end{block}

### 2018-04-24 poster/main.tex
Switch to 3 column layout

@@
 \setlength{\sepwid}{0.024\paperwidth} % Separation width (white space) between columns
-\setlength{\onecolwid}{0.22\paperwidth} % Width of one column
-\setlength{\twocolwid}{0.464\paperwidth} % Width of two columns
+\setlength{\onecolwid}{0.30\paperwidth} % Width of one column
+\setlength{\twocolwid}{0.624\paperwidth} % Width of two columns
 \setlength{\threecolwid}{0.708\paperwidth} % Width of three columns
@@
 Linear RNNs and convolutional models such as strongly typed RNNs, Wavenet, Bytenet,
-quasi-RNNs, and simple recurrent units
+Quasi-RNNs, and simple recurrent units
 have achieved state of the art results on many sequential tasks with rapid training times.
@@
 
-\begin{column}{\twocolwid} % Begin a column which is two columns wide (column 2)
-
-\begin{columns}[t,totalwidth=\twocolwid] % Split up the two columns wide column
-
-\begin{column}{\onecolwid}\vspace{-.6in} % The first column within column 2 (column 2.1)
+\begin{column}{\onecolwid} % Begin a column which is two columns wide (column 2)
 
@@
 
-\end{column} % End of column 2.1
-
-\begin{column}{\onecolwid}\vspace{-.6in} % The second column within column 2 (column 2.2)
-
-%----------------------------------------------------------------------------------------
-%	METHODS
-%----------------------------------------------------------------------------------------
-
-\begin{block}{Scan diagram and complexity??}
-
-Lorem ipsum dolor \textbf{sit amet}, consectetur adipiscing elit.
-Sed laoreet accumsan mattis.
-Integer sapien tellus, auctor ac blandit eget, sollicitudin vitae lorem.
-Praesent dictum tempor pulvinar.
-Suspendisse potenti.
-Sed tincidunt varius ipsum, et porta nulla suscipit et.
-Etiam congue bibendum felis, ac dictum augue cursus a.
-\textbf{Donec} magna eros, iaculis sit amet placerat quis, laoreet id est.
-In ut orci purus, interdum ornare nibh.
-Pellentesque pulvinar, nibh ac malesuada accumsan, urna nunc convallis tortor, ac vehicula nulla tellus eget nulla.
-Nullam lectus tortor, \textit{consequat tempor hendrerit} quis, vestibulum in diam.
-Maecenas sed diam augue.
-
-\end{block}
-
-%----------------------------------------------------------------------------------------
-
-\end{column} % End of column 2.2
-
-\end{columns} % End of the split of column 2 - any content after this will now take up 2 columns width
-
-%----------------------------------------------------------------------------------------
-%	IMPORTANT RESULT
-%----------------------------------------------------------------------------------------
-
-\begin{alertblock}{Important Result (???)}
-
-Lorem ipsum dolor \textbf{sit amet}, consectetur adipiscing elit. Sed commodo molestie porta. Sed ultrices scelerisque sapien ac commodo. Donec ut volutpat elit.
-
-\end{alertblock}
-
-%----------------------------------------------------------------------------------------
-
-\begin{columns}[t,totalwidth=\twocolwid] % Split up the two columns wide column again
-
-\begin{column}{\onecolwid} % The first column within column 2 (column 2.1)
-
 %----------------------------------------------------------------------------------------
@@
 \begin{table}[t]
-  \vspace{2ex}
-  \caption{Parallel kernel speedup for a variety of LS-RNNs.}
 \begin{center}
@@
 \end{center}
+\vspace{2ex}
+\caption{Parallel kernel speedup for a variety of LS-RNNs.}
 \end{table}
@@
 %----------------------------------------------------------------------------------------
-
-\end{column} % End of column 2.1
-
-\begin{column}{\onecolwid} % The second column within column 2 (column 2.2)
 
@@
 
+\end{column} % End of column 2.1
+
 % \begin{table}
@@
 %----------------------------------------------------------------------------------------
-
-\end{column} % End of column 2.2
-
-\end{columns} % End of the split of column 2
-
-\end{column} % End of the second column
 

### 2018-04-24 poster/main.tex
add heatmap

@@
 kernel to accelerate several state of the art RNN architectures by up to 9x and
-to solve a synthetic sequence classification task with a one million timestep
+to solve a synthetic sequence classification task with a one million time step
 dependency.
@@
 Large minibatches are necessary for computational performance but create large memory
-requirements and may damage model generalization ability.
+requirements and damage model generalization ability.
+
+\vspace{1ex}
 
@@
 
+\vspace{1ex}
+
 Given $x_t$, $\lambda_t$ can compute $h_t=\lambda_t h_{t-1} + x_t$ for $t=1\ldots T$ on $p$
-processors in $O(T/p + \log(p))$ with the classic parallel scan algorithm.
+processors in $O(T/p + \log(p))$ with the classic parallel scan algorithm. Backpropagation
+of gradient can also be parallelized with the same algorithm.
 \end{block}
+
+\vspace{2ex}
 
@@
 \includegraphics[width=1.0\linewidth]{cumsum.png}
-\caption{Parallel cumulative sum example}
+\caption{Cumulative sum parallelized over 3 processors}
 \end{figure}
@@
 \begin{block}{Gated Impulse Linear Recurrence}
-We designed a recurrent layer that can take maximium advantage of parallel linear recurrence??
+Given a fast algorithm for evaluating linear recurrences, we introduce a new
+linear recurrent layer called \textbf{gated impulse linear recurrence (GILR)}
 
@@
 as vanilla RNN and LSTM prevents parallelization over sequence length.
+\vspace{1ex}
 
@@
 allows the easy adaption of any existing RNN architecture for parallel computation.
+Several existing RNNs can be viewed as linear surrogate RNNs.
+\vspace{1ex}
 
@@
 by linear recurrence, so a linear surrogate LSTM must only compute a
-linear $\tilde{h}_t$. We let $h = \text{GILR}(x)$ and call it a GILR-LSTM.
-
+linear $\tilde{h}_t$. A \textbf{GILR-LSTM} uses $\tilde{h} = \text{GILR}(x)$
 \end{block}
@@
 
-\begin{block}{Accelerating Existing Models}
-  We implemented several recent linear RNNs with the parallel linear
-  recurrence kernel, as well as a variant with purely serial evaluation.
-  We were able to obtain significant speedups with parallel evaluation
-  in all the architectures at longer sequence lengths.
+\begin{block}{Accelerating Existing Linear Surrogate RNNs}
+We compare the performance of several recent linear RNNs
+implemented with the parallel linear recurrence kernel against
+a baseline serial linear recurrence kernel.
 
-  Since the parallel kernel computes the exact same result as the
-  serial kernel, we obtain this speedup with no tradeoff in terms of
-  reduced accuracy or precision.
-
-  [Can be cut if no space]
-  We used two stacked RNN layers with 256 hidden
-  units, keeping the GPU memory usage constant by fixing $bT = 65,536$
-  for minibatch size $b$ and sequence length $T$. QRNN(2) refers to a
-  QRNN with filter size 2. %\citep{bradbury2017quasi}
-
+\vspace{1ex}
 \begin{table}[t]
@@
 \vspace{2ex}
-\caption{Parallel kernel speedup for a variety of LS-RNNs.}
+\caption{Parallel kernel speedup for a variety of LS-RNNs.   All models use two stacked RNN layers with 256 hidden
+  units, keeping the GPU memory usage constant by fixing $bT = 65,536$
+  for minibatch size $b$ and sequence length $T$. QRNN($k$) refers to a
+  QRNN with filter size $k$. %\citep{bradbury2017quasi}
+}
 \end{table}
@@
 %----------------------------------------------------------------------------------------
-
-%----------------------------------------------------------------------------------------
-%	RESULTS
-%----------------------------------------------------------------------------------------
-
-\begin{block}{Synthetic Long-Term Dependency}
-
-  We tested the GILR-LSTM by comparing to the CuDNN LSTM implementation
-  on a synthetic memorisation task (problem 2b from %\citet{hochreiter1997long}. )
-  We compared the performance on variants with different values of
-  \(n\).  We obtained speedups of over 6x measured in wall clock time to
-  convergence.  Further, the GILR-LSTM attained convergence
-  when the time dependence of the problem had a length of \textbf{one  million time steps}.
-
- \vspace{2em}
+\vspace{-1ex}
 \begin{figure}
-\includegraphics[width=1.0\linewidth]{8k_for_poster.png}
-\caption{Accuracy on the memorisation task with 8,192 sequence length}
+\includegraphics[width=0.9\linewidth]{cudnn_heatmap_gilr.png}
+\caption{Throughput (thousand steps/s) for same sized LSTM and GILR-LSTM}
 \end{figure}
-
-\end{block}
 
@@
 \begin{column}{\onecolwid} % The third column
+
+\begin{block}{Synthetic Long-Term Dependency}
+
+  We tested the GILR-LSTM by comparing to the CuDNN LSTM implementation
+  on a synthetic memorisation task (problem 2b from %\citet{hochreiter1997long}. )
+  We compared the performance on variants with different values of
+  \(n\).  We obtained speedups of over 6x measured in wall clock time to
+  convergence.  Further, the GILR-LSTM attained convergence
+  when the time dependence of the problem had a length of \textbf{one  million time steps}.
+
+ \vspace{2em}
+\begin{figure}
+\includegraphics[width=1.0\linewidth]{8k_for_poster.png}
+\caption{Accuracy on the memorisation task with 8,192 sequence length}
+\end{figure}
+
+\end{block}
 

### 2018-04-24 poster/main.tex
fit middle column better

@@
 allows the easy adaption of any existing RNN architecture for parallel computation.
-Several existing RNNs can be viewed as linear surrogate RNNs.
+Several recent linear RNNs can be viewed as linear surrogate RNNs.
 \vspace{1ex}
@@
 
-\begin{block}{Accelerating Existing Linear Surrogate RNNs}
-We compare the performance of several recent linear RNNs
-implemented with the parallel linear recurrence kernel against
-a baseline serial linear recurrence kernel.
-
-\vspace{1ex}
+\begin{block}{Training Runtime Results}
 \begin{table}[t]
@@
 \vspace{2ex}
-\caption{Parallel kernel speedup for a variety of LS-RNNs.   All models use two stacked RNN layers with 256 hidden
+\caption{Parallel kernel speedup compared to serial linear recurrence for a variety of LS-RNNs
+  All models use two stacked RNN layers with 256 hidden
   units, keeping the GPU memory usage constant by fixing $bT = 65,536$
@@
 \includegraphics[width=0.9\linewidth]{cudnn_heatmap_gilr.png}
-\caption{Throughput (thousand steps/s) for same sized LSTM and GILR-LSTM}
+\caption{Throughput (thousand steps/s) for 2 layer 256 unit cuDNN LSTM and GILR-LSTM as
+a function of batch size and sequence length. LSTM throughput is independent of sequence
+length. GILR-LSTM can achieve much greater throughput at small batch sizes.}
 \end{figure}
@@
 %----------------------------------------------------------------------------------------
-%	ADDITIONAL INFORMATION
-%----------------------------------------------------------------------------------------
-
-\begin{block}{Additional Information}
-
-Maecenas ultricies feugiat velit non mattis. Fusce tempus arcu id ligula varius dictum.
-\begin{itemize}
-\item Curabitur pellentesque dignissim
-\item Eu facilisis est tempus quis
-\item Duis porta consequat lorem
-\end{itemize}
-
-\end{block}
-
-%----------------------------------------------------------------------------------------
 %	REFERENCES
@@
 \setbeamercolor{block title}{fg=red,bg=white} % Change the block title color
-
-\begin{block}{Acknowledgements}
-
-\small{\rmfamily{Nam mollis tristique neque eu luctus.
-Suspendisse rutrum congue nisi sed convallis.
-Aenean id neque dolor.
-Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.}} \\
-
-\end{block}
 

### 2018-04-24 poster/main.tex
Small changes

@@
 
-\institute{Fill in institutes} % Institution(s)
+\institute{FILL IN INSTITUTES} % Institution(s)
 
@@
 %----------------------------------------------------------------------------------------
-\begin{block}{Background}
+\begin{block}{Introduction}
 Large minibatches are necessary for computational performance but create large memory
@@
 processors in $O(T/p + \log(p))$ with the classic parallel scan algorithm. Backpropagation
-of gradient can also be parallelized with the same algorithm.
+of gradient can also be parallelized with the same algorithm. We implemented a parallel linear
+recurrence operation in CUDA and integrated it with TensorFlow.
 \end{block}
-
-\vspace{2ex}
 
@@
 \includegraphics[width=1.0\linewidth]{cumsum.png}
-\caption{Cumulative sum parallelized over 3 processors}
+\caption{Example of parallelizing cumulative sum over 3 processors}
 \end{figure}

### 2018-04-29 poster/main.tex
Add to 3rd column of poster

@@
 
-\author{Eric Martin and Chris Cundy} % Author(s)
+\author{Eric Martin and Chris Cundy*} % Author(s)
 
-\institute{FILL IN INSTITUTES} % Institution(s)
+\institute{*UC Berkeley} % Institution(s)
 
@@
 Given $x_t$, $\lambda_t$ can compute $h_t=\lambda_t h_{t-1} + x_t$ for $t=1\ldots T$ on $p$
-processors in $O(T/p + \log(p))$ with the classic parallel scan algorithm. Backpropagation
+processors in $O(T/p + \log(p))$ with the classic \textbf{parallel scan algorithm}. Backpropagation
 of gradient can also be parallelized with the same algorithm. We implemented a parallel linear
@@
 
-\begin{block}{Synthetic Long-Term Dependency}
+\begin{block}{Learning Long-Term Dependencies}
 
-  We tested the GILR-LSTM by comparing to the CuDNN LSTM implementation
-  on a synthetic memorisation task (problem 2b from %\citet{hochreiter1997long}. )
-  We compared the performance on variants with different values of
-  \(n\).  We obtained speedups of over 6x measured in wall clock time to
-  convergence.  Further, the GILR-LSTM attained convergence
-  when the time dependence of the problem had a length of \textbf{one  million time steps}.
+Problem: Learn to remember 1 bit of information for $T$ time steps.
 
- \vspace{2em}
+\vspace{1ex}
+We measured time until convergence for a 2 layer GILR-LSTM and LSTM for $T$ ranging from
+1,000 to 1,000,000.
+
+\vspace{1ex}
+The GILR-LSTM converged in over \textbf{6x} less wall time. We demonstrated a
+GILR-LSTM could learn a \textbf{one million time step} sequential dependency, which
+is at least a 100x longer dependency than previously learned.
+
+\vspace{2ex}
 \begin{figure}
 \includegraphics[width=1.0\linewidth]{8k_for_poster.png}
-\caption{Accuracy on the memorisation task with 8,192 sequence length}
+\caption{Accuracy on the memorization task with 8,192 sequence length}
 \end{figure}
@@
 
-Nunc tempus venenatis facilisis.
-\textbf{Curabitur suscipit} consequat eros non porttitor.
-Sed a massa dolor, id ornare enim.
-Fusce quis massa dictum tortor \textbf{tincidunt mattis}.
-Donec quam est, lobortis quis pretium at, laoreet scelerisque lacus.
-Nam quis odio enim, in molestie libero.
-Vivamus cursus mi at \textit{nulla elementum sollicitudin}.
+Parallel linear recurrence enables rapid learning on extremely long sequences at small
+minibatch sizes. A significant portion of deep learning's current success can be
+attributed to highly efficient matrix multiplication and convolution kernels.
+We hope that parallel linear recurrence can join these algorithms and
+be to large scale sequence
+modelling what fast convolution algorithms are to image recognition.
+
+\vspace{1ex}
+Possible future work includes parallel training of memory augmented models, applications
+to autoregressive flows, and replacing decay vector $\lambda_t$ with structured matrix
+$\Lambda_t$.
 
@@
 
-\begin{center}
-\begin{tabular}{ccc}
-\includegraphics[width=0.4\linewidth]{logo.png} & \hfill & \includegraphics[width=0.4\linewidth]{logo.png}
-\end{tabular}
-\end{center}
-
 %----------------------------------------------------------------------------------------

### 2018-04-29 poster/main.tex
minor poster rewording

@@
 be to large scale sequence
-modelling what fast convolution algorithms are to image recognition.
+modelling what fast convolution is to image recognition.
 

### 2018-04-29 poster/main.tex
add references to poster

@@
 \nocite{*} % Insert publications even if they are not cited in the poster
-\small{\bibliographystyle{unsrt}
-\bibliography{sample}\vspace{0.75in}}
+\small{\bibliographystyle{unsrt}}
+D.~Balduzzi and M.~Ghifary.
+Strongly-typed recurrent neural networks.
+
+G.~E. Blelloch.
+Prefix sums and their applications.
+
+J.~Bradbury, S.~Merity, C.~Xiong, and R.~Socher.
+Quasi-recurrent neural networks.
+
+N.~Kalchbrenner, et al.
+Neural machine translation in linear time.
+
+T.~Lei, Y.~Zhang,
+Training RNNs as fast as CNNs.
+
+A.~van~den Oord, et al.
+Wavenet: A generative model for raw audio.
 
 \end{block}
-
-%----------------------------------------------------------------------------------------
-%	ACKNOWLEDGEMENTS
-%----------------------------------------------------------------------------------------
-
-\setbeamercolor{block title}{fg=red,bg=white} % Change the block title color
-
-%----------------------------------------------------------------------------------------
-%	CONTACT INFORMATION
-%----------------------------------------------------------------------------------------
-
-\setbeamercolor{block alerted title}{fg=black,bg=norange} % Change the alert block title colors
-\setbeamercolor{block alerted body}{fg=black,bg=white} % Change the alert block body colors
-
-\begin{alertblock}{Contact Information}
-
-\begin{itemize}
-\item Web: \href{http://www.university.edu/smithlab}{http://www.university.edu/smithlab}
-\item Email: \href{mailto:john@smith.com}{john@smith.com}
-\item Phone: +1 (000) 111 1111
-\end{itemize}
-
-\end{alertblock}
 
